### 2020-04-02 llncs2e/figure1.tex
submission

\begin{figure}[t!]
\begin{tikzpicture}[
	bullet/.style={circle,
	fill,
	inner sep=2pt,
	label={#1}},
	>=latex,
	scale=0.5
]
	\draw[color=gray] (0,0) circle (85pt);
	\node[above, color=gray] at (2.5,2) {$\mathcal{S}^1$};
	\node[above] at (-5,-3) {(a)};
	\draw [black,line width=0.5pt,/.list={1/6,1/2,5/6}]
	(210:3) node[color = brick, bullet={below left:1}] (a) {}
	-- (90:3) node[color = brick, bullet={above:2}] (b) {}
	-- (-30:3) node[color = brick, bullet={below right:3}] (c) {}
	-- cycle;
	\node[above] at (-5,2) {$K_0 = \{\{1\}, \{2\}, \{3\}\}$};
	\node[above, fill=white] at (-4,1) {$K_1 = \{\{1,2\}, \{1,3\}, \{2,3\}\}$};
	\node[above, fill=white] at (-5,0) {Coefficients in $\mathbb{Z}/2\mathbb{Z}$};
\end{tikzpicture}
\begin{tikzpicture}[
	bullet/.style={circle,
	fill,
	inner sep=2pt,
	label={#1}},
	>=latex,
	scale=0.5
]
	\node[above, color=gray] at (2.5,2) {$\mathcal{S}^1$};
	\draw[color=gray] (0,0) circle (85pt);
	\node[above] at (-5,-3) {(b)};
	\draw [black,line width=0.5pt,->-/.list={1/6,1/2,5/6}]
	(210:3) node[color = brick, bullet={below left:1}] (a) {}
	-- (90:3) node[color = brick, bullet={above:2}] (b) {}
	-- (-30:3) node[color = brick, bullet={below right:3}] (c) {}
	-- cycle;
	\draw[line width=0.5pt,-{Latex[bend]}] (240:0.5) arc(240:-60:0.5);
	\node[above] at (-5,2) {$K_0 = \{\{1\}, \{2\}, \{3\}\}$};
	\node[above, fill=white] at (-3.9,1) {$K_1 = \{\{1,2\}, \{2,3\}\}, \{1,3\}\}$};
	\node[above] at (-5.7,0) {Coefficients in $\mathbb{R}$};
\end{tikzpicture}
\caption{Two triangulations of $\mathcal{S}^1$. We compute $H_0(\mathcal{S}^1) = \ker(\partial_0) / \Ima(\partial_1)$ for (a) and (b), see Sect. \ref{simcomplexhomtheory}. For (a): $\ker(\partial_0) = \text{span}(\{1\},\{2\},\{3\})$ and $\Ima(\partial_1)= \text{span}(\{\{2\}-\{1\}\},\{\{3\}-\{2\}\})$ as linearly independent elements. Observe that $\{1\}-\{3\}=\{2\}-\{1\}-(\{3\}-\{2\})$ holds. Thus $\ker(\partial_0) \simeq (\mathbb{Z}/2\mathbb{Z})^3$, $\Ima(\partial_1)=(\mathbb{Z}/2\mathbb{Z})^2$ and $H_0(\mathcal{S}^1;\mathbb{Z}/2\mathbb{Z}) = \mathbb{Z}/2\mathbb{Z}$. For (b): $\ker(\partial_0) = \text{span}(\{1\},\{2\},\{3\})$, but due to real coefficients the linear dependency is this time given by $\{2-1\} = \{3-2\}$. Thus $\ker(\partial_0) \simeq \mathbb{R}^3$ and $\Ima(\partial_1)=\mathbb{R}^2$, which yields the homology group $H_0(\mathcal{S}^1;\mathbb{R}) = \mathbb{R}$.}
\label{complexes}
\end{figure}

### 2020-04-02 llncs2e/figure2.tex
submission

\begin{figure}[t!]
\centering
\begin{tikzcd}
& M\simeq \mathbb{R}^n \arrow[ld, "x^{(0)}", bend right = 15] 
& U \simeq \mathbb{R}^{n-k} \arrow[ld, "x^{(1)}" pos=0.8, swap, bend right = 15] \arrow[d, "x^{(2)}"] \arrow[rd, "x^{(L-1)}", bend left = 15] \arrow[l, hook', swap, "\imath"] 
&\\
x^{(0)}(M) \arrow[r, "\varphi^{(0)}"] 
& x^{(1)}(U) \arrow[r, leftrightarrow, "\varphi^{(1)}"]
& x^{(2)}(U) \arrow[r, leftrightarrow, "\varphi^{(L-1)}", dotted]
& x^{(L-1)}(U) \arrow[r, hook, "\varphi^{(L)}"]
& x^{(L)}(M)
\end{tikzcd}
\caption{The architecture of the neural net maps over $\varphi^{(0)}$ into a subspace of $M$, in
which the searched manifold $U$ of the data can be embedded. The maps $\varphi^{(1)}, \cdots, \varphi^{(L-1)}$
are models of bijective, smooth and therefore diffeomorphic functions.}
\label{architecture}
\end{figure}

### 2020-04-02 llncs2e/figure3.tex
submission

\begin{figure}[t!]
\begin{tikzpicture}
\begin{axis}[
	height=2.3in,
	width= 2.7in,
	axis x line = bottom,
	axis y line = left,
	minor x tick num=10,
	ylabel={$\frac{\omega_i + \omega_{j+1}}{2}$},
	xlabel={(a) cifar10.},
	xmin=0,
	xmax=1050,
	ymin=0,
	ymax=650,
	ticklabel style={font=\small},
	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
	y label style={at={(axis description cs:0.20,0.85)},rotate=-90,anchor=south, fill=white},
	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
]
\coordinate (spypoint) at (axis cs:85,50);
\coordinate (magnifyglass) at (axis cs:800,400);
\draw [thick, dotted, draw=gray] (axis cs: 0,580) -- (axis cs: 1050,580) node[pos=0.9, below] {$\infty$};
\end{axis}
\begin{scope}[spy using outlines={circle, magnification=4, connect spies}]
\node[inner sep=0pt] at (2.55,1.9) {\includegraphics[width=0.42\textwidth]{figure3_a.pdf}};
\spy [lightgray, size=2cm] on (spypoint) in node[fill=white] at (magnifyglass);
\end{scope} 
\end{tikzpicture}\hfill
\begin{tikzpicture}
\begin{axis}[
	height=2.3in,
	width= 2.7in,
	axis x line = bottom,
	axis y line = left,
	minor x tick num=10,
	xlabel={(b) cifar100.},
	xmin=0,
	xmax=1050,
	ylabel={$\cdot 10^{3}$},
	ymin=0,
	ymax=5,
	ticklabel style={font=\small, fill=white},
	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
	y label style={at={(axis description cs:0.1,0.9)},rotate=-90,anchor=south},
	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
]
\draw [thick, dotted, draw=gray] (axis cs: 0,4.45) -- (axis cs: 1050,4.45) node[pos=0.9, below] {$\infty$};
\end{axis}
\node[inner sep=0pt] (h) at (2.55,1.9) {\includegraphics[width=0.42\textwidth]{figure3_b.pdf}};
\node[inner sep=0pt] (h) at (4.5,3.1) {{\color{blacky}$\bullet$} {$H_0$}};
\node[inner sep=0pt] (h) at (4.5,2.7) {{\color{gray}$\bullet$} {$H_1$}};
\node[inner sep=0pt] (h) at (4.5,2.3) {{\color{purpur}$\bullet$} {$H_2$}};
\node[inner sep=0pt] (h) at (4.5,1.9) {{\color{brick}$\bullet$} {$H_3$}};
\node[inner sep=0pt] (h) at (4.5,1.5) {{\color{fire}$\bullet$} {$H_4$}};
\end{tikzpicture}
\caption{Mean confident persistence landscapes of the investigated datasets cifar10 and cifar100 with a resolution of $1000$. The mean persistence landscapes were computed on truncated persistence diagrams, with a confidence of $\alpha = 0.95$ for a representative of a persistent homology class tracked on the filtration.}
\label{homgroups}
\end{figure}

### 2020-04-02 llncs2e/figure4.tex
submission

\begin{figure}[t!]
	\begin{tikzpicture}
		\begin{axis}[
		table/col sep=comma,
		xlabel={(a) {cifar10}.},
		ylabel={Validation loss},
		height=2.3in,
		width= 2.7in,
   		ticklabel style={font=\small, fill=white},
		xticklabel style={rotate=90, anchor=near xticklabel},
	    xtick={0,0.1,0.2,0.4,0.6,0.8,1},
    	xticklabels={dummy,0,50,100,150,200,250},
		]
		\end{axis}
		\node[inner sep=0pt] at (2.83,2.28) {\includegraphics[width=0.4\textwidth]{figure4_a.pdf}};
	\end{tikzpicture}
	\begin{tikzpicture}
		\begin{axis}[
		table/col sep=comma,
		xlabel={(b) {cifar100}.},
		height=2.3in,
		width= 2.7in,
		ticklabel style={font=\small, fill=white},
		xticklabel style={rotate=90, anchor=near xticklabel},
		yticklabels={,,}
	    xtick={0,0.1,0.2,0.4,0.6,0.8,1},
    	xticklabels={dummy,0,50,100,150,200,250},
		]
		\end{axis}
		\node[inner sep=0pt] at (2.83,2.28) {\includegraphics[width=0.4\textwidth]{figure4_b.pdf}};
	\end{tikzpicture}
\caption{The experiments show the mean squared error loss function on the validation dataset. We trained a total of $392$ neural nets ten times per dataset. Each line represents one neural net as an averaged loss function. The architecture can be found in Sect. \ref{subsec:invertibleneuralnets}, Appx. \ref{appx:invarch} and Fig. \ref{architecture}. The dimension of the net is chosen as a multiple of $2$, i.e. $2,4,6, \cdots, 784$. The {\color{brick} $\bullet$}--marked loss functions are parameterized above the determined threshold from Tab. \ref{stats}. The gray loss functions are below the defined threshold.}
\label{costfunctions}
\end{figure}

### 2020-04-02 llncs2e/figure5.tex
submission

\begin{table*}[t!]
\centering
\caption{The representatives from the mean confidence persistent landscape are counted per homology group. We give the minimal embedding dimension for some compact object on a connected commutative Lie group,
having approximately as many representatives of the respective homology groups. Further, we give an embedding dimension for an object consisting of all such representatives. We apply Eq. \ref{statespace} to compute the required state space dimension of resnets with different degrees of smoothness.}
\begin{tabular}{@{}rccccccccccccc@{}}
\toprule \multirow{2}{*}{\backslashbox{Data \strut}{\strut Features}} & \phantom{abc} & \multicolumn{5}{c}{Homology groups} &&& \multicolumn{5}{c}{$\approx$ Embedding dim.}\\
\cmidrule{3-7} \cmidrule{10-14} & & $H_0$ & $H_1$ & $H_2$ & $H_3$ & $H_4$ &&& $U$ & $\mathcal{C}^1$ & $\mathcal{C}^2$ & $\mathcal{C}^3$ & $\mathcal{C}^4$ \\
\midrule
cifar10 && 12 & 16 & 40 & 59 & 50 &&& $184 \pm 44$ & x & x & x & x\\
cifar100 && 13 & 18 & 34 & 46 & 48 &&& $$ & x & x & x & x \\
\midrule
Solutions of && $12$ & $16$ & $9 \pm 4$ & $8 \pm 3$ & $7 \pm 15$\\
Eq. \ref{countingsimple} \& Eq. \ref{countingequation} && $13$ & $18$ & $9 \pm 2$ & $8 \pm 10$ & $7 \pm 13$\\
\cmidrule{1-8}
\end{tabular}
\label{stats}
\end{table*}

### 2020-04-02 llncs2e/paper77.tex
submission

\documentclass[envcountsect,runningheads]{llncs}
%
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amscd,amstext}
\usepackage[utf8]{inputenc}
\makeatletter
\newcommand{\@chapapp}{\relax}
\makeatother

\usepackage{pst-node}
\usepackage{tikz,pgf}
\usepackage{tikz-cd} 
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usetikzlibrary{spy,decorations.markings,decorations.text,arrows.meta,bending,3d,shapes.arrows,positioning,fit,backgrounds,fadings}
\let\clipbox\relax

\usepackage{adjustbox}
\usepackage{mathtools}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{mathrsfs}
\usepackage{multibib}
\usepackage{filecontents}
\usepackage{slashbox}
\usepackage{multirow}
\usepackage{booktabs}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

\usepackage[hidelinks]{hyperref}
\hypersetup{
    pdftitle={Approximate Parametrization of Invertible Neural Nets up to Homology Equivalence},    % title
    pdfauthor={Luciano Melodia},     % author
    pdfsubject={Invertible Neural Nets up to Homology Equivalence},   % subject of the document
    pdfcreator={Luciano Melodia},   % creator of the document
    pdfproducer={Luciano Melodia},  % producer of the document
    pdfkeywords={Embedding Dimension} {Parametrization} {Persistent Homology} {Neural Nets} {Manifold Learning}, % list of keywords
    pdfnewwindow=true,
    colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=blue
}

\newcites{main}{References}

\definecolor{blacky}{RGB}{26,26,29}
\definecolor{gray}{RGB}{78,78,80}
\definecolor{purpur}{RGB}{111,34,50}
\definecolor{brick}{RGB}{149,7,64}
\definecolor{fire}{RGB}{195,7,63}

\newcommand{\gradient}[0]{\noindent%
    \begin{tikzpicture}
    \fill[brick, path fading=west, left color=fire, right color=brick] (0,1em) rectangle (1cm,1.5em);
    \end{tikzpicture}%
}

\tikzset{circle dotted/.style={dash pattern=on .05mm off 2mm, line cap=round}}

\let\definition\relax
\let\lemma\relax
\let\corollary\relax
\let\theorem\relax
\let\proposition\relax
\let\example\relax
\let\remark\relax

\spnewtheorem{definition}{Definition}[section]{\bfseries}{\rmfamily}
\spnewtheorem{example}{Example}[definition]{\itshape}{\rmfamily}
\spnewtheorem{theorem}{Theorem}[section]{\bfseries}{\rmfamily}
\spnewtheorem{proposition}{Proposition}[theorem]{\bfseries}{\rmfamily}
\spnewtheorem{propositionsec}{Proposition}[section]{\bfseries}{\rmfamily}
\spnewtheorem{lemma}{Lemma}[theorem]{\bfseries}{\rmfamily}
\spnewtheorem{corollary}{Corollary}[lemma]{\bfseries}{\rmfamily}
\spnewtheorem{remark}{Remark}[example]{\itshape}{\rmfamily}
\spnewtheorem{remarkdef}{Remark}[definition]{\itshape}{\rmfamily}
\spnewtheorem{remarklem}{Remark}[lemma]{\itshape}{\rmfamily}
\spnewtheorem{remarktheo}{Remark}[theorem]{\itshape}{\rmfamily}
\spnewtheorem{remarkprop}{Remark}[proposition]{\itshape}{\rmfamily}
\DeclareMathOperator{\Ima}{im}
\DeclareMathOperator{\rank}{rank}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
\renewcommand{\qed}{\hfill$\square$}
\renewcommand\UrlFont{\color{blue}\rmfamily}
\newcommand{\interior}[1]{%
  {\kern0pt#1}^{\mathrm{o}}%
}

\tikzset{->-/.style={decoration={markings, mark=at position #1 with {\arrow[line width=1pt]{>}}},postaction={decorate}}}
\begin{document}
%
\title{Parametrization of Invertible Resnets up to Homology Equivalence for Abelian Lie Groups\thanks{The work is partially supported by Siemens Power and Gas. The code is published on GitHub: \href{https://github.com/karhunenloeve/PHIA}{https://github.com/karhunenloeve/PHIA}.}}
%
\titlerunning{Invertible Resnets up to Homology Equivalence}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
\and Richard Lenz\orcidID{0000-0003-1551-4824}}
%
\authorrunning{L. Melodia et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Chair of Computer Science 6\\
Friedrich-Alexander University Erlangen-Nürnberg\\
91058 Erlangen, Germany \\
\email{\{luciano.melodia,richard.lenz\}@fau.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Neural nets have been used in an elusive number of scientific disciplines. Nevertheless, their parameterization is largely unexplored. Dense nets are the coordinate transformations of a manifold from which the data is sampled. After processing through a layer, the representation of the original manifold may change. This is crucial for the preservation of its topological structure and should therefore be parameterized correctly. We discuss a method to determine the smallest topology preserving layer for an invertible residual net. We consider the data domain as abelian connected matrix group and observe that it is decomposable into $\mathbb{R}^p \times \mathbb {T}^q$. Persistent homology allows us to count its $k$-th homology groups. Using Künneth's theorem, we count the $n$-th Betti numbers. Since we know the embedding dimension of $\mathbb{R}^p$ and $\mathcal{S}^1$, we parameterize the bottleneck layer with the smallest possible matrix group, which can represent a manifold with those homology groups. Resnets guarantee smaller embeddings due to the dimension of their state space representation.

\keywords{Embedding Dimension \and Parametrization \and Persistent Homology \and Neural Nets \and Manifold Learning.}
\end{abstract}

\section{Motivation}
Neural nets and their underlying principles were studied from the perspective of differential and information geometry \citemain{YangA97}. The natural gradient descent turned out to be Fisher optimal \citemain{Amari1,Amari2}, in its interpretation as a walk on the hypersurface of a neural loss function. Local minima, maxima and saddle points on the hypersurface have always been a problem when training machine learning models. A critical point on this surface corresponding to the global minimum of a smaller model forms a local minimum or saddle point of a larger one \citemain{FukumizuA00}. In other words, the critical values of a function represented by a layer of $(n-1)$-perceptrons will result in many critical values of a layer of $n$-perceptrons. This motivates an appropriately sized model.

The manifold of a dense neural net is in most cases Euclidean, but it is restricted by the loss function or the data manifold to certain matrix groups. A metric is required to describe the geometric properties of such a manifold. Recently \citemain{HauserR17}, the change of coordinate systems and the learned metric tensor of a Riemann manifold during back propagation of neural nets was formulated. Further, it has been shown that dense neural nets can't approximate all functions arbitrarily precisely \citemain{Johnson19}. In response to this result, neural nets with residual connections were examined. Resnets add the output of one layer to a deeper one and therefore bypass some of the layers in between. This type of neural net is an universal function approximator \citemain{LinJ18}. The depth of the net is a parameter depending on the moduli of continuity of a function to be approximated \citemain{LinJ18}. Such a residual layer defines a forward or backward difference operator on a partition of the layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \citemain{HauserGJR19}. Estimating simplicial homology in connection with Künneth's theorem we found a resnet with an approximate minimal number of perceptrons for a very special learning problem.

The idea originates from Lie theory. We use invertible architectures, to guarantee our assumptions and apply ascending sorting on the samples to yield permutation invariance or abelianization of the input domain \citemain{ArdizzoneKRK19,NguyenAK19}. Our interpretation of a neural net is a continuous transformation, transferring invariant structures of the data in the input space to its deepest layer. Thus, it is capable to reconstruct the structure of the original space. This approach corresponds to the concept of homotopy equivalence \citemain{RubinsteinR08,YangY93}, which means that there is a continuous map $f: X \rightarrow Y$ and $g: Y \rightarrow X$, such that $f \circ g$ is homotopic to the identity $\text{Id}_Y$ and $g \circ f$ to $\text{Id}_X$. Thus, the spaces can be deformed continuously into one another. To do so, their homotopy groups must be equal. The zeroth homotopy group gives information about connected components which can be collapsed to a point by continuous functions, the first homotopy group about `holes' and the second about `cavities' and so on. But homotopies are difficult to compute \citemain{klaus2004quick,hatcher2002algebraic}. Homology theory gives us a remedy that guarantees an isomorphism between homotopy groups and homology groups under the conditions of Hurewicz's Theorem \citemain{klaus2004quick}. We use theoretical results from Lie theory and persistent homology theory to predict how many dimensions of embedding are needed to represent the homology groups. Our paper is structured as follows:
\begin{itemize}
    \item In Sect. \ref{sec:manifoldassumption} we discuss the mathematical assumptions and recall the principles of smooth manifolds and Lie theory. In particular, we recall the decomposition of connected commutative Lie groups into $\mathbb{T}^p \times \mathbb{R}^q$.
    \item In Sect. \ref{sec:persistenthomology} we introduce to the theory of persistent homology, define persistence diagrams, persistence landscapes and confidence sets, in order to estimate homology groups despite noisy data. We use Künneth's theorem and derive how to count the homology groups of the input space.
    \item In Sect. \ref{sec:neuralnettheory} we explain the mathematical theory of neural nets and construct invertible deep neural nets to satisfy our assumptions. Resnets, which have a larger state space dimension, offer an architecture with a comparatively small number of parameters to dense nets.
    \item In Sect. \ref{sec:numericalexperiments} we validate our theory training auto encoders on the cifar10 and cifar100 datasets, isolating the innermost layer such that no residual connections bypass it on different embedding dimensions.
\end{itemize}

\section{The Manifold Assumption}
\label{sec:manifoldassumption}
The manifold assumption states that a point $x = \{x_1, \cdots, x_d\} \in \mathbb{R}^n$ lies on an embedded manifold $M$, with $\dim M \ll n$. We specify the concept of smooth manifolds. Then we dive into Lie groups concluding with a central theorem on the isomorphism of abelian connected Lie groups to the product space $\mathbb{T}^p \times \mathbb{R}^q$, which turns out useful to estimate the embedding dimension of the manifold.

\subsection{Background: Smooth Manifolds}
Manifolds generalize the concept of space mathematically. They are described by a collection of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to $\mathbb{R}^{n}$ where $n = \text{dim} \; M$ and an open neighborhood $(U_i,\varphi_i)$ in $M$ which is covered by this map. The union of such maps should cover the entire space, i.e. $M = \bigcup_{i \in I} U_i$ and is called \textit{atlas} $\mathcal{A} = \{(U_i,\varphi_i) \; \vert \; i \in I\}$ \citemain{lee2013smooth}. The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$.

We require that the coordinate system on $M$ is smooth. This is due to the condition that two maps of a differentiable structure must always be compatible with each other. Compatibility means, that the maps $(U,\varphi), (V,\psi)$ are diffeomorphic, such that their composition $\psi \circ \varphi^{-1}:\varphi(U \cap V) \rightarrow \psi(U \cap V)$ is a diffeomorphism between the two open subsets $\varphi(U \cap V)$, $\psi(U \cap V) \subseteq \mathbb{R}^n$, i.e. $\psi \circ \varphi^{-1}$ and $\varphi \circ \psi^{-1}$ are in $\mathcal{C}^\infty(M)$. Therefore, they are also smooth. If we take a step back and neglect smoothness for $M$, then $M$ is a topological manifold.

If we now equip $M$ with such an atlas $\mathcal{A}$ and if this atlas is maximal with respect to inclusion, then it is a \emph{differentiable structure}. Each atlas for $M$ is included in a maximal atlas. Because of the differentiable structure, the maps in $\mathcal{A}$ are also compatible with all other maps of $\mathcal{A}_{max}$ \citemain{lee2013smooth}. Thus, it is sufficient to describe smooth manifolds by an atlas that is not maximal.

\subsection{Background: Lie groups}
\label{liegroups}
A Lie group is a $\mathcal{C}^\infty$-manifold $G$ which is equipped with a group structure such that the maps $\mu:(x,y) \mapsto xy, G \times G \rightarrow G$ and $\imath: x \mapsto x^{-1}, G \rightarrow G$ are smooth. In our case $\mu$ corresponds to the matrix multiplication. The neutral element of a group is denoted by $e = e_G$. Each abelian commutative Lie group $G$ of dimension $\dim G = m$ is isomorphic to a particular product space $\mathbb{T}^p \times \mathbb{R}^q$, with $p + q = m$, see Appx. \ref{connectedcommutativeliegroupdecomposition}. Thus, the group can be decomposed into a product space with a relatively simple topological structure. The $q$-torus $\mathbb{T}^q$ is a surface of revolution. It moves a curve around an axis of rotation. These axes are given by $1$-spheres, so the $q$-torus is the product space $\mathbb{T}^q := \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$. Our initial decomposition of the Lie group into a product space can be further simplified into $G \simeq \mathbb{R}^p \times (\mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q)$. Note, that the $(n-1)$-sphere is given by $\mathcal{S}^{n-1} := \{x \in \mathbb{R}^n \; \vert \; ||x||_2 = 1\}$. Thus, $\mathcal{S}^1$ can be embedded into a two-dimensional Euclidean space. The homology groups and the embedding dimensions of the product spaces, yield by decomposition of connected commutative Lie groups, are therefore known. Any Euclidean space like $\mathbb{R}^p$ is contractible, i.e. zero-homotopic or homotopy equivalent to a constant map, thus the homology group $H_k(\mathbb{R}^p) = 0$ for $k > 0$ and $H_k(\mathbb{R}^p) = \mathbb{Z}$ for $k = 0$. To illustrate the role of orientation, we calculate the homology groups and illustrate the triangulation of the $1$-sphere, both with coefficients in $\mathbb{R}$ and in $\mathbb{Z}/2\mathbb{Z}$ in Fig. \ref{complexes}. The latter field neglects orientation. Note, that for this field $+1 = -1$.

A tool is needed to estimate homology groups of the manifold on which the data points lie. The method can then be inversely applied to the homology groups to estimate the embedding dimension of the data manifold. We introduce persistent homology with its confidence intervals in order to obtain the best possible estimate of the current state of the art by removing topological noise.

\section{Persistent Homology}
\label{sec:persistenthomology}
In algebraic topology spaces are studied as the composition of triangles, the latter being used as the fundamental component of a large class of spaces. All closed surfaces can be constructed by identifying -- or, less formally, `gluing' -- the edges of triangles \citemain[p.~102]{hatcher2002algebraic,lume}. The $n$-dimensional generalization of triangles are $n$-\emph{simplices}. They are the smallest convex set in Euclidean space with $(n+1)$-points $v_0,\cdots,v_n$, which have no solutions for any system of linear equations in $n$-variables. Thus, they do not lie on any hyperplane with a dimension less than $n$ \citemain[p.~103]{hatcher2002algebraic}. We define the $n$-simplex as follows:
\begin{equation}
    \Delta^n = \left\{ [v_0,\cdots,v_n] \in \mathbb{R}^{n+1} \; \bigg\vert \; \sum_{i=0}^{n} v_i = 1 \; \text{and} \; v_i \geq 0 \; \text{for all} \; i \right\}.
\end{equation}
Removing a vertex from the $\Delta^n$ results in a $(n-1)$-simplex called \emph{face} of $\Delta^n$. The combinatorial description of $n$-simplices is called \emph{simplicial complex}.

\subsection{Simplicial Complexes and Homology Theory}
\label{simcomplexhomtheory}
Simplicial homology is studied by $\Delta$-complexes whose simplices are uniquely determined by their vertices. Thus, every $n$-simplex has $(n+1)$ distinct vertices and no other $n$-simplex has the same set of vertices. A \emph{simplicial complex} $K$ is a collection of simplices, such that any face of a simplex of $K$ is a simplex of $K$ and the intersection of any two simplices of $K$ is either empty or a common face of both \cite[p.~11]{boissonnat2018geometric}. This is a unique combinatorial description of vertices together with a collection of sets $\{\Delta^k_1, \cdots, \Delta^k_m\}$ of $k$-simplices, which are $(k+1)$ element subsets of $K$ \citemain[p.~107]{hatcher2002algebraic}. Further, we require every $(k+1)$-element, with $k < n$ to be a $k$-simplex of $\Delta^k \subset K$. Simplicial complexes have a natural order of simplices, which is illustrated in Fig. \ref{complexes}, which plays a central role in their geometric realization \citemain[p.~14]{boissonnat2018geometric}. If we choose a total order for all corners which are located in any simplex of $K$, we get an order for each single $k$-simplex.

A \textit{singular $k$-simplex} is defined as a continuous map $\sigma: \Delta^k \rightarrow X$ into a topological space $X$. Intuitively, every simplicial complex $K$ has some connected components. These form a vector space $C_k(X)$, which is called \textit{$k$-th chain group}, with coefficients in $\mathbb{Z}$ and elements called $k$-\textit{chains}. The $k$-chains are finite sums $\sum_{i=0}^{k} z_i \sigma_i$ with $z_i \in \mathbb{Z}$ \citemain[p.~108]{hatcher2002algebraic}. Having connected components defined as a group implies also the definition of a boundary. As the boundaries of a $k$-simplex are also $(k-1)$-simplices, we can remove all other simplices to project into the lower chain group by the $k$-th boundary operator $\partial_k: C_k(X) \rightarrow C_{k-1}(X)$:
\begin{equation}
    \partial_k(\sigma) = \sum_{i=0}^{k} (-1)^{i} \sigma \bigg\vert {[v_0,\cdots,\hat{v}_i, \cdots, v_k]}.
\end{equation}
Observe that the composition of boundary maps is zero, as a boundary has no boundary by itself, i.e. $\partial_i\partial_{i+1} = 0$ \citemain[pp.~108--109]{hatcher2002algebraic}. Thus, $\ker \; \partial_i$ corresponds to the generating elements of the $i$-th boundary group and $\Ima \; \partial_{i+1}$ for the $i$-th chain group. The singular homology group is the quotient space
\begin{equation}
    H_k(X) = \frac{\ker \left(\partial_k: C_k\left(X\right) \rightarrow C_{k-1}\left(X\right)\right)}{\Ima \left(\partial_{k+1}:C_{k+1}\left(X\right) \rightarrow C_{k}\left(X\right)\right)}.
\end{equation}
The $k$-th singular homology group is isomorphic to the $n$-th simplicial homology group for a trianguliable metric space, i.e. $H_k(X) \simeq H^{\Delta}_k(X)$ for all $k \in \mathbb{Z}$, see Appx. \ref{simplicialsingularequivalence}, thus we can transfer results from one homology theory to the other.

Examining the homology groups of a set of points gives us little information about the structure of our dataset. However, since the homology groups of a simplicial complex are equal to the direct sum of the ones of all disjoint subcomplexes, see Appx. \ref{directsums}, we examine them more closely. In particular, we are interested in a parametrization of the simplicial complex as geometric realization in which the homology groups appear and disappear again. For this purpose we consider all possible subcomplexes that form a \emph{filtration} over the point cloud $X$. Depending on how we vary the parameter $\omega_i$ of the chosen simplicial complex, the following sequence is generated, starting with the empty set:
\begin{align}
	\emptyset &= K^{\omega_0}(X) \subset K^{\omega_1}(X) \subset \cdots \subset K^{\omega_{n+1}}(X) = K(X),\\
	K^{\omega_{i+1}}(X) &= K^{\omega_i}(X) \cup \sigma^{\omega_{i+1}}, \quad \text{for} \; i \in \{0, \cdots, n-1\}.
\end{align}
The filtration has a discrete realization with a fixed $\epsilon = \omega_{i+1}-\omega_{i} = \min(x,y)$ for all $x,y \in X$. We chose a sufficiently small $\epsilon$, in order to capture the birth and death of critical values in the sublevel sets of the simplicial complex upon the data, to be able to map as much structure as possible into the persistence diagram -- which clearly depends on the metric.
\input{figure1.tex}

\subsection{Persistence Diagrams}
Recording the \emph{Betti numbers} of the $k$-th homology group along the filtration, we obtain the \emph{$k$-dimensional persistence diagram}. The Betti numbers are defined by $\rank(H_k(X))$. We write $H^{\omega_i}_k = H_k(K^{\omega_i}(X))$ as $k$-th homology group on the simplicial complex $K$ with parametrization $\omega_i$. Then $H^{\omega_{i}}_{k} \rightarrow H^{\omega_{i+1}}_k$ induces a sequence of homomorphisms on the filtration, see Appx. \ref{persistenthomologyproofs} \citemain{lume,edelsbrunner2008persistent}:
\begin{align}
0 = H^{\omega_{0}}_k \rightarrow H^{\omega_{1}}_k \rightarrow \cdots \rightarrow H^{\omega_{n}}_k \rightarrow H^{\omega_{n+1}}_k = 0.
\end{align}
The image of each homomorphism consists of all $k$-dimensional homology classes which are born in $K^{\omega_i}(X)$ or appear before and die after spanning $K^{\omega_{i+1}}(X)$. Tracking the Betti numbers on the filtration results into a multiplicity
\begin{equation}
\mu^{\omega_{i},\omega_{j}}_k := (\beta_k^{\omega_{i},\omega_{j-1}} - \beta_k^{\omega_i,\omega_{j}})-(\beta_k^{\omega_{i-1},\omega_{j-1}}-\beta_k^{\omega_{i-1},\omega_{j}})
\end{equation}
for the $k$-th homology group and index pairs $(\omega_i,\omega_{j+1}) \in \overline{\mathbb{R}^2} := \mathbb{R}^2 \cup \infty$ with indices $i \leq j$ \citemain{edelsbrunner2008persistent}. The Euclidean space is extended, as the first connected component of our filtration is not merged with any other critical value. We assign the value $\infty$ to it, corresponding to the second coordinate $w_{j+1}$ in the persistence diagram. The first term counts elements born in $K^{\omega_{j-1}}(X)$ and vanished entering $K^{\omega_{j}}(X)$, while the second term counts the representatives of homology classes before $K^{\omega_{j}}(X)$ and vanishing at $K^{\omega_{j}}(X)$. The index pairs denote the \emph{birth} and \emph{death} of a homology class. The $k$-th persistence diagram is thus
\begin{equation}
\mathfrak{P}_k(K(X)) := \left\{(\omega_i, \omega_{j+1}) \in \overline{\mathbb{R}^2} \; \bigg\vert \; \mu^{\omega_{i},\omega_{j+1}}_k = 1 \; \text{for all} \; i,j \in I \right\}.
\end{equation}

\subsection{Confidence Sets}
We are interested in the homology groups of $M$ and not in those of $K(X)$. Nevertheless, the point set $X$ is drawn from a distribution, that is either on or near $M \subset \mathbb{R}^d$. We need to decide whether a persistent homology class comes from this distribution or not. For this purpose, confidence intervals for persistence diagrams have been defined \citemain{bubenik2015statistical,fasy2014confidence}. A confidence interval starts at the off-diagonal of a persistence diagram and ends parallel to it. It determines which topological features are persistent enough, i.e. persist in the filtration long enough to be recognized as such. Otherwise, they are treated as topological noise and removed.

For two compact sets $X,Y \subset \mathbb{R}^n$, the persistent homology on their filtration is stable with respect to the Hausdorff distance \citemain{fasy2014confidence}, which is defined as
\begin{align}
d_H(X,Y) := \max\left\{\sup_{x \in X} \inf_{y \in Y} d(x,y), \sup_{y \in Y} \inf_{x \in X} d(x,y) \right\}.
\end{align}
Furthermore, persistent homology is stable with respect to the bottleneck distance -- an extremal case of the Wasserstein distance -- for some $p \rightarrow \infty$:
\begin{align}
d_B(X,Y) := \inf_{\varphi: X \rightarrow Y} \sup_{x \in X} ||x-\varphi(x)||_p,
\end{align}
with $\varphi$ beeing within the set of all bijections from $X$ into $Y$, yielding
\begin{align}
	d_B\left(\mathfrak{P}\left(K\left(X\right)\right), \mathfrak{P}\left(K\left(Y\right)\right)\right) \leq c \cdot d_H\left(X,Y\right).
\end{align}
Let $\mathbb{P}$ be a distribution of a finite set of points $X$ and $x \in X$ a sample drawn from this distribution. Then we can define a confidence interval with respect to the Hausdorff distance as $\mathbb{P}\left( d_H(x,X) \leq \mu_\alpha \right) \geq 1-\alpha$ \citemain{fasy2014confidence}. A persistence confidence set $\mathscr{C}(K(X))$ is a subset of all persistence diagrams which have a certain distance from the diagonal. We look for all homology classes persisting a minimum `time':
\begin{align}
	\mathscr{C}(K(X)) := \left\{ \mathfrak{P}(K(X)) \; \bigg\vert \; c \cdot d_B\left(\text{Diag}\left(\mathfrak{P}\left(K\left(X\right)\right)\right), \mathfrak{P}\left(K\left(X\right)\right) \right) \leq \mu_\alpha \right\}.
\end{align}
Recall, that no points can appear under the diagonal line. Therefore, we measure the distance to the diagonal. For $d_B(x,y)$ imagine a box around each point with a certain radius $\mu$. If the diagonal of the persistence diagram passes through the box of a point, the point is considered as noise and removed.

\subsection{Persistence Landscapes}
The summaries of topologically persistent features -- with the purpose of computing statistics in a Hilbert space -- was initiated by Bubenik \citemain{bubenik2015statistical}. Bubenik and Dlotko first provided efficient algorithms to compute persistence landscapes \citemain{bubenikD17}. Persistence landscapes give a summary of the topology of a set of points embedded in a given topological manifold. Looking at the points $(\omega_i,\omega_{j+1}) \in \overline{\mathbb{R}^2}$ on the $k$-th persistence diagram $\mathfrak{P}_k(K(X))$, one associates a piecewise linear function $f_{(\omega_i,\omega_{j+1})}: \mathbb{R} \rightarrow [0,\infty)$ with them, which is defined as
\[
    f_{(\omega_i,\omega_{j+1})}(x) =
\begin{dcases}
    0,
    & \text{if} \; x \not\in (\omega_i,\omega_{j+1}),\\
    x - \omega_{i},
    & \text{if} \; x \in \left(\omega_{i},\frac{\omega_i + \omega_{j+1}}{2} \right],\\
    \omega_{j+1} - x 
    & \text{if} \; x \in \left(\frac{\omega_i + \omega_{j+1}}{2},\omega_{j+1} \right).
\end{dcases}
\]
We describe the summaries for the general persistence diagram, beeing the disjoint union of the $k$-th persistence diagrams $\mathfrak{P}(K(X)) := \coprod_{i=0}^{k} \mathfrak{P}_i(K(X))$. A persistence landscape $\mathfrak{L}(\mathfrak{P}(K(X)))$, contains the birth-death pairs $(\omega_i,\omega_{j+1})$, for an $i,j \in \{1, \cdots, n\}$ and is the function sequence $\lambda_k : \mathbb{R} \rightarrow [0,\infty)$ for a $k \in \mathbb{N}$, where $\lambda_k(x)$ denotes the $k$-th greatest value of $f_{(\omega_i,\omega_{j+1})}(x)$. Thus, $\lambda_k(x) = 0$ if $k > n$. In contrast to standard topological summaries, the persistence landscape lies in a completely normed vector space, suited for statistical computations.

\subsection{Counting Betti Numbers}
In the following, the homology groups are computed over $\mathbb{Z}$. Based on our assumption to operate on a connected commutative Lie group, there is an isomorphism for the homology groups of the data space $G$, so that $H_k(G) \simeq H_k(\mathbb{R}^p \times \mathcal{S}^{1}_{1}\times\cdots\times\mathcal{S}^{1}_{q})$ with $p+q = \dim G$, see \ref{decomposition}. The homology groups of this product space are equal to the homology groups of the sought manifold. The neural net should be able to preserve all structures within this space. Since the input domain is a set of points to be described by one or more compact manifolds, all homology classes should be preserved in our learning task. Künneth's theorem states that the $k$-th singular homology group of a topological product space $X \times Y$ is isomorphic to the product of the respective $k$-th homology groups of $X$ and $Y$, see \citemain[p.~268]{hatcher2002algebraic} and Appx. \ref{a:groups}:
\begin{align}
    H_k(X \times Y) \simeq \bigoplus_{i+j=k}H_i(X) \otimes H_j(Y).
\end{align}
This applies to all fields, since modules over a field are always free. If we apply the theorem to the decomposition, we get for the space we are looking for
\begin{align}
	\label{homologykunneth}
    &H_k(\mathcal{S}^{1}_{1}\times\cdots\times \mathcal{S}^{1}_{q}\times\mathbb{R}^p) \simeq \\
    \simeq \bigoplus_{i_{1} + \cdots + i_{r}=k} &H_{i_{1}}(\mathcal{S}^{1}_{1}) \otimes \cdots \otimes H_{i_{\hat{r}}}(\mathcal{S}^{1}_{q}) \otimes H_{i_{r}}(\mathbb{R}^{p}).
\end{align}
We need to know how to count in order to derive the dimension of the sought manifold from the representatives of homology classes within the mean persistence landscapes. Note, in Eq. \ref{homologykunneth} terms remain only for indices $i_{j} \in \{0,1\}$. Thus, the homology group $H_k(\mathbb{T}^{q}) \simeq H_k(\mathcal{S}^{1}_{1} \times \cdots \times \mathcal{S}^{1}_{q}) \simeq \mathbb{Z}^{{q\choose{k}}}$ and $H_0(\mathbb{R}^p)\simeq\mathbb{Z}^{p}$, where $p$ indicates the number of connected components. We denote the amount of representatives from homology classes of the $k$-th homology group appearing in the mean persistence landscape $\bar{\mathfrak{L}}$ by $\sharp^{H}_{k}(\bar{\mathfrak{L}}(\mathfrak{P}(K(X)))$. Thus, we get the following relation for the representatives of the zeroth homology group:
\begin{align}
	\label{countingsimple}
	\sharp_{0}^{H}(\bar{\mathfrak{L}}(K(X))) \approx p.
\end{align}
This is the number of connected components to be represented by the innermost layer of the neural net, corresponding exactly to the number of elements from the zeroth homology group. We only need to consider the homology groups which are represented in the mean persistence landscape. Assuming that our estimate is good, higher homology groups are trivial. The required number of factors of $1$-spheres can be determined for the $k$-th homology group by
\begin{align}
	\label{countingequation}
	\sharp_{k}^{H}(\bar{\mathfrak{L}}(\mathfrak{P}(K(X))) \approx \prod_{i=1}^{k} \frac{q-k+i}{i}
\end{align}

Note, that we assume to work on a smooth manifold, whose de Rham cohomology induces an isomorphism into the simplicial homology with coefficients in $\mathbb{R}$, see Appx. \ref{simplicialsingularequivalence}. We can still work with coefficients in $\mathbb{Z}$ because, according to the Universal Coefficient Theorem, the integral homology groups determine the homology groups over every field \cite{gruenberg1968universal}. Thus, these results can be applied without restriction to any homology theory, see Appx. \ref{simplicialsingularequivalence}.

\section{Neural Nets}
\label{sec:neuralnettheory}
Neural nets are a composition of affine transformations with a non-linear activation function. The projection on the $(l+1)$-th layer can be formalized as
\begin{equation}
\label{neuralnet}
    \varphi^{(l+1)} := \textbf{x}^{(l+1)} = \sigma(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).
\end{equation}
The linear map $\textbf{x} \mapsto \mathbf{W} \textbf{x}$ and the statistical distortion term $\textbf{x} \mapsto \textbf{x}+\textbf{b}$ is determined by stochastic gradient descent.  Note, that the linear transformation $\textbf{x} \mapsto \mathbf{W} \cdot \textbf{x}$ is interpreted as the product of a matrix $\mathbf{W}_{ij} = \mathbf{\delta}_{ij}\mathbf{W}_i$ with the input vector $\textbf{x}$ and the Kronecker delta, while $(\cdot)$ denotes elementwise multiplication.

$\sigma$ denotes a non-linear activation function. Commonly used are functions from the exponential family. A feedforward net is then a map $\varphi^{(l)}:\textbf{x}^{(l)}(M)$ $\rightarrow$ $\left(\varphi^{(l)}\circ \textbf{x}^{(l)}\right)(M)$. $\textbf{x}^{(0)} \in M$ are points on a Riemann manifold $M$. Learning by back propagation can be considered as a change of coordinate representation or shift on the tangent bundle of the underlying manifold \citemain{rumelhart1986learning}.

\subsection{Deep Learning}
\input{figure2.tex}
Neural nets are used in practice with a different number of perceptrons per layer. However, a layer, from the perspective of a matrix group, can also represent the manifold of the previous one. Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map remains constant. Thus, the dimension of the data manifold is the dimension of the smallest possible bottleneck layer of the neural net  $\dim M = \min_l \dim \textbf{x}^{(l)}$ \citemain{HauserR17}. Our contribution provides the theory of how this dimension can be determined for all neural nets that are a composition of invertible layers, i.e. $\textbf{x}^{(l)} \in \text{GL}(n,\mathbb{R}) := \{\varphi: \mathbb{R}^n \rightarrow \mathbb{R}^n \; \vert \; \det \varphi \neq 0 \}$. The concatenation of such maps is called a deep neural net:
\begin{equation}
    \textbf{x}^{(L)}(M) = \left(\varphi^{(l+1)} \circ \varphi^{(l)} \cdots \circ \varphi^{(1)} \circ \varphi^{(0)} \circ \textbf{x}^{(0)}\right)(M).
\end{equation}
Activation functions restrict to a subgroup of Euclidean space. Continuous homomorphisms between Lie groups are automatically smooth. Thus, a second countable locally Euclidean group has exactly one differentiable structure, turning it into a Lie group. We need to show that $\textbf{x}^{(l+1)}$ has the topology of $G$ and is therefore an embedded closed Lie subgroup $\textbf{x}^{(l)} := H \subseteq G$. Cartans Theorem of Closed Subgroups states that a subgroup $H$ of a Lie group $G$ is an embedded Lie subgroup, if it is a closed subgroup \citemain{cartan1952theorie}. We construct our neural nets such that they are a subset of $\text{GL}(n,\mathbb{R})$ and according to the proof in Appx. \ref{liegroups}, this is a closed subgroup of $\text{GL}(n,\mathbb{C})$, which is itself a Lie group. It follows that a layer in the neural net is also a Lie group. Due to invertibility, the subsequent smaller layers are Lie subgroups, as $\sigma^{(l)}$ is smooth.

\subsection{Residual Neural Networks}
\label{resnet}
A \textit{resnet} skips some of the layers by adding outputs to a deeper layers. It has one \textit{skip connection} and is a $\mathcal{C}^1$-continuous dynamic transformation \citemain{he2016deep,HauserGJR19}:
\begin{align}
	\label{residualnet}
	\textbf{x}^{(l+1)} = \textbf{x}^{(l)} + \sigma^{(l)}\left(\textbf{x}^{(l)}\right) \Delta l.
\end{align}
The term $\Delta l$ in Eq. \ref{residualnet} is inserted as a perturbation term. In the work of Hauser et al. \citemain{HauserR17}, neural nets were classified as $\mathcal{C}^k$-architecture, depending on how often a \textit{finite difference operator} is applied to the map $\textbf{x}: M \times I \rightarrow \mathbb{R}^d$. The forward and backward difference operators are defined as $\delta^{+}: \textbf{x}^{(l)} \mapsto \textbf{x}^{(l+1)} - \textbf{x}^{(l)}$ and $\delta^{-}: \textbf{x}^{(l)} \mapsto \textbf{x}^{(l)} - \textbf{x}^{(l-1)}$. In this theory, we can write the difference operators for a general $n \in \mathbb{N}$ with explicit specification of the perturbation $\Delta l$ as
\begin{align}
	\label{generalresidual}
	\delta^+\left( \delta^- \right)^{n-1} \textbf{x}^{(l)} = \sum_{l'=0}^{k} \left[ (-1)^{l'} {k \choose l'} \textbf{x}^{(l+1-l')} \right] = \sigma^{(l)} \left(\textbf{x}^{(l)}\right) \Delta l^k.
\end{align}
We use the same notation as in \citemain{HauserGJR19}, where $(\delta^-)^{k-1} := \delta^- \delta^- \cdots \delta^-$ denotes the application of the difference operator $(k-1)$-times and the binomial coefficient ${k \choose l'}$, with $l'$ indicating the index for summation over the layers. Using one forward difference operator and the remaining $(k-1)$ backward difference operators, the next layer $\textbf{x}^{(l+1)}$ is a function of the $k$ previous layers $\textbf{x}^{(l)}, \textbf{x}^{(l-1)}, \cdots, \textbf{x}^{(l-k+1)}$. To calculate the \textit{state space dimension} of equations of degree $k$ as in Eq. \ref{generalresidual}, we define the state as the difference operators of different transformation order of an input vector $\textbf{x}$ in the layer $l$ according to \citemain{HauserGJR19}:
\begin{align}
	\textbf{q}_n^{(l)} := \left( \delta^- \right)^{n-1} \textbf{x}^{(l)} \quad \text{for all} \; n = 1,2,\cdots,k.
\end{align}

By the recursive relation $\textbf{q}_{n+1}^{(l+1)} = \textbf{q}_n^{(l+1)} - \textbf{q}_n^{(l)}$ for the case $n = k$ we get $\textbf{q}_k^{(l+1)} - \textbf{q}_k^{(l)} = \sigma^{(l)} \left( \textbf{q}_1^{(l)}\right) \Delta l^k$ from Eq. \ref{generalresidual} \citemain{HauserGJR19}. We assume that $\textbf{q}_{n+1}^{(l+1)} = \sum_{l'=n+1}^{(k)} [\textbf{q}_{l'}^{(l)}] + \sigma\left( \textbf{q}_1^{(l)} \right) \Delta l^k$, thus the state space in $(l+1)$ is given by \citemain{HauserGJR19}:
\begin{align}
	\label{statespacepreform}
	\textbf{q}_n^{(l+1)} &= \textbf{q}_n^{(l)} + \textbf{q}_{n+1}^{(l+1)} = \textbf{q}_n^{(l)} + \sum_{l'=n+1}^k \left[\textbf{q}_{l'}^{(l)}\right] + \sigma\left( \textbf{q}_1^{(l)} \right) \Delta l^k = \\
	&= \sum_{l'=n}^k \left[\textbf{q}_{l'}^{(l)}\right] + \sigma\left( \textbf{q}_1^{(l)} \right) \Delta l^k.
\end{align}
Finally, we obtain the state space dimension for a $\mathcal{C}^k$-differentiable neural net. Considering Eq. \ref{statespacepreform}, we can observe that a layer with $d$-perceptrons, i.e. a mapping which sends $\textbf{x}^{(l)} \mapsto \mathbb{R}^d$, with a \textit{smoothness of $k$-th order} can be represented by the state space form $\textbf{q}^{(l)} := [\textbf{q}_1^{(l)},\textbf{q}_2^{(l)},\cdots,\textbf{q}_k^{(l)}]$. Thus, it maps into $\mathbb{R}^{k\cdot d}$. The more general formula for $\mathcal{C}^k$-neural nets is given by \citemain{HauserGJR19}:
\begin{align}
	\label{statespace}
	\textbf{q}_n^{(l+1)} = \sum_{l'=n}^{k} \left[\textbf{q}_{l'}^{(l)}\right] + \sigma\left( \textbf{q}_1^{(l)} \right) \Delta l^k, \quad \text{for all} \quad n = 1,2, \cdots, k.
\end{align}
A neural net with $k$-state variables is transformed by a shared activation function $\sigma^{(l)}$, which has a $(d\times d)$-parameter matrix, but according to the state space theory it is mapped to $(k\cdot d \times k \cdot d)$, which means a reduction of the parameter number by the factor $k^2$ \citemain{HauserGJR19}. We use these nets to keep the number of parameters in our neural net as small as possible and to compare the optimal state space dimension instead of the optimal embedding dimension.

\subsection{Invertible Architecture}
\label{subsec:invertibleneuralnets}
The group $\text{GL}(n,\mathbb{R})$ is the group of invertible matrices, thus we have to equip our neural net with an invertible architecture. The input vector is given by $\varphi^{(l)} := \textbf{x}^{(l)} = (\textbf{u}^{(l)}_{1},\textbf{u}^{(l)}_{1})$, splitting the $l$-th layer into $\textbf{u}^{(l)}_1$ and $\textbf{u}^{(l)}_2$ \citemain{DinhSB17,ArdizzoneKRK19}:
\begin{align}
	\textbf{u}^{(l)}_{1} &= \left(\textbf{v}^{(l)}_1 - \psi_2\left(\textbf{u}^{(l)}_2\right)\right) \cdot \exp\left(-\xi_2\left(\textbf{u}^{(l)}_2\right)\right), \\
	\textbf{u}^{(l)}_{2} &= \left(\textbf{v}^{(l)}_2 - \psi_2\left(\textbf{v}^{(l)}_1\right)\right) \cdot \exp\left(-\xi_1\left(\textbf{v}^{(l)}_1\right)\right).
\end{align}
The two input vectors $\textbf{v}^{(l)}_1$ and $\textbf{v}^{(l)}_2$ of the layer ensure that the output is trivially invertible. The multiplication is to be understood elementwise, as in Eq. \ref{neuralnet}:
\begin{align}
	\textbf{v}_1 = \textbf{u}^{(l)}_{1} \cdot \exp\left(\xi_2\left(\textbf{u}^{(l)}_{2}\right) + \psi_2\left(\textbf{u}^{(l)}_{2}\right)\right), \\
	\textbf{v}_1 = \textbf{u}^{(l)}_{1} \cdot \exp\left(\xi_2\left(\textbf{u}^{(l)}_{2}\right) + \psi_2\left(\textbf{u}^{(l)}_{2}\right)\right).
\end{align} 
We need a completely invertible neural net, therefore all layers are invertible -- except the very first one, see Fig. \ref{architecture} and Sect. \ref{sec:numericalexperiments}. The functions $\psi_i,\xi_j: \textbf{u}^{(l)}\left(M\right) \rightarrow \left(\xi_j \circ \textbf{u}^{(l)}\right)\left(M\right)$, $\textbf{u}^{(l)} \mapsto \sigma\left(W^{(l+1)} \cdot \textbf{u}^{(l)} + \textbf{b}^{(l+1)}\right)$ are implemented as neural subnets by themselves. These nets do not have to be invertible in general.

\section{Numerical Experiments}
\label{sec:numericalexperiments}
\input{figure3.tex}
We train an autoencoder and use as input heavily noisy images and map them to their noiseless original. It is not relevant whether the mean persistence landscapes are computed over the noisy data or over the originals, since the primer can deal with noise. In principle, the persistence landscapes should be calculated for the data manifold. We use invertible $\mathcal{C}^1$-nets and isolate the innermost layer of the autoencoder, see Sect. \ref{resnet} and Fig. \ref{architecture}. To test if our prediction for the embedding dimension, we parameterize the innermost layer using two perceptrons. The nets are trained and evaluated according to the caption of Fig. \ref{costfunctions}.

\paragraph{Persistence Landscapes} The persistence landscapes are used for prediction, computed on the persistence diagrams of the datasets cifar10 and cifar100 \citemain{krizhevsky2009learning}, see Fig. \ref{homgroups}. We use the $\alpha$-complex as filtration, see Melodia et al. \citemain{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is $10$. The maximal edge length is set to $0.1$. All curves in the mean persistent landscapes are smoothened by a Gaussian filter with $\sigma = 2$.
\input{figure5.tex}

\paragraph{Neural Nets} The data manifold is supposed to be a commutative Lie group. Thus, we artificially generate commutativity by sorting input and target vectors beforehand. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.8 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$. These input vectors are trained to the original vector $\textbf{x}^{(0)}$. We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = \frac{1}{n} \sum_{i=1}^{n} \left(\textbf{x}^{(0)}_i-\textbf{y}_i\right)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = \frac{1}{1 + e^{-\mathbf{x}^{(l)}}}$ is applied as elementwise activation function throughout the net. In all experiments dense neural nets are used with a bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}\left(x^{(0)} \right) =$ $\sigma\left( \sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)} \right)$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as $\mathcal{C}^1$-residual invertible layers, see Sect. \ref{resnet} and Sect. \ref{subsec:invertibleneuralnets}. A total of $5$ of these hidden layers are used, all with the same embedding dimension. More details can be found in Appx. \ref{appx:numexp}.

\subsection{Results}
Tab. \ref{stats} shows the representatives of homology classes from their respective groups for cifar10 and cifar100. Further, it shows the estimated minimal embedding dimension for a manifold with these homology groups for $\mathcal{C}^1$ to $\mathcal{C}^4$-resnets. The representatives for the $k$-th homology group $\sharp_k^{H}\left(\bar{\mathfrak{L}}\left(\mathfrak{P}(K(X))\right)\right)$ and the corresponding dataset are counted. Then, we seek for an embedding dimension to represent a manifold with the number of spheres $q$ with these homology groups. A Monte Carlo approach is used to find integer solutions of Eq. \ref{countingequation}. All models below our dimensional treshold show explosions of the gradient, see Fig. \ref{costfunctions}. The dimension of the respective loss function can be identified by the transition \gradient{}, from the lowest tested dimension to the highest, see Fig. \ref{costfunctions}. The models above the treshold remain stable. Higher parameterized tested models are stable. A single net still shows an explosive gradient in Fig. \ref{costfunctions} (b). This could be due to statistical errors in parametrization by mean confidence persistence landscapes. The net showing the deflection of gradient is close to our threshold. We specify the parameterization for an embedding as well as for neural nets of the smoothness degree $1$--$4$. The noise was added after sorting. We can confirm that the neural net is capable of eliminating the additional structure of noise, which is no longer part of the abelian Lie group.
\input{figure4.tex}

\section{Conclusions}
\label{sec:conclusion}
Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to abelianization of the input domain. This forms a solution for a special case of learning problems, allowing to restrict the domain to an almost commutative group. The initial problem could be simplified by decomposing the Lie group into a product space $\mathbb{R}^p \times \mathbb{T}^q$. The homology groups of the $q$-torus and the Euclidean space are well known. The application of fully invertible neural nets to resnets allows to meet the requirements for working with the group of invertible matrices. By applying Künneth's theorem the homology groups of the topological factor spaces could be connected with the ones of their product space. Using mean confident persistence landscapes the homology groups were estimated. A treshold was computed, see sect. \ref{sec:numericalexperiments}, which is determined by the state space dimension of the resets according to a representation up to homology equivalence. The mean persistence landscapes are particularly useful for short-lived isolated collections of data points, which in particular generate elements from homology classes at the beginning of the filtering and become visible at the lower left corner of the persistence landscapes, see Fig. \ref{homgroups} (a). The theoretical work was tested with numerical experiments on toy data. We do not limit the processing of the neural net within the layers to an abelian group. This can cause inaccuracies in the estimation of the dimensional treshold. The explicit abelianization of the invertible resnets solves this problem.
%
%
%
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystylemain{splncs04}
\bibliographymain{biblio}
%
\newpage
\appendix
\section{Supplementary Material for Sect. \ref{sec:manifoldassumption}}
\label{a:mannilliegroup}
\subsection{Groups}
\label{a:groups}
\begin{definition}
A set $G$ together with a map $\mu: G \times G \rightarrow G$, $(x,y) \mapsto xy$ and an element $e = e_G$ is called group if the following conditions are fulfilled:
\begin{enumerate}
	\item[(1)] $(xy)z = x(yz)$ for all $x,y,z \in G$, thus the elements are associative,
	\item[(2)] $xe = ex = x$ for all $x \in G$, so there is a neutral element and
	\item[(3)] $\forall x \in G \; \exists \; x^{-1} \in G$, such that $xx^{-1} = x^{-1}x = e$. 
\end{enumerate}
\end{definition}

\begin{definition}{\cite[p.~196]{hilton2012course}}
Let $G$ be a group and contain a subset $H \subseteq G$. $G$ is called free group with basis $B$, if for each $H$ and each function $\psi: B \rightarrow H$ a unique homomorphism $\varphi: G \rightarrow H$ with $\varphi(x) = \psi(x)$ exists for all $x \in B$.
\end{definition}

\begin{example}
The most prominent example for a free group are the integers with addition $(\mathbb{Z},+)$ and basis $B = \{1\}$. Consider $\psi: B \rightarrow G$ with $g = \psi(1)$ and $g \in G$ and $\varphi(x) = g^n$ for each $n \in \mathbb{Z}$. Then $\varphi$ is the only group homomorphism of $(\mathbb{Z},+)$ into $G$, such that $\varphi(1)=\psi(1)$.
\end{example}

\begin{definition}
Let $\textbf{X} = \{X,\varphi_i\}_{i \in I}$ and $\textbf{Y} = \{Y,\psi_i\}_{i \in I}$ be two algebraic structures of the same type $\sigma = \{n_i\}_{i\in I}$, such that for each $i \in I$ and $n_i \in \mathbb{N}$, the arity of the fundamental operations is denoted with $\varphi_i$ and $\psi_i$. The projection $\varphi: \textbf{X} \rightarrow \textbf{Y}$ is called homomorphism, if for all $x_{i} \in X$
\begin{equation}
\varphi\left(f_i\left(x_{1},\cdots,x_{n}\right)\right) = g_i\left(\varphi\left(x_{1}\right),\cdots,\varphi\left(x_{n}\right)\right).
\end{equation}
\end{definition}

\begin{example}
The function $\varphi: \mathbb{Z} \rightarrow \mathbb{Z}/2\mathbb{Z}$ is a homomorphism of groups. The group $(\mathbb{Z}/2\mathbb{Z},+)$ is the quotient space that assembles all even integers to zero and all odd integers to one, where the respective equivalence class is defined.
\end{example}

\begin{remark}
Recalling group homomomorphisms and isomorphisms helps to gain proper intuition on chain and boundary groups. Isomorphisms from one algebraic structure to another exist if these two structures are considered to be essentially `equal'. A homomorphism is weaker. It preserves `some' of the structure. Since isomorphisms are homomorphisms, this preserved structure can be regarded as a kind of `subset'. Homomorphisms preserve operations, but allow elements of one algebraic structure to collapse into some elements of the other. 
\end{remark}

\begin{definition}
A subgroup $H \subseteq G$ is a group if a homomorphism $\varphi: G \rightarrow H$ exists, which maps $\varphi(e_G)$ to $e_H$ and $\varphi(xy)$ to $\varphi(x)\varphi(y)$ for all $x,y \in G$. A subgroup is called \emph{normal} if $xHx^{-1} = H$, for all $x \in G$.
\end{definition}

\begin{example}{\cite[p.~5]{van2003lie}}
Consider $\varphi:G\rightarrow H$ as a surjective group homomomorphism. Then $K := \ker \varphi$ is a normal subgroup of $G$. There is a unique map $\overline{\varphi}: G/K \rightarrow H$, which yields $\overline{\varphi} \circ \pi = \varphi$, with the canonical projection $\pi$. The \emph{factor map} $\overline{\varphi}$ is an isomorphism of groups.
\end{example}

\begin{proof}
Let $x \in G$ and $k \in K$. Then $\varphi(xkx^{-1})$ $=$ $\varphi(x)\varphi(k)\varphi(x^{-1})$ $=$ $\varphi(x)e_H\varphi(x)^{-1}$ $=$ $e_H$, thus $xkx^{-1}$ is an element of $\ker \varphi = K$. It follows that $xKx^{-1} \subset K$. Due to the properties of the neutral element $x^{-1}Kx \subset K$, i.e. $K \subset xKx^{-1}$, we get $K = xKx^{-1}$. \qed
\end{proof}

\begin{remark}
Similar as subsets behave to sets, \emph{subgroups} behave to groups. 
\end{remark}

\subsection{Lie groups}
\begin{definition}
Let $G$ be a commutative abelian Lie group. The exponential map $\exp := \exp_G : T_eG \rightarrow G$ maps from the tangent space $T_eG$ in $e$ into the group $G$. It also maps in every point into $G$ due to homogeneity. It is defined as $\exp(\mathfrak{X}) = \alpha_{\mathfrak{X}} (1)$ for a left invariant vector field $\mathfrak{X} \in T_eG$ and a maximal integral curve $\alpha_{\mathfrak{X}}$ with a starting point $e$. For all $s,t \in \mathbb{R}$ and each $\mathfrak{X} \in T_eG$ the exponential map fulfils the properties
\begin{itemize}
	\item[(a)] $\exp(s\mathfrak{X}) = \alpha_{\mathfrak{X}}(s)$,
	\item[(b)] $\exp(s+t) \mathfrak{X} = \exp(s\mathfrak{X}) \exp(t \mathfrak{X})$.
\end{itemize}
\end{definition}

\begin{proof}
Let $\gamma(t) = \alpha_{\mathfrak{X}}(st)$ be a curve. Then $\gamma(0) = e$ and
\begin{align}
	\frac{d}{dt}\gamma(t) = s\dot{\alpha}_{\mathfrak{X}}(st) = s v_{\mathfrak{X}}(\alpha_{\mathfrak{X}}(st)) = v_{s\mathfrak{X}}(\gamma(t)).
\end{align}
Let $\mathfrak{X} \in T_eG$, then the integral curve $\alpha_{\mathfrak{X}}$ is defined on $\mathbb{R}$. Let $y \in G$ and $\alpha_1(t) = y\alpha(t)$. Differentiating with respect to $t$ yields
\begin{align}
	\frac{d}{dt}\alpha_{1}(t) = T_{\alpha(t)}y \frac{d}{dt} \alpha(t) = T_{\alpha(t)} y v_{\mathfrak{X}}(\alpha(t)) = v_{\mathfrak{X}}(\alpha_1(t))
\end{align}
by the left invariance of $v_{\mathfrak{X}}$. So $\alpha_1$ is also an integral curve for $v_{\mathfrak{X}}$. Let $I$ be the domain of $\alpha_{\mathfrak{X}}$ and $t_1 \in I$ with $x_1 = \alpha_{\mathfrak{X}}(t_1)$. Then $\alpha_1(t) := x_1 \alpha_{\mathfrak{X}}(t)$ is an integral curve for $v_{\mathfrak{X}}$ with starting point $x_1$ given by $\alpha_2: t \mapsto \alpha_{\mathfrak{X}}(t+t_1)$. The integral curve has domain $I-t_1$. We conclude that $I \subset I-t_1$. It follows that $s+t_1 \in I$ for all $s, t_1 \in I$ so $I$ must be equal to $\mathbb{R}$.

For a fixed $x \in \mathbb{R}$ we get $\gamma: t \mapsto \alpha_{\mathfrak{X}}(s)\alpha_{\mathfrak{X}}(t)$ as maximal integral curve for $v_{\mathfrak{X}}$ with starting point $\alpha_{\mathfrak{X}}(s)$. On the other hand, the same applies to $\epsilon: t \mapsto \alpha_{\mathfrak{X}}(s+t)$. Thus $\gamma = \epsilon$ and we get
\begin{align}
	\exp((s+t)\mathfrak{X}) = \alpha_{\mathfrak{X}}(s+t) = \alpha_{\mathfrak{X}}(s) \alpha_{\mathfrak{X}}(t) = \exp(s\mathfrak{X}) \exp(t\mathfrak{X}).
\end{align} \qed
\end{proof}

\begin{remarkdef}
The exponential map is a smooth map $\exp_G: T_eG \rightarrow G$ and a local diffeomorphism around $0$.
\end{remarkdef}

A Lie group $G$ is also a smooth manifold, as defined in Sect. \ref{liegroups}. Its tangential space is a linear space, which we denote as $\mathfrak{g}$.

\begin{definition}
Each linear space $\mathfrak{g}$ equipped with a bilinear form $[\cdot,\cdot]: \mathfrak{g} \times \mathfrak{g} \rightarrow \mathfrak{g}$ is called Lie algebra, if the bilinear form, also called Lie bracket, is 
\begin{itemize}
	\item[(1)] antisymmetric, i.e. $[\mathfrak{X}, \mathfrak{Y}] = -[\mathfrak{Y}, \mathfrak{X}]$ and the
	\item[(2)] Jacobi identity holds $[\mathfrak{X},[\mathfrak{Y},\mathfrak{Z}]] = [\mathfrak{Y},[\mathfrak{Z},\mathfrak{X}]] + [\mathfrak{Z},[\mathfrak{X},\mathfrak{Y}]]$, for all $\mathfrak{X}, \mathfrak{Y}, \mathfrak{Z} \in \mathfrak{g}$.
\end{itemize}
\end{definition}

\begin{remarkdef}
For a Lie group $G$ with subgroup $H$ both properties are equivalent \cite[pp.~25-27]{van2003lie}.
\end{remarkdef}

\begin{propositionsec}{\cite[p.17]{van2003lie}} Let $G$ be a Lie group and $H$ a subgroup. Then the following properties are equivalent:
\begin{enumerate}
	\item[(1)] There is an open neighborhood $U$ of $e$ such that $U \cap H = \{e\}$.
	\item[(2)] The group $H$ is discrete.
	\item[(3)] For each compact subset $C \subset G$ the intersection $C \cap H$ is finite.
	\item[(4)] The group $H$ is a closed Lie subgroup with Lie algebra $\{0\}$.
\end{enumerate}
\end{propositionsec}
\begin{proof}Take any element $h \in H$. Then $hU$ is an open neighbourhood of $h$ in $G$. $hU \cap H = h(U\cap h^{-1}H)$ and from the group structure of $H$ it follows that $h(U\cap h^{-1}H) = h(U \cap H) = \{h\}$.

Choose $U$ as an open neighborhood centered around $e = e_G$, such that $U \cap H = \{e\}$. Let $g \in G$ be a point in the closure $\overline{H}$. Then there is a sequence $\{h_i\}$ in $H$ which converges to $g$. It follows that $h_{i+1}h_i^{-1} \rightarrow gg^{-1} = e$, with $i \rightarrow \infty$. For an $i$ large enough, $h_{i+1}h_i^{-1} \in U \cap H = \{e\}$, i.e. $h_i = h_{i+1}$. The sequence $h_i$ becomes stationary after a certain index. Therefore $h_i = g$ for some $i$. We conclude that $g \in H$ for any $g$. Further we see that $\overline{H} \subset H$ and therefore $H$ is closed. Since $H$ is closed, the intersection $H \cap C$ is also closed, thus compact. For a $c \in H \cap C$ we choose an open neighborhood $cU$ in $G$, such that $cU \cap H = \{c\}$. Then $\{cU \; | \; c \in H \cap C\}$ is an open cover of $H \cap C$ with no proper subcover. Since $H \cap C$ is compact, the cover must be finite and thus $H \cap C$ is finite.

Let $g \in G$ a point in the closure $\overline{H}$. We choose a compact neighborhood $C$ around $g$, such that $g$ lies in the closure $\overline{H \cap C}$. The latter is finite, thus closed. Therefore $g$ is in the closure $\overline{H \cap C} \subset H$ and contained in $H$. We conclude that $H$ is a closed Lie subgroup. Let $\mathfrak{g}$ be the Lie algebra of $H$ consisting of $\mathfrak{X} \in \mathfrak{g}$ with $\text{exp}(\mathbb{R}\mathfrak{X}) \subset H$. A local diffeomorphism $\text{exp}: \mathfrak{g} \rightarrow G$ to $0$ gives an open neighborhood $\mathfrak{o}$ around $0$ in $\mathfrak{g}$. The exponential function is injective on $\mathfrak{o}$. If we take $\mathfrak{X} \in \mathfrak{g} \setminus \{0\}$, then there is an $\epsilon > 0$, such that $[-\epsilon, \epsilon]\mathfrak{X} \subset \mathfrak{o}$. We define a curve $\gamma: [\epsilon,\epsilon] \rightarrow G$, $t \mapsto \text{exp}(t\mathfrak{X})$, with compact image. Thus $\Ima(\gamma) \cap H$ and $\{t \in [-\epsilon,\epsilon] \; | \; \text{exp}(t \mathfrak{X}) \in H\}$ are finite. Hence, $\mathfrak{X} \notin \mathfrak{h}$ and $\mathfrak{h} = \{0\}$.

$H$ is a closed smooth submanifold of $\mathfrak{X}$, where $\dim H = 0$. It follows that an open neighborhood $U$ exists from $e$ to $G$, such that $U \cap H = \{e\}$. \qed
\end{proof}

\subsection{Decomposition of Connected Commutative Lie groups}
\label{decomposition}
\begin{theorem}{\cite[p.~116,~p.~26]{onishchik1993lie,van2003lie}}
\label{connectedcommutativeliegroupdecomposition}
Every connected commutative (or in other words abelian) Lie group $G$ is isomorphic to some product space $\mathbb{T}^p \times \mathbb{R}^q$ with $\dim G = p+q$, i.e. $G \simeq \mathbb{T}^p \times \mathbb{R}^q$.
\end{theorem}

\begin{proof}
Choosing $G$ as a connected abelian Lie group, the Lie algebra $\mathfrak{g}$ is commutative, such that $[\mathfrak{X}, \mathfrak{Y}] = 0$ for all $\mathfrak{X}, \mathfrak{Y} \in \mathfrak{g}$. It follows that the exponential map $\text{exp}(\mathfrak{X})\text{exp}(\mathfrak{Y}) = \text{exp}(\mathfrak{X}+\mathfrak{Y})$ is given for all $\mathfrak{X}, \mathfrak{Y} \in \mathfrak{g}$. Thus $\text{exp}: \mathfrak{g} \rightarrow G$ is a Lie group homomorphism from $(\mathfrak{g},0,+)$ to $G$. It follows that $\text{exp}(\mathfrak{g})$ is a subgroup of $G$, thus $\text{exp}(\mathfrak{g}) = G_e$. Since $G$ is connected, $G = G_e$. Therefore $\Ima \text{exp}(\mathfrak{g}) = G$ and $\text{exp}: \mathfrak{g} \rightarrow G$ is a surjective Lie group homomorphism. If $\Omega$ is the closed subgroup $\text{ker}(\text{exp})$ of $\mathfrak{g}$, then according to the Isomorphism Theorem for groups: $G \simeq \mathfrak{g}/\Omega$ \cite[pp.~5-7]{van2003lie}. Since $\text{exp}$ is a local diffeomorphism at $0$, there is an open neighborhood $\Gamma$ of $0$ in $\mathfrak{g}$ for which $\text{exp}$ is an injective map. It follows that $\Gamma \cap \Omega = \{0\}$. Thus $\Omega$ is a discrete subgroup of $\mathfrak{g}$.

We need a corresponding basis for a discrete group consisting of finite-dimensional linear independent elements with coefficients in $\mathbb{Z}$.

\begin{lemma}{\cite[p. 27]{van2003lie}}
A subset of a finite dimensional linear space $V$ can be written by linear independent components $\omega_1, \cdots, \omega_p$ of $V$ as a direct sum with coefficients in $\mathbb{Z}$, such that $\Omega = \mathbb{Z}\omega_1 \oplus \cdots \oplus \mathbb{Z}\omega_p$.
\end{lemma}

\begin{proof}
We show that by induction and start with $\dim V = 1$. Choosing a basis, we identify $V$ with $\mathbb{R}$. Then $\Omega$ is a discrete subgroup of $\mathbb{R}$. We assume that $\Omega \neq \{0\}$, then there is an element $x \in \Omega \setminus \{0\}$ with $x > 0$. The section $\Omega \cap [0,x]$ is finite and therefore contains a smallest element $\omega$. The section $\Omega \; \cap \; ]0,1[\omega$ is equal to the empty set, thus $\mathbb{Z} \omega \subset \Omega$. Next we consider an element $y \in \Omega$, then $y \not\in \mathbb{Z}\omega$ would mean that $y \; \in \; ]n,n+1[\omega$ for a corresponding $n \in \mathbb{Z}$. This means that $y-nx \in \Omega \; \cap \; ]0,1[ \omega = \emptyset$. This in turn implies $\Omega \subset \mathbb{Z}\omega$. Thus $\Omega = \mathbb{Z}\omega$.

For the general case we use the idea from van den Ban's proof \cite[p.27]{van2003lie}. We assume that $\dim V = n$ for any $n > 1$. In the trivial case $p = 0$ and $\Omega = \{0\}$ holds. For $\omega \in \Omega \setminus \{0\}$ the intersection $\mathbb{R} \omega \cap \Omega$ is discrete in $\mathbb{R} \omega$, i.e. an integer multiple $\mathbb{Z}\omega_1$. We choose a linear subspace $U \subset V$, in order that $\mathbb{R}\omega_1 \oplus U = V$, with the canonical projection $\pi: U \rightarrow V$ and $C \subset U$ as compact subset:
\begin{align}
\pi^{-1}\left(C\right) &= \mathbb{R} \omega_1 + C = \left([0,1] \omega_1 + C\right) + \mathbb{Z}\omega_1,\\
\implies \pi\left(\Omega\right) \cap C &= \pi\left(\left(C + [0,1] \omega_1\right) \cap \Omega + \mathbb{Z}\omega_1\right)\\
&= \pi\left(\left(C + [0,1]\omega_1\right) \cap \Omega\right). 
\end{align}
The set $C$ is compact, so is $C + [0,1] \omega_1$. It follows that the intersection $(C + [0,1] \omega_1) \cap \Omega$ is also compact. Thus $C$ is a discrete compact subset of $U$. By induction there are linearly independent elements $\overline{\omega}_2, \cdots, \overline{\omega}_p$ of $\pi(\Omega)$, such that $\pi(\Omega) = \mathbb{Z} \overline{\omega}_2 \oplus \cdots \oplus \mathbb{Z}\overline{\omega}_p$. We choose $\pi(\omega_i) = \overline{\omega}_i$, then the elements $\omega_1, \cdots, \omega_p$ are linearly independent and we get the basis $\Omega = \mathbb{Z}\omega_1 \oplus \cdots \oplus \mathbb{Z}\omega_p$. \qed
\end{proof}

Since $\mathfrak{g}$ is a linear space, we can extend the basis $\mathbb{Z}\omega_1 \oplus \cdots \oplus \mathbb{Z} \omega_n$ to the whole Lie algebra $\mathfrak{g}$. Thus we get a linear isomorphism $\varphi: \mathfrak{g} \rightarrow \mathbb{R}^n$, where $\mathbb{R}^n \simeq \mathbb{R}^p \times \mathbb{R}^q$ with $p+q = n$. The map $\text{exp} \circ \varphi^{-1}: \mathbb{R}^n \rightarrow G$ is a surjective Lie group homomorphism, i.e. a group homomorphism that is smooth, with smooth inverse. According to the Isomorphism Theorem for groups, this group also factorizes by a bijective group homomorphism \cite[pp.~5-7]{van2003lie}:
\[
\begin{tikzcd}[column sep=huge]
    \displaystyle \mathfrak{g}
    \arrow[swap]{r}{\varphi}
    \arrow[bend left=25]{rr}{\text{exp}}
    \arrow[swap]{dr}{\pi \; \circ \; \varphi}
&   \mathbb{R}^p \times \mathbb{R}^q
    \arrow{d}{\pi}
    \arrow[swap]{r}{\text{exp} \; \circ \; \varphi^{-1}}
&   G\\	
&   (\mathbb{R}/\mathbb{Z})^p \times \mathbb{R}^q 
	\arrow[swap]{ur}{\overline{\text{exp} \; \circ \; \varphi^{-1}}}
\end{tikzcd}
\]
The canonical map $\pi: \mathbb{R}^n \simeq \mathbb{R}^p \times \mathbb{R}^q \rightarrow (\mathbb{R}/\mathbb{Z})^p \times \mathbb{R}^q$ is locally diffeomorphic and $\overline{\text{exp} \circ \varphi^{-1}} \circ \pi$ as well. Due to the homogeneity of commutative lie groups, the map is everywhere a diffeomorphism. The projection $\overline{\text{exp} \; \circ \; \varphi^{-1}}$ is bijective, thus also a diffeomorphism and thus an isomorphism of Lie groups. By definition $\mathbb{R}^n/\mathbb{Z}^n \simeq (\mathbb{R}/\mathbb{Z})^n \simeq \mathbb{T}^n$ and the proof follows. \qed
\end{proof}

\section{Supplementary Material for Sect. \ref{sec:persistenthomology}}
\label{persistenthomologyproofs}
\subsection{Properties of Simplicial Complexes}
\begin{definition}
A simplicial complex $K$ is \textit{connected} if it cannot be composed of a disjoint union of two or more non-empty subcomplexes.
\end{definition}

\begin{definition}
An embedded simplicial complex is \emph{path-connected} if there is a path from any vertex to any other as a composition of $1$-simplices.
\end{definition}

\begin{propositionsec}
A simplicial complex $K$ is path-connected if it is connected and vice versa. Thus the notions are equivalent for simplicial complexes.
\end{propositionsec}

\begin{proof}
Assume $K$ is path-connected and not connected. Choose two subcomplexes $U,V \subset K$ with $U \cup V = K$. Choose a vertex $u_0 \in U$ and $v_0 \in V$. Thus a path from $u_0$ to $v_0$ exists, due to the path-connectedness of the simplicial complex. If $u_i$ is the last vertex in $U$, then the vertex $u_{i+1}$ can be contained neither in $U$ nor in $V$, due to the disjoint decomposition of the simplicial complex. This contradicts the assumption that it is path-connected.

Consider two vertices $u,v \in K$ that are not path-connected. Define $U,V \subset K$ as path-connected subcomplexes containing $u\in U$ and $v \in V$.  Choose $w \in U \cap V \neq \emptyset$, then there is a path from $u$ to $w$ and also from $w$ to $v$. A composition of paths gives a path from $u$ to $v$, which contradicts the assumption that $u$ and $v$ are not path-connected. Thus $U \cap V = \emptyset$ and $K$ is not connected.\qed
\end{proof}

\subsection{Direct Sums of Homology Groups}
\label{directsums}
\begin{theorem}{\cite[p.~109]{hatcher2002algebraic}}
The $k$-th simplicial homology group $H^{\Delta}_n(K(X))$ of a simplicial complex $K$ for a field $\mathbb{F}$ is isomorphic to the direct sum of the homology groups of all subcomplexes on its filtration, i.e. $H^{\Delta}_n(K(X)) \simeq \bigoplus_{i \in I} H^{\Delta}_n(K^{\omega_{i}}(X))$.
\end{theorem}

\begin{proof}
Let $\{K^{\omega_{i}}\}_{i \in I}$ be the set of all connected components of the simplicial complex $K(X)$ of a point set $X$ and $H^{\Delta}_k\left(K\left(X\right)\right)$ its simplicial homology groups. We denote the homology groups of the components with $H_k^{\Delta}(K^{\omega_{i}}(X))$. With $C_k(K^{\omega_{i}}(X))$, we denote the group of $k$-chains for the connected component of the simplicial complex. Then $C_k(K^{\omega_{i}}(X)) \subseteq C_k(K(X))$ is a subgroup for all $i$.

\begin{lemma}
The group of $k$-chains can be written as a direct sum of the connected components, i.e. $C_k(K(X)) \simeq \bigoplus_{i \in I} C_k(K^{\omega_{i}}(X))$. This also applies to the lower chain groups.
\end{lemma}

\begin{proof}
Using the boundary operator, which is a homomorphism of groups, we get $C_{k-1}(K(X)) \simeq \bigoplus_{i \in I} \partial_k(C_{k}(K^{\omega_{i}}(X)))$. An element in the $(k+1)$-th chain group $v \in C_{k+1}(K(X))$ is represented by $v = \sum_{i \in I} v_{\omega_{i}}$. Thus $\partial_{k+1}(v) \in C_{k}(K(X))$ can be broken down to $\partial_{k+1}(v) = \sum_{i \in I} \partial_{k+1}(v_{\omega_{i}})$ where $v_{\omega_{i}} \in C^\omega_{k+1}(K^{\omega_{i}})$.\qed
\end{proof}

The groups of coundaries can also be decomposed. The kernel of $C_k(K^{\omega_{i}}(X))$ can be written as $\ker \partial_k(C_k(K^{\omega_{i}}(X)))$ $= \ker \partial_k(C_k(K(X)))$ $\cap$ $C_k(K^{\omega_{i}}(X))$.
\begin{align}
	\ker \partial_k &\simeq \bigoplus_{i \in I} \partial_k\left(C_k\left(K^{\omega_{i}}(X)\right)\right), \text{which implies}\\
	&\Rightarrow \frac{\ker \partial_k\left(C_k\left(K\left(X\right)\right)\right)}{\Ima \partial_{k+1}\left(C_{k+1}\left(K\left(X\right)\right)\right)} \simeq \bigoplus_{i \in I} \frac{\ker \partial_k(C_k(K^{\omega_{i}}(X)))}{\Ima \partial_{k+1}(C_{k+1}(K^{\omega_{i}}(X)))}\\
	&\Rightarrow H^{\Delta}_k(K(X)) \simeq \bigoplus_{i \in I} H^{\Delta}_k(K^{\omega_{i}}(X)).
\end{align}
\qed
\end{proof}

\begin{remarklem}
We obtain an isomorphism for the singular homology groups from the homology groups of the set of path components. This can be thought of as a direct sum of all `loops' from every possible path component in space.
\end{remarklem}

\begin{definition}
A \emph{path component} of a topological space $X$ is an equivalence class under the equivalence relation $x \sim y$ if there is a path from $x$ to $y$.
\end{definition}

\begin{remarkdef}
Since for simplicial complexes, the property to be connected corresponds to path-connectedness, the coordinatewise sum of the connected components is a generator of the whole group.
\end{remarkdef}

\begin{theorem}{\cite[p.~129]{hatcher2002algebraic}}
Let $X$ be a topological space and $\{X^{\omega_{i}}\}_{i \in I}$ a set of path components of $X$ and $\imath_{\omega_{i}}: X^{\omega_{i}} \hookrightarrow X$ the inclusion. Then the maps $(\imath_{\omega_{i}})_\star: H_k(X^{\omega_{i}}) \hookrightarrow H_k(X)$ for each $k \geq 0$ induce an isomorphism between the direct sum of the homology groups of all simplicial subcomplexes and the homology groups of $X$:
\begin{equation}
	\bigoplus_{i \in I} H_k(X^{\omega_{i}}) \rightarrow H_k(X).
\end{equation}
\end{theorem}

\begin{proof}Consider $C_k(X^{\omega_{i}})$ to be the free abelian group with basis $\{\sigma: \Delta^k \rightarrow X^{\omega_{i}}\}$ of all singular simplices in $X^{\omega_{i}}$. A generator $\sigma \in C_{k}(X^{\omega_{i}})$ is a simplex in $X$ composed with the inclusion $\imath_{\omega_{i}}$, i.e. $(\imath_{\omega_{i}})_\star: C_k(X^{\omega_{i}}) \rightarrow C_k(X)$, $\sigma \mapsto \imath_{\omega_{i}} \circ \sigma$. If we take all $i \in I$, we can exhaust $C_k(X)$, which results in the isomorphism $\psi: \bigoplus_{i \in I} H_k(X^{\omega_{i}}) \rightarrow H_k(X)$ at chain level. Taking an element $x \in \bigoplus_{i \in I} C_k(X^{\omega_{i}})$ we get a finite sum of the functions $\sigma_1, \cdots, \sigma_n$ with their coefficients $z_1, \cdots, z_n \in \mathbb{Z}$, i.e. $x = \sum_{i=1}^{n} z_i\sigma_i$ and apply the group homomomorphism $\psi$, resulting in $\psi(x) = \sum_{i=1}^{n} z_i \imath_{\omega_i}(\sigma_i)$. Given a generator $\sigma: \Delta^k \rightarrow X$, we know that $\Delta^k$ is connected and $\sigma$ is a continuous function. Thus $\sigma(\Delta^k)$ is contained in a connected component of $X$, namely $\sigma \in C_k(X^{\omega_{j}})$ and $\sigma = \psi(\sigma)$, thus $\psi$ is surjective.\qed
\end{proof}

\subsection{Equivalence of Simplicial and Singular Homology}
\label{simplicialsingularequivalence}
\begin{theorem}{\cite[p.~128]{hatcher2002algebraic}}
Let $k \in \mathbb{Z}$ and $H^\Delta_k(K(X))$ the simplicial $k$-th homology group over a triangulable topological space $X$ and $K$ the triangulation. Let $H_k(X)$ be the respective singular homology group. Then the map $H^{\Delta}_k(K(X)) \rightarrow H_k(X)$ is an isomorphism of groups.
\end{theorem}

\begin{proof}
We follow the proofs of \cite[p.~128, pp.~190-192]{hatcher2002algebraic,munkres2018elements} and start from the perspective of the simplicial homology groups by choosing $K(X)$ as a simplicial complex and $K(X^n)$ as the $n$-skeleton of $X$, which consists of all simplices of dimension $n$ or less. Thus, with $K(X^{n-1}) \subset K(X)$ we obtain the following commutative diagram of exact sequences, for which $\ker(\psi_i) = \Ima(\psi_{i+1})$:
\[
\begin{tikzcd}
& {H^{\Delta}_{k+1}((K(X^n),K(X^{n-1})))} \arrow[d, "\sigma_i"] \arrow[r, "\psi_i"]
& H^{\Delta}_{k}(K(X^{n-1})) \arrow[d, "\sigma_j"] \arrow[dd, bend left=80, "\psi_j"] \\
& {H_{k+1}((X^n,X^{n-1}))} \arrow[r, "\psi'_i"]
& H_{k}(X^{n-1}) \arrow[dd, bend left=80, "\psi'_j"]\\
H^{\Delta}_{n-1}(K(X^{n-1})) \arrow[d, "\sigma_m"]
& {H^{\Delta}_{k}((K(X^n),K(X^{n-1})))} \arrow[l,"\psi_k"] \arrow[d, "\sigma_l"]
& H^{\Delta}_{k}(K(X^n)) \arrow[l, "\psi_l"] \arrow[d, "\sigma_k"]\\
H_{k-1}(X^{n-1})
& {H_{k}((X^n,X^{n-1}))} \arrow[l,"\psi'_k"]
& H_{k}(X^{n}) \arrow[l,"\psi'_l"]
\end{tikzcd}
\]
The quotient space $X^n/X^{n-1}$ contains only simplices of dimension $n$. For $k \neq n$ the group $C_k((X^n,X^{n+1}))$ is zero. For $k = n$, the group $C_k((X^n,X^{n+1}))$ is free and commutative with a basis consisting of $k$-simplices. The cycles $\ker \partial_k$ are a subgroup of $C_k((X^n,X^{n+1}))$ and the boundary group $\Ima \partial_{k+1}$ is empty. We conclude that $H^{\Delta}_k((K(X^n),K(X^{n-1})))$ has a description like $C_k((X^n,X^{n-1}))$.

The continuous maps $\sigma: \Delta^n \rightarrow X$ into the topological space $X$ give the map $\psi: \coprod_i(\Delta^n_i,\Delta^{n-1}_i) \rightarrow (X^n,X^{n-1})$ and for the simplicial quotient spaces $\psi_{\star}: \coprod_i \Delta^n_i / \coprod_i \Delta^{n-1}_i \rightarrow X^n/X^{n-1}$. In view of these homeomorphisms, the following isomorphism of homology groups results:
\begin{align}
	H_k(\coprod_i \Delta^n_i / \coprod_i \Delta^{n-1}_i) \simeq H_k(X^n / X^{n-1}).
\end{align}

A consequence of the Excision Theorem is the existence of an isomorphism $H_k((X,V)) \rightarrow H_k(X/V)$ for all good pairs $(X,V)$ \cite[pp.~175-182]{munkres2018elements}.

\begin{theorem}
Given the topological spaces $U \subset V \subset X$ for which the closure $\overline{U} \subset \interior{V}$ is in the interior of $V$, the embedding $(X-U,V-U) \hookrightarrow (X,V)$ induces an isomorphism $H_k((X-U, V-U)) \rightarrow H_k((X,V))$ for all $k \in \mathbb{Z}$.
\end{theorem}

\begin{proof}
For a proof we refer to \cite[pp.~175-182]{munkres2018elements}. \qed
\end{proof}

\begin{remarktheo}
Recall that the notation $(X^{n},X^{n-1})$ denotes a topological pair, i.e. a shorthand to indicate an inclusion $\imath: X^{n-1} \hookrightarrow X^n$. 
\end{remarktheo}
\begin{remarktheo}
As a reminder, a good pair $(X,V)$ is given if there is an open neighborhood $W \subset X$ with $V \subset W$ such that this deformation retracts onto $V$. 
\end{remarktheo}

Applying transitivity of group operations yields
\begin{align}
	H_k\left(\coprod_i\left(\Delta^n_i,\Delta^{n-1}_i\right)\right) &\simeq H_k\left(\coprod_i \Delta^n_i/\coprod_i \Delta^{n-1}_i\right),\\
	H_k\left(\coprod_i \Delta^n_i/\coprod_i \Delta^{n-1}_i\right) &\simeq H_k\left(\left(X^n,X^{n-1}\right)\right),\\
	H_k\left(\left(X^n,X^{n-1}\right)\right) &\simeq H_k(X^n / X^{n-1}).
\end{align}
We see, that $H_k((X^n,X^{n-1}))$ is zero, for $k \neq n$ and a free abelian group, on the basis of the cycles given by the continuous map $\sigma: \Delta^n \rightarrow X$. It follows that the map $H^\Delta_k((K(X^n),K(X^{n-1}))) \rightarrow H_k((X^n,X^{n-1}))$ is an isomorphism.

In the commutative diagram of simplicial and singular homology groups above, the maps $\sigma_i$ and $\sigma_m$ are isomorphisms. By induction $\sigma_j$ and $\sigma_l$ are also isomorphisms. It remains to be shown that $\sigma_k$ is also an isomorphism. For this purpose we use the Five Lemma \cite{pryde1977five}, from which follows that $\sigma_k$ is an isomorphism and therefore the map $H^{\Delta}_k(K(X^n)) \rightarrow H_k(X^n)$ is an isomorphism. \qed
\end{proof}

\subsection{Singular Homology and Smooth Manifolds}
To complete the argument we use ideas from a proof of de Rham`s Theorem. De Rham's Theorem manifests the fundamental connection of the analytic and topological properties of a smooth manifold, where the pairing of differential forms and singular chains gives an isomorphism from $k$-th de Rham cohomology to $k$-th singular homology and an isomorphism into simplicial homology.

\begin{theorem} The de Rham cohomology of a smooth, oriented and trianguliable manifold $H_{k}^{\text{dR}^{*}}(M)$ has an isomorphism to the real homology group $H_k(K(M);\mathbb{R})$, e.g. calculated by a singular \v{C}ech complex $K$ on some smooth manifold $M$ with real coefficients.
\end{theorem}

\begin{proof}
For a complete proof we refer to  \cite[Chap.~18, Sect.~V.9]{lee2013smooth,bredon2013topology}.
\end{proof}

\begin{theorem}
\label{singularsmoothhomgroup}
The $k$-th smooth singular homology group is defined as
\begin{equation}
	H_{k}^{\infty} (M) = \frac{\ker\left(\partial: C_{k}^{\infty} \rightarrow C_{k-1}^{\infty}\left(M\right)\right)}{\Ima\left(\partial: C_{k+1}^{\infty}\left(M\right) \rightarrow C_{k}^{\infty}\left(M\right)\right)},
\end{equation}
and gives an isomorphism into the singular homology groups on $M$, such that the map $\imath_{\star}: H_{k}^{\infty}(M) \rightarrow H_k(M)$ is an isomorphism of groups.
\end{theorem}

\begin{remarktheo}
For the free abelian group of $k$-chains $C_k(M)$ created by smooth $k$-simplices on $M$ we introduce the notation $C_k^{\infty}(M)$. These are formal sums of finitely many smooth $k$-simplices. The boundary of a smooth $(k+1)$-simplex is thus a smooth $k$-chain.
\end{remarktheo}

\begin{proof}
Let $U \subset \mathbb{R}^n$ be a subset of the Euclidean space and $f: U \rightarrow M$ a smooth map into the manifold $M$. The map is said to be smooth in $p \in U$ if there is a smooth extension of $f$ to an open neighborhood of $p$. If $p$ is arbitrary, then $f$ is smooth in $M$. The standard basis vectors $\Delta_i$ form a smooth $k$-simplex $\Delta^k := [v_0, \ldots, v_k] \subset \mathbb{R}^n$, with a smooth map $\sigma: \Delta^k  \rightarrow M$.

The inclusion $\imath: C_{k}^{\infty}(M) \hookrightarrow C_k(M)$ satisfies $\imath \circ \partial = \partial \circ \imath$, from which we can see, that it induces a map $\imath_{\star}: H_{k}^{\infty}(M) \rightarrow H_k(M)$, for which $\imath_{\star}[x] = [\imath(x)]$.

\begin{lemma}
For the simplex $\Delta^k \subset \mathbb{R}^n$ with the standard basis vectors as vertices, let $\sigma: \partial \Delta^k \rightarrow M$ be the continuous map with a restriction to the boundary of each face. Then $\sigma$ is smooth as a function from $\partial \Delta^k$ into $M$.
\end{lemma}

\begin{proof}
We follow the proof of \cite[Lem.~2.1]{parkproof}. Let $0 \leq i \leq k$ and $\hat{v}_i$ denote the boundary face opposite of $\partial_{i} \Delta^k = [v_0, \cdots, \hat{v}_i, \cdots, v_k]$, where the primer represents the omitted element by the boundary operator. For each $i$ and each $v \in\partial_{i} \Delta^k$ we seek an open neighborhood $v \in U_v \subseteq \mathbb{R}^n$ and a smooth extension $\tilde{\sigma}: U_v \rightarrow M$ of $\sigma$ such that $\tilde{\sigma}$ restricted to $U_v \cap \partial_{i}\Delta^k$ corresponds exactly to $\sigma$.

\begin{proposition}
For each $v \in \partial \Delta^k$ a smooth extension $\tilde{\sigma}$ into $M$ exists.
\label{smoothextension}
\end{proposition}

\begin{proof}
We follow the proof of \cite[Lem.~2.1]{parkproof}. The element $v$ is contained in at least one boundary face of $\Delta^k$, but cannot be contained in all of them. This means that the boundary faces $\partial_1 \Delta^k, \cdots, \partial_j \Delta^k$ for $1 \leq j \leq k$ contain $v$. In particular $v$ is not contained in $\partial_0 \Delta^k$. Let $U = \bigcap_{i=1}^{j} U_i$ be the intersections for each possible $v \in \partial_i\Delta^k$. Using the induction hypothesis, we show that there is a smooth map $\tilde{\sigma}: U \rightarrow M$, whose restriction to $U \cap \left (\bigcup_{i=1}^j \partial_i \Delta^k \right)$ matches $\sigma$. This property is shown locally, such that for small enough $U$, $\sigma$ maps $U$ to a coordinate chart of $f(v)$, which in turn can be identified diffeomorphically as a subspace of $\mathbb{R}^n$. For $j=1$ the proof is trivial, thus assume that $j > 1$ and that a smooth map $\tilde{\sigma}_0 : U \rightarrow M$ exists, whose restriction to $U \cap \left( \bigcup_{i=1}^j \partial_i \Delta^k \right)$ matches $\sigma$. We define $\tilde{\sigma}: U \rightarrow M$ with
\begin{align}
	&\tilde{\sigma} \circ \left(v_1,\cdots,v_{i-1},0,v_{i+1},\cdots,v_n\right) = \\
	= \; &\tilde{\sigma}_0 \circ \left( v_1, \cdots, v_m\right) - \tilde{\sigma}_0 \circ  \left(v_1,\cdots,v_{j-1},0,v_{j+1},\cdots,v_m \right) + \\
	& + \tilde{\sigma}_j\circ \left(v_1,\cdots,v_{j-1},0,v_{j+1},\cdots,v_m \right).
\end{align}
For each $1 \leq i \leq j -1$ the restriction $\tilde{\sigma}$ to $U \cap \partial_i \Delta^k$ is given by
\begin{align}
&\tilde{\sigma} \circ \left(v_1,\cdots,v_{i-1},0,v_{i+1},\cdots, v_n\right) = \\
= \; &\tilde{\sigma}_0 \circ \left(v_1, \cdots, v_{i-1}, 0, v_{i+1}, \cdots, v_n\right) - \\
&- \tilde{\sigma}_0 \circ \left(v_1, \cdots, v_{i-1}, 0, v_{i+1}, \cdots,v_{j-1},0,v_{j+1},\cdots, v_n\right) + \\
&+ \tilde{\sigma}_j \circ \left(v_1, \cdots, v_{i-1}, 0, v_{i+1}, \cdots,v_{j-1},0,v_{j+1}, \cdots, v_n\right) = \\
= \; &\sigma \circ \left(v_1, \cdots, v_{i-1}, 0, v_{i+1}, \cdots, v_n\right).
\end{align}
The singular simplex $\tilde{\sigma}_0$ is equal to $\sigma_i$ for a $v_i \in \Delta^k$ with $v_i = 0$. Similarly, $\tilde{\sigma}_j$ is equal to $\sigma_j$ for a $v_j \in \Delta^k$ with $v_j = 0$.

An analogous argument shows, that for some face of a singular simplex $\tilde{\sigma}$, restricted to $U \cap \partial_j \Delta^k$, the inductive step can be completed as follows:
\begin{align}
&\tilde{\sigma} \circ \left(v_1,\cdots,v_{j-1},0,v_{j+1},\cdots, v_n\right) = \\
= \; &\tilde{\sigma}_0 \circ \left(v_1, \cdots, x_{j-1}, 0, v_{j+1}, \cdots, v_n\right) - \\
&- \tilde{\sigma}_0 \circ \left(v_1, \cdots, v_{j-1}, 0, v_{j+1}, \cdots, v_n\right) + \\
&+ \tilde{\sigma}_j \circ \left(v_1, \cdots, v_{j-1}, 0, v_{j+1}, \cdots, v_n\right) = \\
= \; &\sigma \circ \left(v_1, \cdots, v_{i-1}, 0, v_{i+1}, \cdots, v_n\right).
\end{align}
\qed
\end{proof}
Prop. \ref{smoothextension} proofs the lemma, as every $v_i \in \partial \Delta^k$ has a smooth extension for arbitrary $k$, every element of any simplex has a smooth extension.
\qed
\end{proof}

For each simplex we construct a homotopy to a smooth simplex, such that the homotopies respect the restriction to the boundary faces.

\begin{lemma}
\label{smoothhomtop}
For every $k$-simplex $\sigma: \Delta^k \rightarrow M$ from the standard $k$-simplex of $\mathbb{R}^n$ to $M$, there exists a continous map $h_\sigma: \Delta^k \times I \rightarrow M$, beeing a homotopy, satisfying the following conditions:
\begin{itemize}
	\item[(1)] $h_\sigma$ maps from $\sigma(v) = h_\sigma(v,0)$ to a smooth $k$-simplex $\tilde{\sigma}(v) = h_\sigma(v,1)$.
	\item[(2)] For each boundary face inclusion $s_{i,k}: \Delta^{k-1} \simeq \partial_{i} \Delta^k \hookrightarrow \Delta^{k}$, we get \\$h_{\sigma \circ s_{i,k}}(v,t)$ $=$ $h_\sigma(s_{i,k}(v),t)$ for $(v,t) \in \Delta^{k-1} \times I$.
	\item[(3)] For a smooth $k$-simplex $\sigma$, $h_\sigma$ is the constant homotopy, i.e. $h_{\sigma}(v,t) = \sigma(v)$.
\end{itemize}
\end{lemma}

\begin{proof}
For a proof we refer to \cite[Lem.~2.3]{parkproof}. \qed
\end{proof}
To finalize our proof of Theo. \ref{singularsmoothhomgroup}, we follow \cite[Theo.~2.4]{parkproof}. Let $\imath_0: \Delta^k \hookrightarrow \Delta^k \times I$, $v \mapsto (v,0)$ and $\imath_1: \Delta^k \hookrightarrow \Delta^k \times I$, $v \mapsto (v,1)$. The smooth maps allow to define a homomorphism $f: C_k(M) \rightarrow C^{\infty}_{k}(M)$ by mapping $\sigma \mapsto h_\sigma \circ \imath_1$ and extending linearly. Then $f$ is a map of chains, due to
\begin{align}
f \circ \partial(\sigma) &= f\left( \sum_{i=0}^{k} (-1)^{i} h_{\sigma \circ s_{i,k}} \circ \imath_1 \right) =\\
&= \sum_{i=0}^{k} (-1)^i h_\sigma \circ (s_{i,k} \times \text{Id}_{I}) \circ \imath_1 =\\
&= \sum_{i=0}^{k} (-1)^i h_\sigma \circ \imath_1 \circ s_{i,k} =\\
&= \partial(h_\sigma \circ \imath_1).
\end{align}
Thus $f$ induces a homomorphism $f_{\star}:H_k(M) \rightarrow H^{\infty}_{k}(M)$.

\begin{proposition}
The homomorphism $f_{\star}:H_k(M) \rightarrow H^{\infty}_{k}(M)$ is the inverse of $\imath_{\star}: H^{\infty}_{k}(M) \rightarrow H_k(M)$, such that $\imath_{\star} = f_{\star}^{-1}$.
\end{proposition}
\begin{proof}
By Lem. \ref{smoothhomtop} (3) $f \circ \imath = \text{Id}_{C_{k}^{\infty}(M)}$, such that $f_{\star} \circ \imath_{\star} = \text{Id}_{H_{k}^{\infty}(M)}$. Vice versa, it suffices to show that a chain homotopy $h: C_k(M) \rightarrow C_{k+1}(M)$ exists, such that $\partial \circ h + h \circ \partial = \imath \circ f - \text{Id}_{C_{k}(M)}$. Let $V \subset \mathbb{R}^n$ and $A: \Delta^{k} \rightarrow V$, an affine transformation which sends $v_i \mapsto p_i$ denoted as $A(v_0, \cdots, v_k)$. We define a map $t_{i,k}: \Delta^{k} \rightarrow \Delta^{k} \times I$ such that
\begin{align}
	t_{i,k} := A\left( \left( v_0,0 \right),\cdots,\left( v_{i-1},0 \right),\left( v_i,0 \right),\cdots,\left( v_k,0 \right) \right).
\end{align}
\qed
\end{proof}

We use the proof from \cite[Theo.~2.4]{parkproof} for the final step. The map $t_{i,k}: \Delta^{k+1} \rightarrow \Delta^k \times I$ is smooth and well defined. The compositions $t_{i,k} \circ s_{i,k+1}$ and $t_{i-1,k} \circ s_{i,k+1}$ coincide and map from $\Delta^k$ into $\Delta^k \times I$. This is stated as
\begin{align}
\label{interchange}
	t_{i,k} \circ s_{i,k+1} &= t_{i-1,k} \circ s_{i,k+1} = \\
	&= A\left(\left( v_0,0 \right), \cdots, (v_{i-1},0), (v_i,1),\cdots,(v_k,1)\right).
\end{align}
Analogous to this, the map $(s_{i,k} \times \text{Id}_{I}) \circ t_{j,k-1}: \Delta^{k} \rightarrow \Delta^{k} \times I$ is given by
\begin{align}
(s_{i,k} \times \text{Id}_{I}) \circ t_{j,k-1} =
	\begin{dcases}
	    t_{j+1,k} \circ s_{i,k+1} & \text{if} \; i \leq j, \\
	    t_{j,k} \circ s_{i+1,k+1} & \text{if} \; i > j.
	\end{dcases}
\end{align}
We define the inverse boundary operator $\partial^{-1}: C_{k}(M) \rightarrow C_{k+1}(M)$ into the $(k+1)$-st singular chain group:
\begin{align}
	\partial^{-1}(\sigma) := \sum_{i=0}^{k} (-1)^{i} h_\sigma \circ t_{j,k}.
\end{align}
According to \cite[Lem.~2.3]{parkproof}, applying the inverse boundary operator to the ordinary boundary operator yields
\begin{align}
	\partial^{-1} \circ \partial(\sigma) = & \partial^{-1}\left( \sum_{i=0}^{k} (-1)^i \sigma \circ s_{i,k} \right) =\\
	= &\sum_{i=0}^{k} \sum_{j=0}^{k-1} (-1)^{i+j} h_\sigma \circ s_{i,k} \circ t_{j,k-1} = \\
	= &\sum_{i=0}^{k}\sum_{j=0}^{k-1} (-1)^{i+j} h_\sigma \circ \left( s_{i,k} \times \text{Id}_I \right) \circ t_{j,k-1}=\\
	\label{smoothsingular1}
	= &\left( \sum_{0 \leq i \leq j \leq k-1} (-1)^{i+j} h_\sigma \circ t_{j+1,k} \circ s_{i,k+1} \right) + \\
	\label{smoothsingular2}
	  &+ \left( \sum_{0 \leq j \leq i \leq k-1} (-1)^{i+j} h_\sigma \circ t_{j,k} \circ s_{i+1,k+1} \right).
\end{align}
We interchange both operators and split the equations into subsums with $i < j-1$ for $j \neq 0$ and $i = j-1$, $j \neq k$ and $i = j$ and finally $i > j$. Then the negative of the following first and last sums in Eq. \ref{singularsmooth1} and \ref{singularsmooth2} equal the negative of the expressions in the last two sums of Eq. \ref{smoothsingular1} and Eq. \ref{smoothsingular2} \cite[Theo.~2.4]{parkproof}:
\begin{align}
	\partial \circ \partial^{1}(\sigma) = &\partial \left( \sum_{i=0}^{k} (-1)^i h_\sigma \circ t_{i,k} \right) = \\
	= & \sum_{i=0}^{k}\sum_{j=0}^{k+1} (-1)^{i+j} h_\sigma \circ t_{i,k} \circ s_{j,k+1} = \\
	= & \left( \sum_{0 \leq i < j-1} \sum_{j=0}^{k+1} h_\sigma \circ t_{i,k} \circ s_{j,k+1}\right) -\\
	\label{second}
	&- \left( \sum_{j=1}^{k+1} h_\sigma \circ t_{j-1,k} \circ s_{j,k+1} \right) + \\
	\label{singularsmooth1}
	&+ \left( \sum_{j=1}^{k+1} h_\sigma \circ t_{j,k} \circ s_{j,k+1} \right) + \\
	\label{singularsmooth2}
	&+ \left( \sum_{0 \leq j < i \leq k} (-1)^{i+j} h_\sigma \circ t_{i,k} \circ s_{j,k+1} \right).
\end{align}
Substituting the maps, by applying Eq. \ref{interchange}, cancels out Eq. \ref{second} and Eq. \ref{singularsmooth2}. The composition of the boundary operator and its inverse counterpart simplifies to
\begin{align}
	\partial \circ \partial^{-1}(\sigma) &= h_\sigma \circ t_{0,k} \circ s_{0,k+1} - h_\sigma \circ t_{k,k} \circ s_{k+1,k+1} = \\
	&= h_\sigma \circ A((v_0,0), \cdots, (v_k,0)) - h_\sigma \circ A((v_0,1), \cdots, (v_k,1)) = \\
	&= h_\sigma \circ \imath_1 - h_\sigma \circ \imath_0.
\end{align}
This shows, that the following composition cancels out, as needed:
\begin{align}
	\partial^{-1} \partial(\sigma) + \partial \circ \partial^{-1}(\sigma) &= h_\sigma \circ \imath_1 - h_\sigma \circ \imath_0 = \\
	&= \imath \circ f(\sigma) - \sigma.
\end{align}
\qed
\end{proof}

\section{Supplementary Material for Sect. \ref{sec:numericalexperiments}}
\label{appx:numexp}
We implement the persistence diagrams, persistence statistics and persistence landscapes using the GUDHI library \cite{gudhi:urm}. For the neural nets we use the Tensorflow backend \cite{tensorflow2015-whitepaper}, with the Keras environment as wrapper \cite{chollet2015keras}. We use Python in the version $3.7$ as programming language. The computations of the neural nets are conducted on Intel Core i7 9700K processors and the homology related parts are computed on a NVIDIA Quadro P4000 graphic card.

\subsubsection{Invertible $\mathcal{C}^1$-Net}
\label{appx:invarch}
The neural net used in Sect. \ref{sec:numericalexperiments} is a partially invertible architecture. It consists of a total of five hidden layers. The first layer is a dense layer, which connects the input $\mathbf{x}^{(0)} \mapsto \text{BN}\left(\sigma\left(\sum_{i=1}^{n} \textbf{W}^{(1)}_{ji}\mathbf{x}^{(0)}_{i} + \mathbf{b}^{(1)}_{i}\right)\right)$ to a subspace of $\mathbf{x}^{(0)} \in \mathbb{R}^n$. The subspace is initially $\mathbb{R}^{n-k}$ with $(n-k) \leq n$, according to Fig. \ref{architecture}. The sigmoid function is used as activation function. The weights are initialized uniformly distributed and the bias term is with zeros, respectively. The batch normalization can be expressed for a batch $k$ at the $l$-th layer as
\begin{align}
	\text{BN}(\mathbf{x}^{(l)}_{(k)})_{\gamma^{(l)}_{(k)},\beta^{(l)}_{(k)}} = \frac{\gamma}{\sqrt{\text{Var}[\mathbf{x}^{(l)}_{(k)}] + \epsilon}} \mathbf{x}^{(l)}_{(k)} + \left( \beta - \frac{\gamma \text{E}[\mathbf{x}^{(l)}_{(k)}]}{\sqrt{\text{Var}[\mathbf{x}^{(l)}_{(k)}]+\epsilon}} \right),
\end{align}
where $\gamma,\beta$ and $\epsilon$ are free parameters updated during back propagation. The very first and last dense layer is not invertible. Its purpose is a learned projection into the desired subspace of dimension $(n-k)$ for some $k$ and afterwards projecting into the original space in an encoding fashion, see Fig. \ref{architecture}.

\subsubsection{Non-invertible $\xi_i$- and $\psi_i$-Architecture}
The invertible architecture, which is described in Sect. \ref{subsec:invertibleneuralnets}, uses the partially linear functions $\xi_1,\xi_2$ and $\psi_1,\psi_2$, which are parameterized during the learning process. The architecture is taken from the work of Ardizzone et al. \cite{ArdizzoneKRK19}. These functions are to be learned and are composed in a trivially invertible way, see Sect. \ref{subsec:invertibleneuralnets}. The functions themselves are implemented as neural subnets. Each function is a $\mathcal{C}^1$-net as described in Sect. \ref{resnet}, with a total of $3$ dense layers. The vector $\mathbf{u}_{1}^{(l)} \in \text{M}(n,\mathbb{R})$ is transformed by $\xi_1,\xi_2,\psi_1,\psi_2: \text{M}(n,\mathbb{R}) \rightarrow \text{M}(n,\mathbb{R})$, such that the dimension of the vector does not change. The leaky rectifying linear unit activation function has been selected for every layer in the subnet:
\begin{align}
	\text{LeakyReLu}(\mathbf{x}^{(l)}) = 
	\begin{dcases}
	    \mathbf{x}^{(l)}_{i},
	    & \text{if} \; \mathbf{x}^{(l)}_{i} \geq 0, \\
	    5.5 \cdot \mathbf{x}^{(l)}_{i}
	    & \text{if} \; \mathbf{x}^{(l)}_{i} < 0.
	\end{dcases}
\end{align}
We apply batch normalization after the activation function, see Appx. \ref{appx:invarch}. The partition of the vector is still an open research question \cite{ArdizzoneKRK19}. We divide the flattened vector precisly into two halfs, since the samples are square matrices:
\begin{align}
\mathbf{x}^{(l)} = (\mathbf{u}^{(l)}_1, 0, \cdots, 0) + (0, \cdots, 0, \mathbf{u}^{(l)}_2).
\end{align}

\subsubsection{Hyperparameters for Learning} We use a batch size of $128$ for every model. We train the neural nets for $250$ epochs with the mean squared error as loss function, see Sect. \ref{sec:numericalexperiments}. As optimization algorithm we use ADAM \cite{KingmaB14}, with a learning rate of $10^{-4}$. The dataset is randomly shuffled. The high batch size should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Sect. \ref{sec:numericalexperiments}.

\bibliographystyle{splncs04}
\renewcommand{\refname}{References for Appendices}
\bibliography{biblio}
\end{document}

### 2020-04-02 expAutoencoder.py

@@ -1,0 +1,52 @@
+#!/usr/bin/env python
+
+from autoencoderInvertible import *
+from keras.models import Model
+from keras.losses import kullback_leibler_divergence, mean_squared_error
+from keras.optimizers import Adam
+from keras.datasets import mnist, cifar10, cifar100, boston_housing
+from keras.callbacks import TensorBoard, CSVLogger
+from multiprocessing import Pool
+
+from os import listdir
+from os.path import isfile, join
+
+
+def list_diff(list1, list2): 
+    return (list(set(list1) - set(list2))) 
+
+def execute_experiments(i):
+    print("Current dimension: " + str(i))
+    invertible_autoencoder = Model(
+    input_img, dense_invertible_subspace_autoencoder(input_img, units=i, invertibleLayers=5, groupLayers=3)
+    )
+    invertible_autoencoder.summary()
+    invertible_autoencoder.compile(optimizer=Adam(lr=0.0001), loss=mean_squared_error)
+    csv_logger = CSVLogger('results/cifar100/c_1_log_units'+ str(i) + '.csv', append=True, separator=',')
+    invertible_autoencoder.fit(
+        add_gaussian_noise(x_train, 0.8),
+        x_train,
+        epochs=250,
+        batch_size=128,
+        shuffle=True,
+        validation_data=(add_gaussian_noise(x_test, 0.8), x_test),
+        callbacks = [csv_logger],
+        verbose=0
+    )    
+
+def homgroups_dimension():
+    cifar10 = [12,16,40,59,50]
+    cifar100 = [13,18,34,46,48]
+
+    epsilon = 30.
+
+    for homologygroup in range(1,5):
+        homcount = cifar100[homologygroup]
+        for p in range(3, 1000):
+            number = 1.
+            for i in range(1, homologygroup + 1):
+                number = number * ((p - homologygroup + i) / i)
+            if number > (homcount - epsilon) and number < (homcount + epsilon):
+                print("The amount of needed spheres for H" + str(homologygroup) + " is approximately " + str(p) + " for epsilon = " + str(number - homcount))
+
+homgroups_dimension()

### 2020-04-02 persistenceLandscapes.py

@@ -1,0 +1,233 @@
+#!/usr/bin/env python
+
+import matplotlib.pyplot as plt
+import numpy as np
+import gudhi as gd
+import math
+import os
+import gudhi.representations
+import tikzplotlib
+import itertools
+import persistenceStatistics as ps
+import tensorflow as tf
+
+from sklearn.kernel_approximation import RBFSampler
+from sklearn.preprocessing import MinMaxScaler
+from keras.datasets import cifar10, cifar100, fashion_mnist
+from tqdm import tqdm
+from scipy.ndimage.filters import gaussian_filter1d
+
+colorScheme = {
+    "black": "#1A1A1D",
+    "gray": "#4E4E50",
+    "purpur": "#6F2232",
+    "brick": "#950740",
+    "fire": "#C3073F",
+}
+
+
+def concatenate_landscapes(
+    persLandscape1: np.ndarray, persLandscape2: np.ndarray, resolution: int
+) -> list:
+    """
+        **This function concatenates the persistence landscapes according to homology groups.**
+
+        The computation of homology groups requires a certain resolution for each homology class.
+        According to this resolution the direct sum of persistence landscapes has to be concatenated
+        in a correct manner, such that the persistent homology can be plotted according to the `n`-dimensional
+        persistent homology groups.
+
+        + param **persLandscape1**: persistence landscape, type `np.ndarray`.
+        + param **persLandscape2**: persistence landscape, type `np.ndarray`.
+        + return **concatenatedLandscape**: direct sum of persistence landscapes, type `list`.
+    """
+    numberPartition = int(len(persLandscape2[0]) / resolution)
+    splitLandscape1 = np.split(persLandscape1[0], numberPartition)
+    splitLandscape2 = np.split(persLandscape2[0], numberPartition)
+    concatenatedLandscape = []
+
+    for i in range(0, numberPartition):
+        concatenatedLandscape.append(
+             np.fmax(splitLandscape1[i], splitLandscape2[i]))
+
+    return [np.array(concatenatedLandscape).flatten()]
+
+
+def compute_persistence_landscape(
+    data: np.ndarray,
+    res: int = 1000,
+    persistenceIntervals: int = 1,
+    maxAlphaSquare: float = 1e12,
+    filtration: str = ["alphaComplex", "vietorisRips", "tangential"],
+    maxDimensions: int = 10,
+    edgeLength: float = 0.1,
+    plot: bool = False,
+    smoothen: bool = False,
[…]

### 2020-04-02 llncs2e/paper77.tex
typo
typos

@@
 Persistent homology allows us to count its $k$-th homology groups.
-Using Künneth's theorem, we count the $n$-th Betti numbers.
+Using Künneth's theorem, we count the $k$-th Betti numbers.
 Since we know the embedding dimension of $\mathbb{R}^p$ and $\mathcal{S}^1$, we parameterize the bottleneck layer with the smallest possible matrix group, which can represent a manifold with those homology groups.
@@
 \ref{sec:manifoldassumption} we discuss the mathematical assumptions and recall the principles of smooth manifolds and Lie theory.
-In particular, we recall the decomposition of connected commutative Lie groups into $\mathbb{T}^p \times \mathbb{R}^q$.
+In particular, we recall the decomposition of connected commutative Lie groups into $\mathbb{R}^p \times \mathbb{T}^q$.
     \item In Sect.
@@
     \item In Sect.
-\ref{sec:numericalexperiments} we validate our theory training auto encoders on the cifar10 and cifar100 datasets, isolating the innermost layer such that no residual connections bypass it on different embedding dimensions.
+\ref{sec:numericalexperiments} we validate our theory training autoencoders on the cifar10 and cifar100 datasets, isolating the innermost layer such that no residual connections bypass it on different embedding dimensions.
 \end{itemize}
@@
 We specify the concept of smooth manifolds.
-Then we dive into Lie groups concluding with a central theorem on the isomorphism of abelian connected Lie groups to the product space $\mathbb{T}^p \times \mathbb{R}^q$, which turns out useful to estimate the embedding dimension of the manifold.
+Then we dive into Lie groups concluding with a central theorem on the isomorphism of abelian connected Lie groups to the product space $\mathbb{R}^p \times \mathbb{T}^q$, which turns out useful to estimate the embedding dimension of the manifold.
 
@@
 The neutral element of a group is denoted by $e = e_G$.
-Each abelian commutative Lie group $G$ of dimension $\dim G = m$ is isomorphic to a particular product space $\mathbb{T}^p \times \mathbb{R}^q$, with $p + q = m$, see Appx.
+Each abelian commutative Lie group $G$ of dimension $\dim G = m$ is isomorphic to a particular product space $\mathbb{R}^p \times \mathbb{T}^q$, with $p + q = m$, see Appx.
 \ref{connectedcommutativeliegroupdecomposition}.
@@
 A tool is needed to estimate homology groups of the manifold on which the data points lie.
-The method can then be inversely applied to the homology groups to estimate the embedding dimension of the data manifold.
+The method can then be inversely applied to the homology groups to estimate its embedding dimension.
 We introduce persistent homology with its confidence intervals in order to obtain the best possible estimate of the current state of the art by removing topological noise.
@@
 \end{equation}
-Removing a vertex from the $\Delta^n$ results in a $(n-1)$-simplex called \emph{face} of $\Delta^n$. The combinatorial description of $n$-simplices is called \emph{simplicial complex}.
+Removing a vertex from $\Delta^n$ results in a $(n-1)$-simplex called \emph{face} of $\Delta^n$. The combinatorial description of $n$-simplices is called \emph{simplicial complex}.
 

### 2020-04-02 llncs2e/paper77.tex
corrected assumptions

@@
 %
-\title{Parametrization of Invertible Resnets up to Homology Equivalence for Abelian Lie Groups\thanks{The work is partially supported by Siemens Power and Gas.
-The code is published on GitHub: \href{https://github.com/karhunenloeve/PHIA}{https://github.com/karhunenloeve/PHIA}.}}
+\title{Parametrization of Invertible Resnets up to Homology Equivalence for Abelian Lie Groups}
 %
@@
 Simplicial complexes have a natural order of simplices, which is illustrated in Fig.
-\ref{complexes}, which plays a central role in their geometric realization \citemain[p.~14]{boissonnat2018geometric}.
+\ref{complexes}.
+It plays a central role in their geometric realization \citemain[p.~14]{boissonnat2018geometric}.
 If we choose a total order for all corners which are located in any simplex of $K$, we get an order for each single $k$-simplex.
@@
 \end{equation}
-The $k$-th singular homology group is isomorphic to the $n$-th simplicial homology group for a trianguliable metric space, i.e.
+The $k$-th singular homology group is isomorphic to the $k$-th simplicial homology group for a trianguliable metric space, i.e.
 $H_k(X) \simeq H^{\Delta}_k(X)$ for all $k \in \mathbb{Z}$, see Appx.
@@
 Recording the \emph{Betti numbers} of the $k$-th homology group along the filtration, we obtain the \emph{$k$-dimensional persistence diagram}.
-The Betti numbers are defined by $\rank(H_k(X))$.
+The Betti numbers $\beta_k$ are defined by $\rank(H_k(X))$.
 We write $H^{\omega_i}_k = H_k(K^{\omega_i}(X))$ as $k$-th homology group on the simplicial complex $K$ with parametrization $\omega_i$.
@@
 	\label{countingequation}
-	\sharp_{k}^{H}(\bar{\mathfrak{L}}(\mathfrak{P}(K(X))) \approx \prod_{i=1}^{k} \frac{q-k+i}{i}
-\end{align}
-
+	\sharp_{k}^{H}(\bar{\mathfrak{L}}(\mathfrak{P}(K(X))) \approx \prod_{i=1}^{k} \frac{q-k+i}{i}.
+\end{align}
 Note, that we assume to work on a smooth manifold, whose de Rham cohomology induces an isomorphism into the simplicial homology with coefficients in $\mathbb{R}$, see Appx.
@@
 Thus, the dimension of the data manifold is the dimension of the smallest possible bottleneck layer of the neural net  $\dim M = \min_l \dim \textbf{x}^{(l)}$ \citemain{HauserR17}.
-Our contribution provides the theory of how this dimension can be determined for all neural nets that are a composition of invertible layers, i.e.
-$\textbf{x}^{(l)} \in \text{GL}(n,\mathbb{R}) := \{\varphi: \mathbb{R}^n \rightarrow \mathbb{R}^n \; \vert \; \det \varphi \neq 0 \}$.
+Our contribution provides the theory of how this dimension can be determined for all neural nets that are a composition of invertible layers acting on an abelian Lie group as domain of the data.
 The concatenation of such maps is called a deep neural net:
@@
 \end{align} 
-We need a completely invertible neural net, therefore all layers are invertible -- except the very first one, see Fig.
+We operate on the Lie group $\text{GL}(n,\mathbb{R})$, therefore all layers are invertible -- except the very first one, see Fig.
 \ref{architecture} and Sect.
@@
 \ref{architecture}.
-To test if our prediction for the embedding dimension, we parameterize the innermost layer using two perceptrons.
-The nets are trained and evaluated according to the caption of Fig.
+To test our prediction for the embedding dimension, we parameterize the innermost layer using a multiple of $2$ perceptrons.
+The nets are trained and evaluated according to Fig.
 \ref{costfunctions}.
@@
 Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to abelianization of the input domain.
-This forms a solution for a special case of learning problems, allowing to restrict the domain to an almost commutative group.
+This forms a solution for a special case of learning problems, allowing to restrict the domain to a commutative group.
 The initial problem could be simplified by decomposing the Lie group into a product space $\mathbb{R}^p \times \mathbb{T}^q$.
 The homology groups of the $q$-torus and the Euclidean space are well known.
-The application of fully invertible neural nets to resnets allows to meet the requirements for working with the group of invertible matrices.
+The application of fully invertible neural nets to resnets demonstrates our results on a different Lie group from Euclidean space.
 By applying Künneth's theorem the homology groups of the topological factor spaces could be connected with the ones of their product space.
@@
 We do not limit the processing of the neural net within the layers to an abelian group.
-This can cause inaccuracies in the estimation of the dimensional treshold.
-The explicit abelianization of the invertible resnets solves this problem.
+Instead, we assume that the net learns this structure from the input domain.
+This seems to hold, according to our experiments.
+Still, it can cause inaccuracies in the estimation of the dimensional treshold.
+The abelianization of the invertible layers would solve this problem.
+
+\paragraph{Acknowledgements} The work is partially supported by Siemens Power and Gas. The authors would like to thank Noah Becker and David Haller for proofreading.
+
+\paragraph{Code} The code can be found at: \href{https://github.com/karhunenloeve/PHIA}{https://github.com/karhunenloeve/PHIA}.
 %

### 2020-04-02 llncs2e/paper77.tex
avoid misunderstanding

@@
 We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.8 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$.
-These input vectors are trained to the original vector $\textbf{x}^{(0)}$.
+These input vectors are trained to the sorted vector $\textbf{x}^{(0)}$.
 We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = \frac{1}{n} \sum_{i=1}^{n} \left(\textbf{x}^{(0)}_i-\textbf{y}_i\right)^2$ as loss function.
@@
 Further, it shows the estimated minimal embedding dimension for a manifold with these homology groups for $\mathcal{C}^1$ to $\mathcal{C}^4$-resnets.
-The representatives for the $k$-th homology group $\sharp_k^{H}\left(\bar{\mathfrak{L}}\left(\mathfrak{P}(K(X))\right)\right)$ and the corresponding dataset are counted.
+The representatives for the $k$-th homology group $\sharp_k^{H}\left(\bar{\mathfrak{L}}\left(\mathfrak{P}(K(X))\right)\right)$ of the respective dataset are counted.
 Then, we seek for an embedding dimension to represent a manifold with the number of spheres $q$ with these homology groups.
@@
 The noise was added after sorting.
-We can confirm that the neural net is capable of eliminating the additional structure of noise, which is no longer part of the abelian Lie group.
+The neural net is capable of eliminating the additional structure of noise, which is no longer part of the abelian Lie group.
 \input{figure4.tex}

### 2020-04-02 countHomgroups.py
code

@@ -15,7 +15,7 @@ from os.path import isfile, join
 def list_diff(list1, list2): 
     return (list(set(list1) - set(list2))) 
 
-def execute_experiments(i):
+def execute_experiments(x_train: np.ndarray, i: int):
     print("Current dimension: " + str(i))
     invertible_autoencoder = Model(
     input_img, dense_invertible_subspace_autoencoder(input_img, units=i, invertibleLayers=5, groupLayers=3)
@@ -48,5 +48,3 @@ def homgroups_dimension():
                 number = number * ((p - homologygroup + i) / i)
             if number > (homcount - epsilon) and number < (homcount + epsilon):
                 print("The amount of needed spheres for H" + str(homologygroup) + " is approximately " + str(p) + " for epsilon = " + str(number - homcount))
-
-homgroups_dimension()

### 2020-04-02 llncs2e/paper77.tex
clarification

@@
 In other words, the critical values of a function represented by a layer of $(n-1)$-perceptrons will result in many critical values of a layer of $n$-perceptrons.
-This motivates an appropriately sized model.
 
@@
 
-The idea originates from Lie theory.
-We use invertible architectures, to guarantee our assumptions and apply ascending sorting on the samples to yield permutation invariance or abelianization of the input domain \citemain{ArdizzoneKRK19,NguyenAK19}.
+We use invertible architectures, to demonstrate our hypothesis on a different lie Group from Euclidean space \citemain{NguyenAK19}.
 Our interpretation of a neural net is a continuous transformation, transferring invariant structures of the data in the input space to its deepest layer.
@@
 To do so, their homotopy groups must be equal.
-The zeroth homotopy group gives information about connected components which can be collapsed to a point by continuous functions, the first homotopy group about `holes' and the second about `cavities' and so on.
+The fundamental group gives information about connected components which can be collapsed to a point by continuous functions, the first homotopy group about `holes' and the second about `cavities' and so forth.
 But homotopies are difficult to compute \citemain{klaus2004quick,hatcher2002algebraic}.
-Homology theory gives us a remedy that guarantees an isomorphism between homotopy groups and homology groups under the conditions of Hurewicz's Theorem \citemain{klaus2004quick}.
-We use theoretical results from Lie theory and persistent homology theory to predict how many dimensions of embedding are needed to represent the homology groups.
+Homology theory gives a remedy that guarantees an isomorphism between homotopy groups and homology groups under the conditions of Hurewicz's Theorem \citemain{klaus2004quick}.
+We use theoretical results from Lie theory and persistent homology theory to estimate how many dimensions of embedding are needed to represent the homology groups.
 Our paper is structured as follows:
@@
 
-\paragraph{Neural Nets} The data manifold is supposed to be a commutative Lie group.
-Thus, we artificially generate commutativity by sorting input and target vectors beforehand.
+\paragraph{Neural Nets} The data manifold is supposed to be some commutative Lie group.
 We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.8 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$.
-These input vectors are trained to the sorted vector $\textbf{x}^{(0)}$.
+These input vectors are trained to the original vector $\textbf{x}^{(0)}$.
 We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = \frac{1}{n} \sum_{i=1}^{n} \left(\textbf{x}^{(0)}_i-\textbf{y}_i\right)^2$ as loss function.
@@
 We specify the parameterization for an embedding as well as for neural nets of the smoothness degree $1$--$4$.
-The noise was added after sorting.
 The neural net is capable of eliminating the additional structure of noise, which is no longer part of the abelian Lie group.
@@
 \label{sec:conclusion}
-Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to abelianization of the input domain.
-This forms a solution for a special case of learning problems, allowing to restrict the domain to a commutative group.
+Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some commutative Lie group.
+This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group.
 The initial problem could be simplified by decomposing the Lie group into a product space $\mathbb{R}^p \times \mathbb{T}^q$.

### 2020-04-02 llncs2e/paper77.tex
notation

@@
 \label{connectedcommutativeliegroupdecomposition}
-Every connected commutative (or in other words abelian) Lie group $G$ is isomorphic to some product space $\mathbb{T}^p \times \mathbb{R}^q$ with $\dim G = p+q$, i.e.
-$G \simeq \mathbb{T}^p \times \mathbb{R}^q$.
+Every connected commutative (or in other words abelian) Lie group $G$ is isomorphic to some product space $\mathbb{R}^p \times \mathbb{T}^q$ with $\dim G = p+q$, i.e.
+$G \simeq \mathbb{R}^p \times \mathbb{T}^q$.
 \end{theorem}
@@
 &   G\\	
-&   (\mathbb{R}/\mathbb{Z})^p \times \mathbb{R}^q 
+&   (\mathbb{R})^p \times \mathbb{R/\mathbb{Z}}^q 
 	\arrow[swap]{ur}{\overline{\text{exp} \; \circ \; \varphi^{-1}}}
@@
 \]
-The canonical map $\pi: \mathbb{R}^n \simeq \mathbb{R}^p \times \mathbb{R}^q \rightarrow (\mathbb{R}/\mathbb{Z})^p \times \mathbb{R}^q$ is locally diffeomorphic and $\overline{\text{exp} \circ \varphi^{-1}} \circ \pi$ as well.
+The canonical map $\pi: \mathbb{R}^n \simeq \mathbb{R}^p \times \mathbb{R}^q \rightarrow \mathbb{R}^p \times (\mathbb{R}/\mathbb{Z})^q$ is locally diffeomorphic and $\overline{\text{exp} \circ \varphi^{-1}} \circ \pi$ as well.
 Due to the homogeneity of commutative lie groups, the map is everywhere a diffeomorphism.
 The projection $\overline{\text{exp} \; \circ \; \varphi^{-1}}$ is bijective, thus also a diffeomorphism and thus an isomorphism of Lie groups.
-By definition $\mathbb{R}^n/\mathbb{Z}^n \simeq (\mathbb{R}/\mathbb{Z})^n \simeq \mathbb{T}^n$ and the proof follows.
+By definition $\mathbb{R}^q/\mathbb{Z}^q \simeq (\mathbb{R}/\mathbb{Z})^q \simeq \mathbb{T}^q$ and the proof follows.
 \qed
@@
 The functions themselves are implemented as neural subnets.
-Each function is a $\mathcal{C}^1$-net as described in Sect.
-\ref{resnet}, with a total of $3$ dense layers.
+Each function is a $\mathcal{C}^1$-net with a total of $3$ dense layers, see Sect.
+\ref{resnet}.
 The vector $\mathbf{u}_{1}^{(l)} \in \text{M}(n,\mathbb{R})$ is transformed by $\xi_1,\xi_2,\psi_1,\psi_2: \text{M}(n,\mathbb{R}) \rightarrow \text{M}(n,\mathbb{R})$, such that the dimension of the vector does not change.

### 2020-04-02 llncs2e/paper77.tex
typos

@@
 We denote the homology groups of the components with $H_k^{\Delta}(K^{\omega_{i}}(X))$.
-With $C_k(K^{\omega_{i}}(X))$, we denote the group of $k$-chains for the connected component of the simplicial complex.
+With $C_k(K^{\omega_{i}}(X))$, we denote the group of $k$-chains for the connected components of the simplicial complex.
 Then $C_k(K^{\omega_{i}}(X)) \subseteq C_k(K(X))$ is a subgroup for all $i$.
@@
 An element in the $(k+1)$-th chain group $v \in C_{k+1}(K(X))$ is represented by $v = \sum_{i \in I} v_{\omega_{i}}$.
-Thus $\partial_{k+1}(v) \in C_{k}(K(X))$ can be broken down to $\partial_{k+1}(v) = \sum_{i \in I} \partial_{k+1}(v_{\omega_{i}})$ where $v_{\omega_{i}} \in C^\omega_{k+1}(K^{\omega_{i}})$.\qed
-\end{proof}
-
-The groups of coundaries can also be decomposed.
+Thus $\partial_{k+1}(v) \in C_{k}(K(X))$ can be broken down to $\partial_{k+1}(v) = \sum_{i \in I} \partial_{k+1}(v_{\omega_{i}})$ where $v_{\omega_{i}} \in C^\omega_{k+1}(K^{\omega_{i}}(X))$.\qed
+\end{proof}
+
+The groups of boundaries can also be decomposed.
 The kernel of $C_k(K^{\omega_{i}}(X))$ can be written as $\ker \partial_k(C_k(K^{\omega_{i}}(X)))$ $= \ker \partial_k(C_k(K(X)))$ $\cap$ $C_k(K^{\omega_{i}}(X))$.
@@
 \begin{remarkdef}
-Since for simplicial complexes, the property to be connected corresponds to path-connectedness, the coordinatewise sum of the connected components is a generator of the whole group.
+Since for simplicial complexes the property to be connected corresponds to path-connectedness, the coordinatewise sum of the connected components is a generator of the whole group.
 \end{remarkdef}

### 2020-04-02 llncs2e/paper77.tex
data

@@
 
-An analogous argument shows, that for some face of a singular simplex $\tilde{\sigma}$, restricted to $U \cap \partial_j \Delta^k$, the inductive step can be completed as follows:
+An analogous argument shows, that for any face of a singular simplex $\tilde{\sigma}$, restricted to $U \cap \partial_j \Delta^k$, the inductive step can be completed as follows:
 \begin{align}
@@
 \label{smoothhomtop}
-For every $k$-simplex $\sigma: \Delta^k \rightarrow M$ from the standard $k$-simplex of $\mathbb{R}^n$ to $M$, there exists a continous map $h_\sigma: \Delta^k \times I \rightarrow M$, beeing a homotopy, satisfying the following conditions:
+For every $k$-simplex $\sigma: \Delta^k \rightarrow M$ from the standard $k$-simplex of $\mathbb{R}^n$ to $M$, there exists a continuous map $h_\sigma: \Delta^k \times I \rightarrow M$, beeing a homotopy, satisfying the following conditions:
 \begin{itemize}
@@
 \end{align}
-where $\gamma,\beta$ and $\epsilon$ are free parameters updated during back propagation.
+where $\gamma,\beta$ and $\epsilon$ are parameters updated during back propagation.
 The very first and last dense layer is not invertible.

### 2020-04-03 llncs2e/paper77.tex
typo

@@
 
-We use invertible architectures, to demonstrate our hypothesis on a different lie Group from Euclidean space \citemain{NguyenAK19}.
+We use invertible architectures, to demonstrate our hypothesis on a different Lie group from Euclidean space \citemain{NguyenAK19}.
 Our interpretation of a neural net is a continuous transformation, transferring invariant structures of the data in the input space to its deepest layer.
@@
 
-\paragraph{Code} The code can be found at: \href{https://github.com/karhunenloeve/PHIA}{https://github.com/karhunenloeve/PHIA}.
+\paragraph{Code} The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.
 %

### 2020-04-04 llncs2e/paper77.tex
title
error

@@
 %
-\title{Parametrization of Invertible Resnets up to Homology Equivalence for Abelian Lie Groups}
+\title{Parametrization of Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups}
 %
@@
 
-\paragraph{Neural Nets} The data manifold is supposed to be some commutative Lie group.
+\paragraph{Neural Nets} The data manifold is supposed to be some connected commutative Lie group.
 We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.8 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$.

### 2020-04-04 llncs2e/paper77.tex
lekyRelu

@@
 We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = \frac{1}{n} \sum_{i=1}^{n} \left(\textbf{x}^{(0)}_i-\textbf{y}_i\right)^2$ as loss function.
-The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = \frac{1}{1 + e^{-\mathbf{x}^{(l)}}}$ is applied as elementwise activation function throughout the net.
+The leaky ReLu Eq.
+\ref{LeakyReLu} activation function $\sigma(\mathbf{x}^{(l)}) = \frac{1}{1 + e^{-\mathbf{x}^{(l)}}}$ is applied as elementwise activation function throughout the net.
 In all experiments dense neural nets are used with a bias term.
@@
 \ref{architecture}.
-The sigmoid function is used as activation function.
+The leaky ReLu function is used as activation function.
 The weights are initialized uniformly distributed and the bias term is with zeros, respectively.
@@
 \begin{align}
+\label{LeakyReLu}
 	\text{LeakyReLu}(\mathbf{x}^{(l)}) = 
@@
 	    & \text{if} \; \mathbf{x}^{(l)}_{i} \geq 0, \\
-	    5.5 \cdot \mathbf{x}^{(l)}_{i}
+	    -5.5 \cdot \mathbf{x}^{(l)}_{i}
 	    & \text{if} \; \mathbf{x}^{(l)}_{i} < 0.

### 2020-04-04 llncs2e/paper77.tex
minor errors

@@
 \hypersetup{
-    pdftitle={Approximate Parametrization of Invertible Neural Nets up to Homology Equivalence},    % title
+    pdftitle={Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups},    % title
     pdfauthor={Luciano Melodia},     % author
@@
 %
-\title{Parametrization of Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups}
+\title{Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups}
 %

### 2020-04-04 llncs2e/paper77.tex
error

@@
 \label{sec:conclusion}
-Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some commutative Lie group.
+Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group.
 This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group.

### 2020-04-04 llncs2e/paper77.tex
minor corrections

@@
 Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group.
-This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group.
+This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold.
 The initial problem could be simplified by decomposing the Lie group into a product space $\mathbb{R}^p \times \mathbb{T}^q$.
@@
 Still, it can cause inaccuracies in the estimation of the dimensional treshold.
-The abelianization of the invertible layers would solve this problem.
+The restriction of the layers to a maximal abelian subgroup of $\text{GL}(n,\mathbb{R})$ solves this problem.
 

### 2020-04-04 llncs2e/paper77.tex
minor errors
errors

@@
 We discuss a method to determine the smallest topology preserving layer for an invertible residual net.
-We consider the data domain as abelian connected matrix group and observe that it is decomposable into $\mathbb{R}^p \times \mathbb {T}^q$.
+We consider the data domain as abelian connected Lie group and observe that it is decomposable into $\mathbb{R}^p \times \mathbb {T}^q$.
 Persistent homology allows us to count its $k$-th homology groups.
@@
 By applying Künneth's theorem the homology groups of the topological factor spaces could be connected with the ones of their product space.
-Using mean confident persistence landscapes the homology groups were estimated.
-A treshold was computed, see sect.
-\ref{sec:numericalexperiments}, which is determined by the state space dimension of the resets according to a representation up to homology equivalence.
-The mean persistence landscapes are particularly useful for short-lived isolated collections of data points, which in particular generate elements from homology classes at the beginning of the filtering and become visible at the lower left corner of the persistence landscapes, see Fig.
+Using mean confidence persistence landscapes the homology groups were estimated.
+A treshold was computed, see Sect.
+\ref{sec:numericalexperiments}, which is determined by the state space dimension of the resnets according to a representation up to homology equivalence.
+The mean persistence landscapes are particularly useful for short-lived isolated collections of data points, which in particular generate elements from homology classes at the beginning of the filtration and become visible at the lower left corner of the persistence landscapes, see Fig.
 \ref{homgroups} (a).
 The theoretical work was tested with numerical experiments on toy data.
-We do not limit the processing of the neural net within the layers to an abelian group.
-Instead, we assume that the net learns this structure from the input domain.
-This seems to hold, according to our experiments.
-Still, it can cause inaccuracies in the estimation of the dimensional treshold.
-The restriction of the layers to a maximal abelian subgroup of $\text{GL}(n,\mathbb{R})$ solves this problem.
+We do not limit the processing of the neural net within the layers to an abelian Lie group.
+Instead, we assume that the net learns some map to a structure close to such an abelian connected Lie group with the estimated homology groups, on which the data could lie.
+This seems to hold, according to our exper<iments.
+Still, this method does only work exact in the case that the data lies on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural network to such a manifold.
 

### 2020-04-04 llncs2e/paper77.tex
paper

@@
 Instead, we assume that the net learns some map to a structure close to such an abelian connected Lie group with the estimated homology groups, on which the data could lie.
-This seems to hold, according to our exper<iments.
-Still, this method does only work exact in the case that the data lies on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural network to such a manifold.
+This seems to hold, according to our experiments.
+Still, this method does only approximate the true data manifold in the case that the data lies already on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural network to such a manifold.
 

### 2020-04-04 llncs2e/paper77.tex
sigmoid

@@
 We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = \frac{1}{n} \sum_{i=1}^{n} \left(\textbf{x}^{(0)}_i-\textbf{y}_i\right)^2$ as loss function.
-The leaky ReLu Eq.
-\ref{LeakyReLu} activation function $\sigma(\mathbf{x}^{(l)}) = \frac{1}{1 + e^{-\mathbf{x}^{(l)}}}$ is applied as elementwise activation function throughout the net.
+The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = \frac{1}{1 + e^{-\mathbf{x}^{(l)}}}$ is applied as elementwise activation function throughout the net.
 In all experiments dense neural nets are used with a bias term.

### 2020-04-04 llncs2e/paper77.tex
paper

@@
 This is crucial for the preservation of its topological structure and should therefore be parameterized correctly.
-We discuss a method to determine the smallest topology preserving layer for an invertible residual net.
-We consider the data domain as abelian connected Lie group and observe that it is decomposable into $\mathbb{R}^p \times \mathbb {T}^q$.
+We discuss a method to determine the smallest topology preserving layer considering the data domain as abelian connected Lie group and observe that it is decomposable into $\mathbb{R}^p \times \mathbb {T}^q$.
 Persistent homology allows us to count its $k$-th homology groups.

### 2020-04-04 llncs2e/paper77.tex
errors

@@
 Thus, the dimension of the data manifold is the dimension of the smallest possible bottleneck layer of the neural net  $\dim M = \min_l \dim \textbf{x}^{(l)}$ \citemain{HauserR17}.
-Our contribution provides the theory of how this dimension can be determined for all neural nets that are a composition of invertible layers acting on an abelian Lie group as domain of the data.
+Our contribution provides the theory of how this dimension can be determined for all neural nets acting on an abelian connected Lie group as domain of the data.
 The concatenation of such maps is called a deep neural net:
@@
 \ref{costfunctions} (b).
-This could be due to statistical errors in parametrization by mean confidence persistence landscapes.
+This is due to the fact that the neural network layers themself are not connected nor abelian and are not explicitly restricted to such a structure.
 The net showing the deflection of gradient is close to our threshold.
 We specify the parameterization for an embedding as well as for neural nets of the smoothness degree $1$--$4$.
-The neural net is capable of eliminating the additional structure of noise, which is no longer part of the abelian Lie group.
 \input{figure4.tex}
@@
 \label{sec:conclusion}
-Based on the theory of Lie groups and persistent homology a method for invertible resnets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group.
+Based on the theory of Lie groups and persistent homology a method for neural networks was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group.
 This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold.

### 2020-04-07 llncs2e/paper77.tex
paper title change2
typo

@@
 \hypersetup{
-    pdftitle={Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups},    % title
+    pdftitle={Parametrization of Neural Networks with Connected Abelian Lie Groups as Data Manifold},    % title
     pdfauthor={Luciano Melodia},     % author
-    pdfsubject={Invertible Neural Nets up to Homology Equivalence},   % subject of the document
+    pdfsubject={Neural Networks on Connected Abelian Lie Groups},   % subject of the document
     pdfcreator={Luciano Melodia},   % creator of the document
@@
 %
-\title{Invertible Resnets up to Homology Equivalence for Connected Abelian Lie Groups}
-%
-\titlerunning{Invertible Resnets up to Homology Equivalence}
+\title{Parametrization of Neural Networks with Connected Abelian Lie Groups as Data Manifold}
+%
+\titlerunning{Neural Networks on Connected Abelian Lie Groups}
 % If the paper title is too long for the running head, you can set
@@
 \ref{stats} shows the representatives of homology classes from their respective groups for cifar10 and cifar100.
-Further, it shows the estimated minimal embedding dimension for a manifold with these homology groups for $\mathcal{C}^1$ to $\mathcal{C}^4$-resnets.
+Further, it shows the estimated minimal embedding dimension for a manifold with these homology groups for $\mathcal{C}^1$ to $\mathcal{C}^4$-invertible resnets.
 The representatives for the $k$-th homology group $\sharp_k^{H}\left(\bar{\mathfrak{L}}\left(\mathfrak{P}(K(X))\right)\right)$ of the respective dataset are counted.

### 2020-05-06 llncs2e/paper77.tex
corrections equation
typos

@@
 \begin{abstract}
-Neural nets have been used in an elusive number of scientific disciplines.
-Nevertheless, their parameterization is largely unexplored.
+Neural nets have been used in an substantial number of scientific disciplines.
+Despite this, their parameterization is largely unexplored.
 Dense nets are the coordinate transformations of a manifold from which the data is sampled.
-After processing through a layer, the representation of the original manifold may change.
+After a layer processes the data, the representation of the original manifold may change.
 This is crucial for the preservation of its topological structure and should therefore be parameterized correctly.
@@
 
-We use invertible architectures, to demonstrate our hypothesis on a different Lie group from Euclidean space \citemain{NguyenAK19}.
+In the following we use invertible architectures, to demonstrate our hypothesis on a different Lie group from Euclidean space \citemain{NguyenAK19}.
 Our interpretation of a neural net is a continuous transformation, transferring invariant structures of the data in the input space to its deepest layer.
-Thus, it is capable to reconstruct the structure of the original space.
+Thus, it is able to reconstruct the structure of the original space.
 This approach corresponds to the concept of homotopy equivalence \citemain{RubinsteinR08,YangY93}, which means that there is a continuous map $f: X \rightarrow Y$ and $g: Y \rightarrow X$, such that $f \circ g$ is homotopic to the identity $\text{Id}_Y$ and $g \circ f$ to $\text{Id}_X$.
@@
 To do so, their homotopy groups must be equal.
-The fundamental group gives information about connected components which can be collapsed to a point by continuous functions, the first homotopy group about `holes' and the second about `cavities' and so forth.
+The fundamental group gives information about connected components.
+These can be collapsed to a point by continuous functions, the first homotopy group about `holes' and the second about `cavities' and so forth.
 But homotopies are difficult to compute \citemain{klaus2004quick,hatcher2002algebraic}.
@@
     \item In Sect.
-\ref{sec:persistenthomology} we introduce to the theory of persistent homology, define persistence diagrams, persistence landscapes and confidence sets, in order to estimate homology groups despite noisy data.
+\ref{sec:persistenthomology} we introduce the theory of persistent homology, define persistence diagrams, persistence landscapes and confidence sets, in order to estimate homology groups despite noisy data.
 We use Künneth's theorem and derive how to count the homology groups of the input space.
@@
 The manifold assumption states that a point $x = \{x_1, \cdots, x_d\} \in \mathbb{R}^n$ lies on an embedded manifold $M$, with $\dim M \ll n$.
-We specify the concept of smooth manifolds.
+First, we specify the concept of smooth manifolds.
 Then we dive into Lie groups concluding with a central theorem on the isomorphism of abelian connected Lie groups to the product space $\mathbb{R}^p \times \mathbb{T}^q$, which turns out useful to estimate the embedding dimension of the manifold.
@@
 Manifolds generalize the concept of space mathematically.
-They are described by a collection of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to $\mathbb{R}^{n}$ where $n = \text{dim} \; M$ and an open neighborhood $(U_i,\varphi_i)$ in $M$ which is covered by this map.
+They are described by a collection of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to $\mathbb{R}^{n}$, where $n = \text{dim} \; M$, and an open neighborhood $(U_i,\varphi_i)$ in $M$ which is covered by this map.
 The union of such maps should cover the entire space, i.e.
@@
 The $k$-chains are finite sums $\sum_{i=0}^{k} z_i \sigma_i$ with $z_i \in \mathbb{Z}$ \citemain[p.~108]{hatcher2002algebraic}.
-Having connected components defined as a group implies also the definition of a boundary.
+Having connected components defined as a group also implies the definition of a boundary.
 As the boundaries of a $k$-simplex are also $(k-1)$-simplices, we can remove all other simplices to project into the lower chain group by the $k$-th boundary operator $\partial_k: C_k(X) \rightarrow C_{k-1}(X)$:
@@
 We assign the value $\infty$ to it, corresponding to the second coordinate $w_{j+1}$ in the persistence diagram.
-The first term counts elements born in $K^{\omega_{j-1}}(X)$ and vanished entering $K^{\omega_{j}}(X)$, while the second term counts the representatives of homology classes before $K^{\omega_{j}}(X)$ and vanishing at $K^{\omega_{j}}(X)$.
+The first term counts elements born in $K^{\omega_{j-1}}(X)$ and which vanish entering $K^{\omega_{j}}(X)$, while the second term counts the representatives of homology classes before $K^{\omega_{j}}(X)$ and which vanish at $K^{\omega_{j}}(X)$.
 The index pairs denote the \emph{birth} and \emph{death} of a homology class.
@@
 \end{align}
-We need to know how to count in order to derive the dimension of the sought manifold from the representatives of homology classes within the mean persistence landscapes.
+We need to count the representatives of homology classes within the mean persistence landscapes in order to derive the dimension of the sought manifold.
 Note, in Eq.
@@
 	\label{countingsimple}
-	\sharp_{0}^{H}(\bar{\mathfrak{L}}(K(X))) \approx p.
+	\sharp_{0}^{H}(\bar{\mathfrak{L}}(\mathfrak{P}(K(X)))) \approx p.
 \end{align}
@@
 Commonly used are functions from the exponential family.
-A feedforward net is then a map $\varphi^{(l)}:\textbf{x}^{(l)}(M)$ $\rightarrow$ $\left(\varphi^{(l)}\circ \textbf{x}^{(l)}\right)(M)$.
+As a result a feedforward net is then a map $\varphi^{(l)}:\textbf{x}^{(l)}(M)$ $\rightarrow$ $\left(\varphi^{(l)}\circ \textbf{x}^{(l)}\right)(M)$.
 $\textbf{x}^{(0)} \in M$ are points on a Riemann manifold $M$.
@@
 \input{figure2.tex}
-Neural nets are used in practice with a different number of perceptrons per layer.
+In practice neural nets are used with a different number of perceptrons per layer.
 However, a layer, from the perspective of a matrix group, can also represent the manifold of the previous one.
@@
 
-\paragraph{Acknowledgements} The work is partially supported by Siemens Power and Gas. The authors would like to thank Noah Becker and David Haller for proofreading.
-
-\paragraph{Code} The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.
+\subsubsection*{Acknowledgements} The work is partially supported by Siemens Power and Gas. The authors thank Noah Becker and David Haller for proofreading.
+
+\subsubsection*{Code} The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.
 %
@@
 \begin{proof}
-We show that by induction and start with $\dim V = 1$.
+We show the Lemma by induction and start with $\dim V = 1$.
 Choosing a basis, we identify $V$ with $\mathbb{R}$.

### 2020-05-06 llncs2e/paper77.tex
correction

@@
 
-\paragraph{Persistence Landscapes} The persistence landscapes are used for prediction, computed on the persistence diagrams of the datasets cifar10 and cifar100 \citemain{krizhevsky2009learning}, see Fig.
+\subsubsection*{Persistence Landscapes} The persistence landscapes are used for prediction, computed on the persistence diagrams of the datasets cifar10 and cifar100 \citemain{krizhevsky2009learning}, see Fig.
 \ref{homgroups}.
@@
 
-\paragraph{Neural Nets} The data manifold is supposed to be some connected commutative Lie group.
+\subsubsection*{Neural Nets} The data manifold is supposed to be some connected commutative Lie group.
 We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.8 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$.

### 2020-05-06 llncs2e/paper77.tex
net

@@
 \hypersetup{
-    pdftitle={Parametrization of Neural Networks with Connected Abelian Lie Groups as Data Manifold},    % title
+    pdftitle={Parametrization of Neural Nets with Connected Abelian Lie Groups as Data Manifold},    % title
     pdfauthor={Luciano Melodia},     % author
-    pdfsubject={Neural Networks on Connected Abelian Lie Groups},   % subject of the document
+    pdfsubject={Neural Nets on Connected Abelian Lie Groups},   % subject of the document
     pdfcreator={Luciano Melodia},   % creator of the document
@@
 %
-\title{Parametrization of Neural Networks with Connected Abelian Lie Groups as Data Manifold}
-%
-\titlerunning{Neural Networks on Connected Abelian Lie Groups}
+\title{Parametrization of Neural Nets with Connected Abelian Lie Groups as Data Manifold}
+%
+\titlerunning{Neural Nets on Connected Abelian Lie Groups}
 % If the paper title is too long for the running head, you can set
@@
 
-\subsection{Residual Neural Networks}
+\subsection{Residual Neural Nets}
 \label{resnet}
@@
 \ref{costfunctions} (b).
-This is due to the fact that the neural network layers themself are not connected nor abelian and are not explicitly restricted to such a structure.
+This is due to the fact that the neural layers themself are not connected nor abelian and are not explicitly restricted to such a structure.
 The net showing the deflection of gradient is close to our threshold.
@@
 \label{sec:conclusion}
-Based on the theory of Lie groups and persistent homology a method for neural networks was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group.
+Based on the theory of Lie groups and persistent homology a method for neural nets was developed to parameterize them with good precision due to the assumption that the data lies on or near by some connected commutative Lie group.
 This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold.
@@
 This seems to hold, according to our experiments.
-Still, this method does only approximate the true data manifold in the case that the data lies already on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural network to such a manifold.
+Still, this method does only approximate the true data manifold in the case that the data lies already on a product space $\mathbb{R}^p \times \mathbb{T}^q$ or is mapped by the neural net to such a manifold.
 

### 2020-05-11 llncs2e/paper77.tex
added comment

@@
 The homology groups of the $q$-torus and the Euclidean space are well known.
-The application of fully invertible neural nets to resnets demonstrates our results on a different Lie group from Euclidean space.
+The application of fully invertible neural nets to resnets demonstrates our results on a different Lie group from Euclidean space and and guarantees that the experiment is learnt via the neural network on a coordinate representation up to the diffeomorphism cf.
+Fig.
+\ref{architecture}.
 By applying Künneth's theorem the homology groups of the topological factor spaces could be connected with the ones of their product space.

### 2020-05-26 llncs2e/paper77.tex
typo

@@
 \subsubsection*{Neural Nets} The data manifold is supposed to be some connected commutative Lie group.
-We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.8 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$.
+We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$ with $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$.
 These input vectors are trained to the original vector $\textbf{x}^{(0)}$.

### 2020-06-17 llncs2e/figure1.tex
update paper

@@
-\begin{figure}[t!]
-\begin{tikzpicture}[
-	bullet/.style={circle,
-	fill,
-	inner sep=2pt,
-	label={#1}},
-	>=latex,
-	scale=0.5
-]
-	\draw[color=gray] (0,0) circle (85pt);
-	\node[above, color=gray] at (2.5,2) {$\mathcal{S}^1$};
-	\node[above] at (-5,-3) {(a)};
-	\draw [black,line width=0.5pt,/.list={1/6,1/2,5/6}]
-	(210:3) node[color = brick, bullet={below left:1}] (a) {}
-	-- (90:3) node[color = brick, bullet={above:2}] (b) {}
-	-- (-30:3) node[color = brick, bullet={below right:3}] (c) {}
-	-- cycle;
-	\node[above] at (-5,2) {$K_0 = \{\{1\}, \{2\}, \{3\}\}$};
-	\node[above, fill=white] at (-4,1) {$K_1 = \{\{1,2\}, \{1,3\}, \{2,3\}\}$};
-	\node[above, fill=white] at (-5,0) {Coefficients in $\mathbb{Z}/2\mathbb{Z}$};
-\end{tikzpicture}
-\begin{tikzpicture}[
-	bullet/.style={circle,
-	fill,
-	inner sep=2pt,
-	label={#1}},
-	>=latex,
-	scale=0.5
-]
-	\node[above, color=gray] at (2.5,2) {$\mathcal{S}^1$};
-	\draw[color=gray] (0,0) circle (85pt);
-	\node[above] at (-5,-3) {(b)};
-	\draw [black,line width=0.5pt,->-/.list={1/6,1/2,5/6}]
-	(210:3) node[color = brick, bullet={below left:1}] (a) {}
-	-- (90:3) node[color = brick, bullet={above:2}] (b) {}
-	-- (-30:3) node[color = brick, bullet={below right:3}] (c) {}
-	-- cycle;
-	\draw[line width=0.5pt,-{Latex[bend]}] (240:0.5) arc(240:-60:0.5);
-	\node[above] at (-5,2) {$K_0 = \{\{1\}, \{2\}, \{3\}\}$};
-	\node[above, fill=white] at (-3.9,1) {$K_1 = \{\{1,2\}, \{2,3\}\}, \{1,3\}\}$};
-	\node[above] at (-5.7,0) {Coefficients in $\mathbb{R}$};
-\end{tikzpicture}
-\caption{Two triangulations of $\mathcal{S}^1$.
-We compute $H_0(\mathcal{S}^1) = \ker(\partial_0) / \Ima(\partial_1)$ for (a) and (b), see Sect.
-\ref{simcomplexhomtheory}.
-For (a): $\ker(\partial_0) = \text{span}(\{1\},\{2\},\{3\})$ and $\Ima(\partial_1)= \text{span}(\{\{2\}-\{1\}\},\{\{3\}-\{2\}\})$ as linearly independent elements.
-Observe that $\{1\}-\{3\}=\{2\}-\{1\}-(\{3\}-\{2\})$ holds.
-Thus $\ker(\partial_0) \simeq (\mathbb{Z}/2\mathbb{Z})^3$, $\Ima(\partial_1)=(\mathbb{Z}/2\mathbb{Z})^2$ and $H_0(\mathcal{S}^1;\mathbb{Z}/2\mathbb{Z}) = \mathbb{Z}/2\mathbb{Z}$.
-For (b): $\ker(\partial_0) = \text{span}(\{1\},\{2\},\{3\})$, but due to real coefficients the linear dependency is this time given by $\{2-1\} = \{3-2\}$.
-Thus $\ker(\partial_0) \simeq \mathbb{R}^3$ and $\Ima(\partial_1)=\mathbb{R}^2$, which yields the homology group $H_0(\mathcal{S}^1;\mathbb{R}) = \mathbb{R}$.}
-\label{complexes}
+\begin{figure}[t]
+    \begin{overpic}[width=\textwidth,tics=10]{figure1.pdf}
+        \put (0.9,58) {\textbf{a})}
+        \put (0.9,28) {\textbf{b})}
+        \put (70,58) {\textbf{c})}
+        \put (8.8,29.6) {$0$}
+        \put (32.8,29.6) {$0$}
+        \put (58,29.6) {$0$}
+        \put (74.8,43.6) {\small{1}}
+        \put (90.5,43.6) {\small{2}}
+        \put (84.6,22.4) {\small{3}}
+        \put (84.6,1) {\small{4}}
+        \put (7,13) {$K_1$}
+        \put (31,13) {$K_2$}
+        \put (57,13) {$K_3$}
+        \put (21,13) {$\subseteq$}
+        \put (45,13) {$\subseteq$}
+        \put (32,39) {$B_k$}
+        \put (32,49) {$Z_k$}
+        \put (32,55) {$C_k$}
+        \put (56,39) {$B_{k-1}$}
+        \put (56,49) {$Z_{k-1}$}
+        \put (56,55) {$C_{k-1}$}
+        \put (7,39) {$B_{k+1}$}
+        \put (7,49) {$Z_{k+1}$}
+        \put (7,55) {$C_{k+1}$}
+        \put (19.5,56) {$\partial_{k+1}$}
+        \put (44.5,56) {$\partial_{k}$}
+        \put (92,37) {{\color{red5} $\bullet$} $H_0$}
+        \put (92,34) {{\color{red4} $\bullet$} $H_1$}
+        \put (92,31) {{\color{red3} $\bullet$} $H_2$}
+        \put (92,28) {{\color{red2} $\bullet$} $H_3$}
+        \put (92,25) {$\cdots$}
+        \put (92,15.5) {{\color{red5} $\bullet$} $H_0$}
+        \put (92,12.5) {{\color{red4} $\bullet$} $H_1$}
+        \put (92,9.5) {{\color{red3} $\bullet$} $H_2$}
+        \put (92,6.5) {{\color{red2} $\bullet$} $H_3$}
+        \put (92,3.5) {$\cdots$}
+    \end{overpic}\\[0.1cm]
+    \caption{\textbf{a}) Illustration of the chain complex following Zomorodian et al.
+\cite{ZomorodianC05}.
+\textbf{b}) Sublevel sets of a simplicial complex connected by inclusion.
+\textbf{c}) Pipeline for persistent homology: \textbf{c})1.
+Loading point sets.
+\textbf{c})2.
+Computation of a filtration.
+\textbf{c})3.
+Computation of persistent homology with confidence band.
+\textbf{c})4.
+Computation of persistence landscapes.}
+    \label{figure1}
 \end{figure}

### 2020-06-17 llncs2e/figure2.tex
update paper

@@
 \centering
-\begin{tikzcd}
-& M\simeq \mathbb{R}^n \arrow[ld, "x^{(0)}", bend right = 15] 
-& U \simeq \mathbb{R}^{n-k} \arrow[ld, "x^{(1)}" pos=0.8, swap, bend right = 15] \arrow[d, "x^{(2)}"] \arrow[rd, "x^{(L-1)}", bend left = 15] \arrow[l, hook', swap, "\imath"] 
-&\\
-x^{(0)}(M) \arrow[r, "\varphi^{(0)}"] 
-& x^{(1)}(U) \arrow[r, leftrightarrow, "\varphi^{(1)}"]
-& x^{(2)}(U) \arrow[r, leftrightarrow, "\varphi^{(L-1)}", dotted]
-& x^{(L-1)}(U) \arrow[r, hook, "\varphi^{(L)}"]
-& x^{(L)}(M)
-\end{tikzcd}
-\caption{The architecture of the neural net maps over $\varphi^{(0)}$ into a subspace of $M$, in
-which the searched manifold $U$ of the data can be embedded. The maps $\varphi^{(1)}, \cdots, \varphi^{(L-1)}$
-are models of bijective, smooth and therefore diffeomorphic functions.}
-\label{architecture}
+\begin{tikzpicture}
+\begin{axis}[
+	height=2.2in,
+	width= 2.6in,
+	axis x line = bottom,
+	axis y line = left,
+	minor x tick num=10,
+	ylabel={$\frac{\epsilon_i + \epsilon_{j+1}}{2}$},
+	xlabel={\textbf{a}) Persistence landscape \texttt{cifar10}.},
+	xmin=0,
+	xmax=1050,
+	ymin=0,
+	ymax=650,
+	ticklabel style={font=\small},
+	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
+	y label style={at={(axis description cs:0.20,0.85)},rotate=-90,anchor=south, fill=white},
+	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
+]
+\coordinate (spypoint) at (axis cs:88,50);
+\coordinate (magnifyglass) at (axis cs:800,400);
+\draw [thick, dotted, draw=gray] (axis cs: 0,580) -- (axis cs: 1050,580) node[pos=0.9, below] {$\infty$};
+\end{axis}
+\begin{scope}[spy using outlines={circle, magnification=4, connect spies}]
+\node[inner sep=0pt] at (2.38,1.77) {\includegraphics[width=0.39\textwidth]{figure2_a.pdf}};
+\spy [lightgray, size=2cm] on (spypoint) in node[fill=white] at (magnifyglass);
+\end{scope} 
+\end{tikzpicture}
+\begin{tikzpicture}
+\begin{axis}[
+	height=2.2in,
+	width= 2.6in,
+	axis x line = bottom,
+	axis y line = left,
+	minor x tick num=10,
+	xlabel={\textbf{b}) Persistence landscape \texttt{cifar100}.},
+	xmin=0,
+	xmax=1050,
+	ylabel={$\cdot 10^{3}$},
+	ymin=0,
+	ymax=5,
+	ticklabel style={font=\small, fill=white},
+	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
+	y label style={at={(axis description cs:0.1,0.9)},rotate=-90,anchor=south},
+	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
+]
+\draw [thick, dotted, draw=gray] (axis cs: 0,4.45) -- (axis cs: 1050,4.45) node[pos=0.9, below] {$\infty$};
+\end{axis}
+\node[inner sep=0pt] (h) at (2.38,1.77) {\includegraphics[width=0.39\textwidth]{figure2_b.pdf}};
+\node[inner sep=0pt] (h) at (4.5,2.7) {{\color{red5}$\bullet$} {$H_0$}};
+\node[inner sep=0pt] (h) at (4.5,2.3) {{\color{red4}$\bullet$} {$H_1$}};
+\node[inner sep=0pt] (h) at (4.5,1.9) {{\color{red3}$\bullet$} {$H_2$}};
+\node[inner sep=0pt] (h) at (4.5,1.5) {{\color{red2}$\bullet$} {$H_3$}};
+\node[inner sep=0pt] (h) at (4.5,1.1) {{\color{red1}$\bullet$} {$H_4$}};
+\end{tikzpicture}
+\begin{tikzpicture}
+	\begin{axis}[
+	table/col sep=comma,
+	xlabel={\textbf{c}) Validation loss \texttt{cifar10}.},
+	height=2.2in,
+	width= 2.6in,
+   	ticklabel style={font=\small, fill=white},
+	xticklabel style={rotate=90, anchor=near xticklabel},
+	x label style={at={(axis description cs:0.5,-0.15)},anchor=north},
+	xtick={0,0.1,0.2,0.4,0.6,0.8,1},
+    xticklabels={dummy,0,50,100,150,200,250},
+	]
+	\end{axis}
+	\node[inner sep=0pt] at (2.51,2.08) {\includegraphics[width=0.343\textwidth, height=3.84cm]{figure2_c.pdf}};
+\end{tikzpicture}
+\begin{tikzpicture}
+	\begin{axis}[
+	table/col sep=comma,
+	xlabel={\textbf{d}) Validation loss \texttt{cifar100}.},
+	height=2.2in,
+	width= 2.6in,
+	ticklabel style={font=\small, fill=white},
+	xticklabel style={rotate=90, anchor=near xticklabel},
+	x label style={at={(axis description cs:0.5,-0.15)},anchor=north},
+	xtick={0,0.1,0.2,0.4,0.6,0.8,1},
+    xticklabels={dummy,0,50,100,150,200,250},
+	]
+	\end{axis}
+	\node[inner sep=0pt] at (2.51,2.08) {\includegraphics[width=0.343\textwidth, height=3.84cm]{figure2_d.pdf}};
+\end{tikzpicture}
+\caption{a) Persistence landscape for \texttt{cifar10} and b) for the \texttt{cifar100} dataset \cite{krizhevsky2009learning}.
+Persistence landscapes are computed up to a resolution of $10^3$.
+c) and d) show the \texttt{MSE} loss function on the validation dataset using a $7:3$ split.}
+\label{figure2}
 \end{figure}

### 2020-06-17 llncs2e/figure3.tex
update paper

@@
-\begin{figure}[t!]
-\begin{tikzpicture}
-\begin{axis}[
-	height=2.3in,
-	width= 2.7in,
-	axis x line = bottom,
-	axis y line = left,
-	minor x tick num=10,
-	ylabel={$\frac{\omega_i + \omega_{j+1}}{2}$},
-	xlabel={(a) cifar10.},
-	xmin=0,
-	xmax=1050,
-	ymin=0,
-	ymax=650,
-	ticklabel style={font=\small},
-	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
-	y label style={at={(axis description cs:0.20,0.85)},rotate=-90,anchor=south, fill=white},
-	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
-]
-\coordinate (spypoint) at (axis cs:85,50);
-\coordinate (magnifyglass) at (axis cs:800,400);
-\draw [thick, dotted, draw=gray] (axis cs: 0,580) -- (axis cs: 1050,580) node[pos=0.9, below] {$\infty$};
-\end{axis}
-\begin{scope}[spy using outlines={circle, magnification=4, connect spies}]
-\node[inner sep=0pt] at (2.55,1.9) {\includegraphics[width=0.42\textwidth]{figure3_a.pdf}};
-\spy [lightgray, size=2cm] on (spypoint) in node[fill=white] at (magnifyglass);
-\end{scope} 
-\end{tikzpicture}\hfill
-\begin{tikzpicture}
-\begin{axis}[
-	height=2.3in,
-	width= 2.7in,
-	axis x line = bottom,
-	axis y line = left,
-	minor x tick num=10,
-	xlabel={(b) cifar100.},
-	xmin=0,
-	xmax=1050,
-	ylabel={$\cdot 10^{3}$},
-	ymin=0,
-	ymax=5,
-	ticklabel style={font=\small, fill=white},
-	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
-	y label style={at={(axis description cs:0.1,0.9)},rotate=-90,anchor=south},
-	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
-]
-\draw [thick, dotted, draw=gray] (axis cs: 0,4.45) -- (axis cs: 1050,4.45) node[pos=0.9, below] {$\infty$};
-\end{axis}
-\node[inner sep=0pt] (h) at (2.55,1.9) {\includegraphics[width=0.42\textwidth]{figure3_b.pdf}};
-\node[inner sep=0pt] (h) at (4.5,3.1) {{\color{blacky}$\bullet$} {$H_0$}};
-\node[inner sep=0pt] (h) at (4.5,2.7) {{\color{gray}$\bullet$} {$H_1$}};
-\node[inner sep=0pt] (h) at (4.5,2.3) {{\color{purpur}$\bullet$} {$H_2$}};
-\node[inner sep=0pt] (h) at (4.5,1.9) {{\color{brick}$\bullet$} {$H_3$}};
-\node[inner sep=0pt] (h) at (4.5,1.5) {{\color{fire}$\bullet$} {$H_4$}};
-\end{tikzpicture}
-\caption{Mean confident persistence landscapes of the investigated datasets cifar10 and cifar100 with a resolution of $1000$.
-The mean persistence landscapes were computed on truncated persistence diagrams, with a confidence of $\alpha = 0.95$ for a representative of a persistent homology class tracked on the filtration.}
-\label{homgroups}
-\end{figure}
+\begin{table*}[t!]
+\centering
+\caption{Counts of the representatives per homology group from $\text{PL}(X)$.}
+\label{figure3}
+\begin{tabular}{@{}rcccccccccccccc@{}}
+\toprule \multirow{2}{*}{\backslashbox{\texttt{Data} \strut}{\strut \texttt{Features}}} & \phantom{abc} & \multicolumn{5}{c}{\texttt{Homology groups}} &&& \multicolumn{6}{c}{\texttt{$\approx$ embedding dimension}}\\
+\cmidrule{3-7} \cmidrule{10-15} & & $H_0$ & $H_1$ & $H_2$ & $H_3$ & $H_4$ &&& $p$ & $q \vert {H_1}$ & $q \vert {H_2}$ & $q \vert {H_3}$ & $q \vert {H_4}$ & $\dim U$\\
+\midrule
+\texttt{cifar10} && 12 & 16 & 40 & 59 & 50 &&& $12$ & $16$ & $9 \scriptstyle{\pm 4}$ & $8 \scriptstyle{\pm 3}$ & $7 \scriptstyle{\pm 15}$ & $92 \scriptstyle{\pm 44}$\\
+\texttt{cifar100} && 13 & 18 & 34 & 46 & 48 &&& $13$ & $18$ & $9 \scriptstyle{\pm 2}$ & $8 \scriptstyle{\pm 10}$ & $7 \scriptstyle{\pm 13}$ & $97 \scriptstyle{\pm 50}$\\
+\midrule
+\end{tabular}
+\label{stats}
+\end{table*}

### 2020-06-17 llncs2e/samplepaper.tex
update paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads,orivec]{llncs}

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{euler}
\usepackage{amsmath,amssymb,amscd,amstext}
\usepackage{xcolor}
\usepackage{pst-node}
\usepackage{tikz,pgf}
\usepackage{tikz-cd} 
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\pgfplotsset{compat=newest}
\usetikzlibrary{spy,decorations.markings,decorations.text,arrows.meta,bending,3d,shapes.arrows,positioning,fit,backgrounds,fadings}
\let\clipbox\relax
\usepackage{slashbox}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{adjustbox}
\usepackage{setspace}
\usepackage{xcolor}
\usepackage{multibib}
\usepackage{filecontents}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{mathtools}
\usepackage{commath}
\usepackage[percent]{overpic}
\usepackage[hidelinks]{hyperref}
\hypersetup{
    pdftitle={Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory},    % title
    pdfauthor={Luciano Melodia},     % author
    pdfsubject={Estimate of the Neural Net Dimension},   % subject of the document
    pdfcreator={Luciano Melodia},   % creator of the document
    pdfproducer={Luciano Melodia},  % producer of the document
    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {Neural nets} {Manifold learning}, % list of keywords
    pdfnewwindow=true,
    colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=purple
}
\definecolor{red1}{RGB}{227,226,223}
\definecolor{red2}{RGB}{227,175,188}
\definecolor{red3}{RGB}{238,76,124}
\definecolor{red4}{RGB}{154,23,80}
\definecolor{red5}{RGB}{93,0,30}
\makeatletter
\newcommand{\@chapapp}{\relax}
\newcommand{\xdashrightarrow}[2][]{\ext@arrow 0359\rightarrowfill@@{#1}{#2}}
\newcommand{\xdashleftarrow}[2][]{\ext@arrow 3095\leftarrowfill@@{#1}{#2}}
\newcommand{\xdashleftrightarrow}[2][]{\ext@arrow 3359\leftrightarrowfill@@{#1}{#2}}
\def\rightarrowfill@@{\arrowfill@@\relax\relbar\rightarrow}
\def\leftarrowfill@@{\arrowfill@@\leftarrow\relbar\relax}
\def\leftrightarrowfill@@{\arrowfill@@\leftarrow\relbar\rightarrow}
\def\arrowfill@@#1#2#3#4{%
  $\m@th\thickmuskip0mu\medmuskip\thickmuskip\thinmuskip\thickmuskip
   \relax#4#1
   \xleaders\hbox{$#4#2$}\hfill
   #3$%
}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
\makeatother
\begin{document}
%
\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\titlerunning{Estimate of the Neural net Dimension}

\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
\and Richard Lenz\orcidID{0000-0003-1551-4824}}
%
\authorrunning{L. Melodia et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Chair of Computer Science 6\\
Friedrich-Alexander University Erlangen-Nürnberg\\
91058 Erlangen, Deutschland \\
\email{\{luciano.melodia,richard.lenz\}@fau.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
In this paper we present an approach to determine the smallest possible number of perceptrons in a neural net in such a way that the topology of the input space can be learned sufficiently well. We introduce a general procedure based on persistent homology to investigate topological invariants of the manifold on which we suspect the data set. We specify the required dimensions precisely, assuming that there is a smooth manifold on or near which the data are located. Furthermore, we require that this space is connected and has a commutative group structure in the mathematical sense. These assumptions allow us to derive a decomposition of the underlying space whose topology is well known. We use the representatives of the $k$-dimensional homology groups from the persistence landscape to determine an integer dimension for this decomposition. This number is the dimension of the embedding that is capable of capturing the topology of the data manifold. We derive the theory and validate it experimentally on toy data sets.

\keywords{Embedding dimension \and Parameterization \and Persistent homology \and Neural nets \and Manifold learning.}
\end{abstract}

\section{Introduction}
Since the development of deep neural nets, their parameterization, in particular the smallest possible number of perceptual units in a layer, has been studied. This number is of importance for auto-encoding tasks that require extrapolation or interpolation of data, such as blind source separation or super-resolution. If this number is overestimated, unnecessary resources are consumed during training. If the number is too small, no sufficiently good estimate can be given. To solve this problem, we make the following contribution in this paper:
\begin{itemize}
    \item We study topological invariants using statistical summaries of persistent homology on a filtered simplicial complex on the data.
    \item Using the theory of Lie groups, we specify the smallest possible embedding dimension so that the neural net is able to estimate a projection onto a space with the determined invariants.
\end{itemize}

\section{Related work}
The related work is divided into two sections. In the first section we present chronologically the milestones in the consideration of neural nets from a differential geometric perspective. In the second section we present work on earlier attempts to determine their minimum embedding dimension.

\subsubsection*{Riemannian geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean. Very recently, there have been results in the direction of other manifolds on which neural nets can operate. The spherical neural nets operate on a manifold with commutative group structure \cite{CohenGKW18,EstevesAMD20}.

A metric is required to describe the geometric properties of such a manifold. The change of coordinate systems and the learned metric tensor of a Riemann manifold during back propagation of neural nets has been investigated \cite{HauserR17}. Further, it has been shown that dense neural nets can't approximate all functions arbitrarily precisely \cite{Johnson19}. In response to this result, neural nets with residual connections have been examined. Residual nets add the output of one layer to a deeper one and therefore bypass some of the layers in between. This type of neural net is indeed an universal function approximator \cite{LinJ18}.

\subsubsection*{Embedding dimension of neural nets} Cybenko showed that dense neural nets with a single hidden layer can approximate any continuous function with a bounded domain with arbitrary small error \cite{Cybenko92}. The upper bound of the embedding dimension for ReLU nets has recently been proven to be $(n+4)$ with $n$ being the sample dimension \cite{RaghuPKGS17}. Bartlett et al. show that a nearly-tight bound for the VC dimension of a neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.

The depth of a neural net has been determined to be a parameter depending on the moduli of continuity of a function to be approximated \cite{LinJ18}. Inspired by the theory of finite difference methods, such a residual layer has been defined as a forward or backward difference operator on a partition of the layers. The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}. Thus, such a layer is capable of embedding into $k\cdot n$ dimensional space. The view that a neural net operates on a smooth manifold inspired us to investigate the invariants of this manifold for the theory discussed in this paper.

To these fundamental results we contribute a topological approach to parameterization problems. Similar to Futagami et al. \cite{FutagamiYS19}, we study the manifold on which we suspect the data. Our approach is applicable to topological spaces in general and is based on persistent homology. We are looking for the dimension of a space with the same topological properties as data indicates and of which we know that it can be approximated. We use the simplest decomposition into a product space of realaxes and $1$-spheres to obtain the same homology groups as the filtration of the data has. Thus, we obtain a computable lower bound for the embedding dimension of any neural net with differentiable structure.

\newpage
\section{Smooth manifolds, Lie groups and persistent homology}
The assumption that a set of points lies on a manifold is more accurate than handling points in Euclidean space. A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that at any two points $x,y \in M$ there are always two open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty. This behaviour provides the intuitive feeling of the position of points in space. The topology $(M, \tau)$ is a set system that describes the structure of a geometric object. Thereby $M$ itself and $\emptyset$ must be contained in $\tau$, and any arbitrary union of open sets from $\tau$ and any intersections from $\tau$ must be in $\tau$. If this set system has a smaller set that generates the topology $\tau$ with the aid of finitely many unions, then this smaller set is a \emph{basis} of the topology. We demand of a manifold that the basis is at most countable. At last, $M$ is locally Euclidean, i.e. each point has a neighborhood which can be homeomorphically mapped to a subset of $\mathbb{R}^n$. The integer $n$ is the dimension of $M$. Further, we demand some additional smooth structure.

\subsection{Smooth manifolds} A smooth structure is a stronger condition than a topological manifold which can be described by a family of continuous functions. This assumption can be justified by the theorem of Stone-Weierstrass which proves that any continuous function can be approximated arbitrarily precisely by a smooth one \cite{stone1948generalized}. Smooth manifolds are described by a collection of local coordinate charts $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$, homeomorphic to $\mathbb{R}^{n}$ and an open neighborhood $(U_i,\varphi_i)$ in $M$ which is covered by this map. The family of such maps that cover $M$ is the \textit{atlas} $\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}. The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$. If the atlas is maximal with respect to inclusion, then it is a \emph{differentiable structure}. Each atlas for $M$ is included in a maximal atlas. Because of the differentiable structure, the maps in $\mathscr{A}$ are also compatible with all other maps of $\mathscr{A}_{\text{max}}$. Thus, it is sufficient to describe smooth manifolds by an atlas that is not maximal.

As long as the activation functions are smooth and of the same differentiable structure, they are also compatible and the manifold does not change while propagating the data trough the layers. Only the coordinate representation changes.

\subsubsection{Lie groups} The study of group theory deals with symmetries which can be expressed algebraically. A pair $(M,\bullet)$, consisting of a map $\bullet: M\times M \rightarrow M$ is called group if the map is associative, $x \bullet (y \bullet z) = (x \bullet y) \bullet z$, has a neutral element so that $x \bullet e_{M} = x$ and has an inverse element to each element of the group, $x \bullet x^{-1} = e_{M}$. A Lie group is a smooth manifold which is equipped with a group structure such that the maps $\bullet:M \times  M \rightarrow  M, (x,y) \mapsto xy, $ and $\imath: M \rightarrow  M, x \mapsto x^{-1}$ are smooth. We call a space \emph{connected} if it cannot be divided into disjoint open neighbourhoods. A group is called Abelian or commutative when all elements under the group operation commute.

We apply a theorem from Lie theory, which states that each connected Abelian Lie group $(M,\bullet)$ of dimension $\dim M = n$ is isomorphic to a product space $\mathbb{R}^p \times \mathbb{T}^q$ with $p+q =n$, for a proof we refer to \cite[p.~116]{onishchik1993lie}. The $q$-torus $\mathbb{T}^q$ is a surface of revolution, which means it moves a curve around an axis of rotation. These axes are given by $1$-spheres, such that the $q$-torus is a product space $\mathbb{T}^q = \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$. The initial decomposition of connected commutative Lie groups can be further simplified into $M \simeq \mathbb{R}^p \times (\mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q)$. Recall, that the $(n-1)$-sphere is given by $\mathcal{S}^{n-1} := \{x \in \mathbb{R}^n \; \vert \; \norm{x}_2 = 1\}$. Thus, $\mathcal{S}^1$ can be embedded into a two-dimensional Euclidean space. Under our assumption, we can derive how many dimensions are at least needed for a suitable embedding. For each $1$-sphere we count two dimensions and for each real line accordingly one. Finally, we have to estimate from data how many factor spaces are necessary to represent the topology of the underlying data manifold sufficiently accurately.

\subsection{Persistent homology}
Algebraic topology provides a computable tool to study not the topology of a set of points itself but an Abelian group which is attached to the topology. The core interest in this discipline lies in \emph{homotopy equivalences} of an equivalence class to which objects belong that are continuously deformable into one another. This term is much broader than a homeomorphism. For two topological spaces $X$ and $Y$ we are looking for a function $h: X \times I \rightarrow Y$, which gives the identity at the time $h_0(X) = X$ and for $h_1(X) = X$ a mapping into the other topological space. If this mapping is continuous with respect to its arguments, it is called \emph{homotopy}. Let us consider two functions $f,g: I \rightarrow X$ so that $f(1) = g(0)$. Then there is a \emph{composition} of product paths $f \cdot g$, which pass first through $f$ and then through $g$ and which is defined as $f \cdot g (s) = f(2s)$ for $0 \leq s \leq 1/2$ and $g(2s - 1)$, for $1/2 \leq s \leq 1$. We first run $f$ at double speed up to $1/2$ and from $1/2$ to $1$ we run the function $g$ at double speed. In addition, suppose that we have a family of functions $f: I \rightarrow X$ that have the same start and end point $f(0) = f(1) = x_0 \in X$. These functions intuitively form a \emph{loop}. The set of all homotopy classes $[f]$ of loops $f: I \rightarrow X$ at base point $x_0$ is noted as $\pi_1(X,x_0)$ and is called first homotopy group or fundamental group. The group operation is the product $[f][g] = [f \cdot g]$ of equivalence classes. If we do not look at the interval $I$, but at mappings considering the unit cube $I^n$, we obtain the $n$-th homotopy group $\pi_n(X,x_0)$ by analogy. With the help of the homotopy groups we can study the connected components of a topological space for the $0$th group, the loops for the $1$th group, the cavities for the $2$th and so forth. Since homotopy groups are very difficult to compute, we resort to an algebraic variant, the \emph{homology groups}.

\subsubsection{Simplices} In data analysis, we do not study the topological spaces themselves, but points that we assume are located on or near this space. However, all closed surfaces can be triangulated, i.e. completely covered with simplices. For a proof and illustrations we refer to \cite[p.~102]{hatcher2002algebraic,lume}. The concept of the triangle is too specific, so the \emph{$n$-simplices} are defined as a generalization. They are the smallest convex set in Euclidean space with $(n+1)$-points $v_0,\cdots,v_n$, having no solutions for any system of linear equations in $n$-variables. Thus, we say they lie in \emph{general position with respect to $\mathbb{R}^n$}, because they do not lie on any hyperplane with dimension less than $n$ \cite[p.~103]{hatcher2002algebraic}. We define the $n$-simplex as follows:
\begin{equation}
    \label{simplex}
    [\sigma] = \left\{ [v_0,\cdots,v_n] \in \mathbb{R}^{n+1} \; \bigg\vert \; \sum_{i=0}^{n} v_i = 1 \; \text{and} \; v_i \geq 0 \; \text{for all} \; i \right\}.
\end{equation}
Removing a vertex from $[\sigma]$ results in a $(n-1)$-simplex called \emph{face} of $[\sigma]$. A $0$-simplex is a point, a $1$-simplex is a path between two $0$-simplices, a $2$-simplex is an area enclosed by three $1$-simplices and so on. So they follow the intuition and generalize triangles to polyhedra, including points by definition.

\input{figure1.tex}
\subsubsection{Simplicial complexes} A \emph{simplicial complex} $K$ is a set of simplices, such that any face of a simplex of $K$ is a simplex of $K$ and the intersection of any two simplices of $K$ is either empty or a common face of both \cite[p.~11]{boissonnat2018geometric}. Further, we require every $(k+1)$-element, with $k < n$ to be a $k$-simplex of $[\sigma] \subset K$. This definition means that the simplicial complex is also a topological space. The properties described ensure that simplices are added to this set in a certain way, namely so that they are only \emph{glued} edge to edge, point to point, and surface to surface. Thus, every $n$-simplex has $n+1$ distinct vertices and no other $n$-simplex has the same set of vertices.  This is a unique combinatorial description of vertices together with a collection of sets $\{[\sigma]_1, \cdots, [\sigma]_m\}$ of $k$-simplices, which are $k+1$ element subsets of $K$ \cite[p.~107]{hatcher2002algebraic}. This set system can be realized geometrically with a homeomorphism, by a corresponding mapping into an Euclidean space.

\subsubsection{Associated Abelian groups} Algebraically, the simplices can be realized as a system of linear combinations as $\sum_i \lambda_i [\sigma]_{i}$ following Eq. \ref{simplex}. Together with addition as an operation, they form a group structure called $k$th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used. The group is commutative or Abelian. The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges is given here by negative and positive signs, so that $[\sigma] = -[\tau]$, iff $[\sigma] = [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations. The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$. The groups are connected by a homomorphism, a mapping which respects the algebraic structure, the boundary operator: $\partial_k: C_k \rightarrow C_{k-1}$, $\partial_k [\sigma] \mapsto \sum_i (-1)^i [v_0,v_1,\cdots,\hat{v}_i, \cdots, v_n]$. The $i$th face is omitted, alternatingly. Using the boundary operator we yield the following sequence of Abelian groups, which is a chain complex cf. Fig. \ref{figure1} a):
\begin{equation}
    0 \xrightarrow[]{\partial_{k+1}} C_k \xrightarrow[]{\partial_{k}} C_{k-1} \xrightarrow[]{\partial_{k-1}} \cdots \xrightarrow[]{\partial_{2}} C_1 \xrightarrow[]{\partial_{1}} C_0 \xrightarrow[]{\partial_{0}} 0.
\end{equation}

For the calculation of homology groups we have to choose a field from which to obtain the coefficients $\lambda_i$. Since, according to the Universal Coefficient Theorem, all homology groups are completely determined by homology groups with integral coefficients, we choose coefficients in $\mathbb{Z}$, see the proof in \cite{gruenberg1968universal}.

The $k$th chain groups contain two more Abelian subgroups that behave normal to it. First, the so-called cycle groups $\langle Z_k, +\rangle$, which are defined as $Z_k := \ker \partial_k = \{[\sigma] \in C_k \; \vert \; \partial_k [\sigma] = \emptyset \}$. This follows the intuition of all elements that form a loop, i.e. have the same start and end point. Some of these loops have the peculiarity of being a boundary of a subcomplex. These elements form the boundary group $\langle B_k, +\rangle$, defined by $B_k := \text{im } \partial_{k+1} = \{[\sigma] \in C_k \; \vert \; \exists [\tau] \in C_{k+1}: [\sigma] = \partial_{k+1}[\tau]\}$. Now we define the homology groups as quotient of groups. Homology -- analogous to homotopy theory -- gives information about connected components, loops and higher dimensional holes in the simplicial complex:
\begin{equation}
    H_k(K) := \frac{\ker \partial_k C_k(K)}{\text{im } \partial_{k+1} C_{k+1}(K)} = \frac{Z_k(K)}{B_k(K)}.
\end{equation}

\subsubsection{Homological persistence} Examining the homology groups of a set of points gives little information about the structure of a dataset. Instead, we are interested in a parametrization of the simplicial complex as geometric realization in which the homology groups appear and disappear again. For this purpose we consider all possible subcomplexes that form a \emph{filtration} over the point set $X$. We denote $K^{\epsilon} := K^{\epsilon}(X)$. Depending on how we vary the parameter $\epsilon_i$ of the chosen simplicial complex, the following sequence is generated, connected by inclusion and starting with the empty set cf. Fig. \ref{figure1} b):
\begin{align}
    \emptyset &= K^{\epsilon_0} \subset K^{\epsilon_1} \subset \cdots \subset K^{\epsilon_{n+1}}= K,\\
    K^{\epsilon_{i+1}} &= K^{\epsilon_i} \cup \sigma^{\epsilon_{i+1}}, \quad \text{for} \; i \in \{0, \cdots, n-1\}.
\end{align}
The filtration has a discrete realization with a fixed $\epsilon = \epsilon_{i+1}-\epsilon_{i} = \min(x,y)$ for all $x,y \in X$. Through filtration we are able to investigate the homology groups during each step of the parameterization. We record when elements from a homology group appear and when they disappear again. Intuitively speaking, we can see when $k$-dimensional holes appear and disappear in the filtration. We call this process \emph{birth} and \emph{death} of topological features. Recording the \emph{Betti numbers} of the $k$-th homology group along the filtration, we obtain the \emph{$k$-dimensional persistence diagram}, see Fig. \ref{figure1} c)3. The Betti numbers $\beta_k$ are defined by $\text{rank } H_k$. We write $H^{\epsilon_i}_k$ as $k$-th homology group on the simplicial complex $K$ with parametrization $\epsilon_i$. Then $H^{\epsilon_{i}}_{k} \rightarrow H^{\epsilon_{i+1}}_k$ induces a sequence of homomorphisms on the filtration, for a proof we refer to \cite{edelsbrunner2008persistent}:
\begin{align}
0 = H^{\epsilon_{0}}_k \rightarrow H^{\epsilon_{1}}_k \rightarrow \cdots \rightarrow H^{\epsilon_{n}}_k \rightarrow H^{\epsilon_{n+1}}_k = 0.
\end{align}
The image of each homomorphism consists of all $k$-dimensional homology classes which are born in $K^{\epsilon_i}$ or appear before and die after spanning $K^{\epsilon_{i+1}}$. Tracking the Betti numbers on the filtration results into a multiplicity
\begin{equation}
\mu^{\epsilon_{i},\epsilon_{j}}_k = (\beta_k^{\epsilon_{i},\epsilon_{j-1}} - \beta_k^{\epsilon_i,\epsilon_{j}})-(\beta_k^{\epsilon_{i-1},\epsilon_{j-1}}-\beta_k^{\epsilon_{i-1},\epsilon_{j}})
\end{equation}
for the $k$-th homology group and index pairs $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2} := \mathbb{R}^2 \cup \infty$ with indices $i \leq j$. The Euclidean space is extended, as the very first connected component on the filtration remains connected. Thus, we assign to it infinite persistence, corresponding to the second coordinate $\epsilon_{j+1}$. The first term counts elements born in $K^{\epsilon_{j-1}}$ and which vanish entering $K^{\epsilon_{j}}$, while the second term counts the representatives of homology classes before $K^{\epsilon_{j}}$ and which vanish at $K^{\epsilon_{j}}$. The $k$-th persistence diagram is then defined as
\begin{equation}
\text{PH}_{k}(X) := \left\{(\epsilon_i, \epsilon_{j+1}) \in \overline{\mathbb{R}^2} \; \bigg\vert \; \mu^{\epsilon_{i},\epsilon_{j+1}}_k = 1 \; \text{for all} \; i,j \in I \right\}.
\end{equation}

\subsubsection{Persistence landscapes} Persistence landscapes give a statistical summary of the topology of a set of points embedded in a given topological manifold \cite{bubenik2015statistical,bubenikD17}. Looking at the points $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2}$ on the $k$-th persistence diagram $\text{PH}_{k}(X)$, one associates a piecewise linear function $\lambda^{\epsilon_i}_{\epsilon_{j+1}}: \mathbb{R} \rightarrow [0,\infty)$ with those points:
\begin{align}
    &\text{If } x \not\in (\epsilon_i,\epsilon_{j+1}), \quad \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0,\\
    &\text{if } x \in \left(\epsilon_{i},(\epsilon_i + \epsilon_{j+1})/2 \right], \quad \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = x - \epsilon_{i} \; \text{and}\\
    &\text{if } x \in \left((\epsilon_i + \epsilon_{j+1})/2,\epsilon_{j+1} \right), \quad \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = \epsilon_{j+1} - x.
\end{align}
The summaries for the general persistence diagram are the disjoint union of the $k$-th persistence diagrams $\text{PH}(X) := \coprod_{i=0}^{k} \text{PH}_i(X)$. A persistence landscape $\text{PL}(X)$, contains the birth-death pairs $(\epsilon_i,\epsilon_{j+1})$, for an $i,j \in \{1, \cdots, n\}$ and is the function sequence $\Lambda_k : \mathbb{R} \rightarrow [0,\infty)$ for a $k \in \mathbb{N}$, where $\Lambda_k(x)$ denotes the $k$-th greatest value of $\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x)$, see Fig. \ref{figure1} c)4. Thus, $\Lambda_k(x) = 0$ if $k > n$. $\text{PL}(X)$ lies in a completely normed vector space, suited for statistical computations.

\section{Neural nets}
\label{sec:neuralnetworks}
Neural nets are a composition of \emph{affine transformations} with a non-linear activation function. This transformation obtains collinearity. It also preserves parallelism and partial relationships. The projection onto the $(l+1)$-th layer of such a network can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural net.

The linear map $\textbf{x} \mapsto \mathbf{W} \textbf{x}$ and the statistical distortion term $\textbf{x} \mapsto \textbf{x}+\textbf{b}$ are determined by stochastic gradient descent.  Note, that the linear transformation $\textbf{x} \mapsto \mathbf{W} \cdot \textbf{x}$ can be interpreted as the product of a matrix $\mathbf{W}_{ij} = \mathbf{\delta}_{ij}\mathbf{W}_i$ with the input vector $\textbf{x}$ using the Kronecker-$\delta$, while $(\cdot)$ denotes elementwise multiplication. Commonly used are functions from the exponential family. As a result a neural net is a map $\varphi^{(l)}:\textbf{x}^{(l)}(M)$ $\rightarrow$ $(\varphi^{(l)} \circ \textbf{x}^{(l)})(M)$. Learning by back propagation can in this way be considered as a change of coordinate charts of a smooth manifold. For a detailed formulation of back propagation as a shift on the tangent bundle of this manifold we refer to Hauser et al. \cite{HauserR17}.

In practice neural nets are used with a different number of perceptrons per layer. However, a layer can represent the manifold of the previous one. Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map does not change. Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural net \cite{HauserR17}.

\section{Counting Betti numbers}
Based on our assumption that data lies on a connected Abelian Lie group, we can decompose the data space into a product of realaxes and tori. In the following we refer to this Lie group as $G \simeq \mathbb{R}^p \times \mathbb{T}^q$. By this decomposition there is also an isomorphism of the homology groups of $G$ with those of $\mathbb{R}^p \times \mathbb{T}^q$, so that $H_k(G) \simeq H_k(\mathbb{R}^p \times \mathcal{S}^{1}_{1}\times\cdots\times\mathcal{S}^{1}_{q})$ with $p+q = \dim G$. We demand from a neural net that these homology groups can all be approximated, i.e. that all topological structure of the input space can be represented sufficiently well.

One might wonder to what extent the homology groups of a simplicial complex can be used to estimate the homology groups of the manifolds underlying the data. In this section we denote the simplicial homology groups as $H^{\Delta}_{\bullet}(X)$. The $\bullet$ is a proxy for all possible indices. Through the construction of continuous mappings, so-called singular simplices, from a simplex into a topological space, the singular homology groups $H_{\bullet}(X)$ can be constructed. We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a detailed introduction with proofs. This allows to assign Abelian groups to any topological space. If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{\bullet}(X)$, see for a proof \cite[pp.~473]{lee2013smooth}. Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{\bullet}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}. In summary, we have homology theories for simplicial complexes, topological spaces, a smooth variant on topological spaces and on smooth manifolds. We consider simplicial homology to be determined on the basis of a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space. However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}. This legitimizes our approach.

Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space $X \times Y$ is isomorphic to the direct sum of the tensor product of its $k$th homology groups, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
\begin{align}
    H_k(X \times Y) \simeq \bigoplus_{i+j=k}H_i(X) \otimes H_j(Y).
\end{align}
This applies to all fields, since modules over a field are always free. If we apply the theorem to the decomposition, we get for the space we are looking for
\begin{align}
    \label{homologykunneth}
    &H_k(\mathcal{S}^{1}_{1}\times\cdots\times \mathcal{S}^{1}_{q}\times\mathbb{R}^p) \simeq \\
    \bigoplus_{i_{1} + \cdots + i_{r}=k} &H_{i_{1}}(\mathcal{S}^{1}_{1}) \otimes \cdots \otimes H_{i_{\hat{r}}}(\mathcal{S}^{1}_{q}) \otimes H_{i_{r}}(\mathbb{R}^{p}).
\end{align}
By the isomorphism of simplicial, singular, smooth singular and De Rham cohomology, we use the theorem in accordance to our assumption. Thus, we count the representatives of homology classes in the persistence landscapes to derive the dimension of the sought manifold. The homology groups of the $1$-sphere are
\begin{align}
    H_0(\mathcal{S}^1) &= H_1(\mathcal{S}^1) = \mathbb{Z},\\
    H_i(\mathcal{S}^1) &= 0, \; \text{for all } i \geq 2.
\end{align}
Note, in Eq. \ref{homologykunneth} terms remain only for indices $i_{j} \in \{0,1\}$. Thus, we get
\begin{align}
    H_0(\mathbb{R}^p) &\simeq\mathbb{Z}^{p},\\
    \label{determineq}
    H_k(\mathbb{T}^{q}) &\simeq H_k(\mathcal{S}^{1}_{1} \times \cdots \times \mathcal{S}^{1}_{q}) \simeq \mathbb{Z}^{{q\choose{k}}},
\end{align}
where $p$ indicates the number of connected components. The number of local maxima for a homology group in the persistence landscape is the cardinality of the set of points $(\epsilon_i,\epsilon_{j+1})$ with derivatives $d\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0 \; \text{and} \; d^2 \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) < 0$, see Fig. \ref{figure2} b) c). Also, it is the solution for the binomial coefficient:
\begin{align}
     {q \choose k} = \frac{q}{1} \cdot \frac{q-1}{2} \cdots \frac{q-(k-1)}{k} = \prod_{i=1}^{k} \frac{q+1-i}{i}.
\end{align} 
We count in $PL(X)$ the elements of the $0$th homology group. The amount of elements from higher homology groups correspond to $q$ (Eq. \ref{determineq}) and are computed using Monte-Carlo methods for an (approximate) integer solution, see Tab. \ref{figure3}.

\section{Experimental setting}
We train an auto-encoder using as input heavily noisy images and map them to their noiseless original. We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$. The embedding dimension is increased for each experiment by $2$, i.e. $2,4,6, \cdots, 784$. $392$ neural nets have been trained $\times 100$. Each line of Fig. \ref{figure2} c), d) represents one neural architecture as averaged loss function.
\input{figure2.tex}
\subsubsection*{Persistence landscapes} We use the Delaunay-complex for the filtration, according to Melodia et al. \cite{lume}. The maximum $\alpha$-square is set to $e^{12}$. The maximal expansion dimension of the simplicial complex is set to $10$. The maximal edge length is set to $0.1$. The persistent landscapes are smoothened by a Gaussian filter $G(x) = 1 / \sqrt{2 \pi \sigma^2} \cdot e^{- x^2 / 2 \sigma^2}$ with $\sigma = 2$ for visualization purposes. We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}. Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.
\subsubsection*{Hyperparameters} The batch size is set to $128$, which should help to better recognize good parametrization. Higher batch size causes explosions of the gradient, which we visualize in Fig. \ref{figure2} c), d). For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$. The dataset is randomly shuffled. We implement in \texttt{Python} $3.7$. We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}. The neural training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.
\input{figure3.tex}
\subsubsection*{Neural net} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function. The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold. Dense neural nets are used with bias term. The datasets are flattened from $(32,32,3)$ into $(3072)$. The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e. $\varphi^{(0)}(x^{(0)}) =$ $\sigma(\sum_{i=1}^{n} \textbf{W}_{ji}^{(1)} \textbf{x}_{i}^{(0)} + \textbf{b}_{i}^{(1)})$. It transfers the data to the desired embedding dimension. The subsequent layers are implemented as residual invertible layers following Dinh et al. \cite{DinhSB17}. We use invertible architectures to avoid degenerated cases and to stay on the same differentiable structure during learning within the layers to be tested. A total of $5$ of these hidden layers are used, all with the same embedding dimension. This diagram describes our architecture:
    \begin{align*}
    x^{(0)}(M) \xrightarrow[]{\varphi^{(0)}}
    x^{(1)}(U) \xleftrightarrow[]{\varphi^{(1)}}
    x^{(2)}(U) \xdashleftrightarrow[]{\varphi^{(L-1)}} 
    x^{(L-1)}(U) \xhookrightarrow[]{\varphi^{(L)}} 
    x^{(L)}(M).
    \end{align*}

\subsubsection*{Results}

\section{Conclusion}
Based on the theory of Lie groups and persistent homology a method for neural nets has been developed to parameterize them using the assumption that the data lies on or near by some connected commutative Lie group. This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold.

Applying Künneth's theorem, the homology groups of the topological factor spaces could be connected with the ones of their product space. Using persistence landscapes the elements originating from some homology groups on the filtration were estimated. With numerical experiments we predicted near ideal embedding dimensions and could confirm, according to Tab. \ref{figure3}, that such a neural embedding delivers a loss function on the validation data set with the smallest fluctuations and the best reconstruction results. We pose following open research questions:
\begin{itemize}
    \item We do not provide the neural layers with the explicit structure of an Abelian Lie group. How can the construction of spherical CNNs \cite{CohenGKW18} be used to always represent a product space $\mathbb{R}^p \times \mathbb{T}^q$ from layer to layer to generate exactly the porposed estimate of a manifold?
    \item Our approach can be applied to any kind of manifold, as far as its minimal representation is known. By applying decompositions, the procedure can be generalized and extended to obtain more accurate estimates. What decomposition can we use if we do not require the manifolds to be connected?
\end{itemize}

\bibliographystyle{splncs04}
\bibliography{biblio}
\end{document}

### 2020-06-17 llncs2e/anonymous.tex
anon and stuff

--- 2020-06-17 llncs2e/samplepaper.tex
@@
 \usepackage[hidelinks]{hyperref}
-\hypersetup{
-    pdftitle={Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory},    % title
-    pdfauthor={Luciano Melodia},     % author
-    pdfsubject={Estimate of the Neural Net Dimension},   % subject of the document
-    pdfcreator={Luciano Melodia},   % creator of the document
-    pdfproducer={Luciano Melodia},  % producer of the document
-    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {Neural nets} {Manifold learning}, % list of keywords
-    pdfnewwindow=true,
-    colorlinks,
-	citecolor=black,
-	filecolor=black,
-	linkcolor=black,
-	urlcolor=purple
-}
 \definecolor{red1}{RGB}{227,226,223}
@@
 %
-\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.}}
+\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{}{We will provide code and a link to the repository}.}}
 %
@@
 
-\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
-\and Richard Lenz\orcidID{0000-0003-1551-4824}}
+\author{Anonymous}
 %
-\authorrunning{L. Melodia et al.}
+\authorrunning{Anonymous}
 % First names are abbreviated in the running head.
@@
 %
-\institute{Chair of Computer Science 6\\
-Friedrich-Alexander University Erlangen-Nürnberg\\
-91058 Erlangen, Deutschland \\
-\email{\{luciano.melodia,richard.lenz\}@fau.de}}
 %
@@
 
-\subsubsection*{Results}
+\subsubsection*{Results} In Fig.
+\ref{figure2} c) d) we colored the neural nets according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.
+The embedding dimension can be taken from Tab.
+\ref{figure3}, due to the trivial invertibility of our neural nets this integer must be doubled, see \cite{DinhSB17}.
+All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig.
+\ref{figure2} c) d).
+The models above the treshold remain stable.
+A single net still shows an explosive gradient colored ${\color{red4}\bullet}$ in Fig.
+\ref{figure2} d).
+The net showing the deflection of gradient is very close to our threshold, identified as $292$ dimensional embedding.
+The closer we come to the treshold, the sparser the fluctuations of loss.
+We justified that the minimal embedding dimension does not lie below our treshold.
 
@@
 \begin{itemize}
-    \item We do not provide the neural layers with the explicit structure of an Abelian Lie group.
-How can the construction of spherical CNNs \cite{CohenGKW18} be used to always represent a product space $\mathbb{R}^p \times \mathbb{T}^q$ from layer to layer to generate exactly the porposed estimate of a manifold?
-    \item Our approach can be applied to any kind of manifold, as far as its minimal representation is known.
-By applying decompositions, the procedure can be generalized and extended to obtain more accurate estimates.
-What decomposition can we use if we do not require the manifolds to be connected?
+    \item The neural layers do not have the explicit structure of an Abelian Lie group.
+How can spherical CNNs \cite{CohenGKW18} be used to always represent a product space $\mathbb{R}^p \times \mathbb{T}^q$ from layer to layer in order to operate on the proposed Lie group?
+    \item This approach can be applied to any kind of manifold, as far as its minimal representation is known.
+Using decompositions, one may generalize this result.
+What decomposition allows to neglect the connectedness?
 \end{itemize}

### 2020-06-17 llncs2e/paper12.tex
anon and stuff

--- 2020-06-17 llncs2e/anonymous.tex
@@
 \usepackage[hidelinks]{hyperref}
+\hypersetup{
+    pdftitle={Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory},    % title
+    pdfauthor={Luciano Melodia},     % author
+    pdfsubject={Estimate of the Neural Net Dimension},   % subject of the document
+    pdfcreator={Luciano Melodia},   % creator of the document
+    pdfproducer={Luciano Melodia},  % producer of the document
+    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {Neural nets} {Manifold learning}, % list of keywords
+    pdfnewwindow=true,
+    colorlinks,
+	citecolor=black,
+	filecolor=black,
+	linkcolor=black,
+	urlcolor=purple
+}
 \definecolor{red1}{RGB}{227,226,223}
@@
 %
-\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{}{We will provide code and a link to the repository}.}}
+\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.}}
 %
@@
 
-\author{Anonymous}
+\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
+\and Richard Lenz\orcidID{0000-0003-1551-4824}}
 %
-\authorrunning{Anonymous}
+\authorrunning{L. Melodia et al.}
 % First names are abbreviated in the running head.
@@
 %
+\institute{Chair of Computer Science 6\\
+Friedrich-Alexander University Erlangen-Nürnberg\\
+91058 Erlangen, Deutschland \\
+\email{\{luciano.melodia,richard.lenz\}@fau.de}}
 %

### 2020-06-17 llncs2e/paper12.tex
corrections

@@
 \section{Related work}
-The related work is divided into two sections.
-In the first section we present chronologically the milestones in the consideration of neural nets from a differential geometric perspective.
-In the second section we present work on earlier attempts to determine their minimum embedding dimension.
+We present chronologically relevant work on neural nets from a differential geometric perspective.
+Further, we summarize earlier attempts to determine depth and width of a neural net, depending on the data to be processed.
 
-\subsubsection*{Riemannian geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean.
-Very recently, there have been results in the direction of other manifolds on which neural nets can operate.
-The spherical neural nets operate on a manifold with commutative group structure \cite{CohenGKW18,EstevesAMD20}.
+\subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean.
+Recently there have been results in the direction of other manifolds on which neural nets can operate.
+The spherical neural nets operate on a manifold with a commutative group structure \cite{CohenGKW18,EstevesAMD20}.
 
-A metric is required to describe the geometric properties of such a manifold.
-The change of coordinate systems and the learned metric tensor of a Riemann manifold during back propagation of neural nets has been investigated \cite{HauserR17}.
-Further, it has been shown that dense neural nets can't approximate all functions arbitrarily precisely \cite{Johnson19}.
-In response to this result, neural nets with residual connections have been examined.
-Residual nets add the output of one layer to a deeper one and therefore bypass some of the layers in between.
-This type of neural net is indeed an universal function approximator \cite{LinJ18}.
+A metric describes the geometric properties of a manifold.
+The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}.
+Further, it has been shown that dense neural nets cannot approximate all functions arbitrarily precisely \cite{Johnson19}.
+Thereupon neural nets with residual connections were investigated.
+Residual nets add the output of one layer to a deeper one, bypassing some of the layers in between.
+Indeed, they are universal function approximators \cite{LinJ18}.
+Inspired by finite difference methods, such a residual layer was defined as a forward or backward difference operator on a partition of layers.
+The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}.
+Such a layer is able to embed in $k\cdot n$ dimensional Euclidean space.
+The smooth manifold perspective inspired us to investigate its invariants leading into the theory discussed in this paper.
 
-\subsubsection*{Embedding dimension of neural nets} Cybenko showed that dense neural nets with a single hidden layer can approximate any continuous function with a bounded domain with arbitrary small error \cite{Cybenko92}.
-The upper bound of the embedding dimension for ReLU nets has recently been proven to be $(n+4)$ with $n$ being the sample dimension \cite{RaghuPKGS17}.
+\subsubsection*{Embedding dimension of neural nets} Cybenko showed that dense neural nets with one hidden layer can approximate any continuous function with a bounded domain with arbitrarily small error \cite{Cybenko92}.
+The upper bound of the embedding dimension for ReLU nets was recently determined to be $(n+4)$, where $n$ is the sample dimension \cite{RaghuPKGS17}.
 Bartlett et al.
-show that a nearly-tight bound for the VC dimension of a neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
+show an almost tight bound for VC dimension.
+A neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
+Its depth was determined as a parameter depending on the moduli of continuity of the function to be approximated \cite{LinJ18}.
 
-The depth of a neural net has been determined to be a parameter depending on the moduli of continuity of a function to be approximated \cite{LinJ18}.
-Inspired by the theory of finite difference methods, such a residual layer has been defined as a forward or backward difference operator on a partition of the layers.
-The state space dimension with residual connections is homeomorphic to $\mathbb{R}^{k \cdot n}$, where $n = \text{dim} \; M$ and $k$ is the number of times the difference operator was used \cite{HauserGJR19}.
-Thus, such a layer is capable of embedding into $k\cdot n$ dimensional space.
-The view that a neural net operates on a smooth manifold inspired us to investigate the invariants of this manifold for the theory discussed in this paper.
+To these fundamental results we contribute a topological approach to parameterization.
+Similar to Futagami et al.\cite{FutagamiYS19}, we investigate the manifold on which we suspect the data.
+Our approach is applicable to topological spaces in general and is based on persistent homology.
+We look for the dimension of a space with the same topological properties that the data indicates and which can be approximated.
+We assume a simple decomposition into a product space of realaxes and $1$-spheres to get the same homology groups as the persistent ones on a filtration of data.
+The dimension of the decomposition is a computable lower bound.
 
-To these fundamental results we contribute a topological approach to parameterization problems.
-Similar to Futagami et al.
-\cite{FutagamiYS19}, we study the manifold on which we suspect the data.
-Our approach is applicable to topological spaces in general and is based on persistent homology.
-We are looking for the dimension of a space with the same topological properties as data indicates and of which we know that it can be approximated.
-We use the simplest decomposition into a product space of realaxes and $1$-spheres to obtain the same homology groups as the filtration of the data has.
-Thus, we obtain a computable lower bound for the embedding dimension of any neural net with differentiable structure.
+\section{Smooth manifolds, Lie groups and persistent homology}
+The assumption that a set of points lies on a manifold is more accurate than the treatment in Euclidean space.
+A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that at any two points $x,y \in M$ have always open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty.
+This behaviour gives an intuitive feeling for the position of points in space.
+The topology $(M, \tau)$ is a set system that describes the structure of a geometric object.
+Here $M$ itself and $\emptyset$ must be contained in $\tau$, and any union of open sets from $\tau$ and any intersection of $\tau$ must be contained in $\tau$.
+If this set system has a subset which generates the topology $\tau$ with finitely many unions, then this subset is a \emph{basis} of the topology.
+We demand of a manifold that the basis is at most countable.
+In addition, $M$ is locally Euclidean, i.e.
+each point has a neighborhood which can be mapped homeomorphically to a subset of $\mathbb{R}^n$.
+The integer $n$ is the dimension of $M$.
 
-\newpage
-\section{Smooth manifolds, Lie groups and persistent homology}
-The assumption that a set of points lies on a manifold is more accurate than handling points in Euclidean space.
-A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that at any two points $x,y \in M$ there are always two open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty.
-This behaviour provides the intuitive feeling of the position of points in space.
-The topology $(M, \tau)$ is a set system that describes the structure of a geometric object.
-Thereby $M$ itself and $\emptyset$ must be contained in $\tau$, and any arbitrary union of open sets from $\tau$ and any intersections from $\tau$ must be in $\tau$.
-If this set system has a smaller set that generates the topology $\tau$ with the aid of finitely many unions, then this smaller set is a \emph{basis} of the topology.
-We demand of a manifold that the basis is at most countable.
-At last, $M$ is locally Euclidean, i.e.
-each point has a neighborhood which can be homeomorphically mapped to a subset of $\mathbb{R}^n$.
-The integer $n$ is the dimension of $M$.
-Further, we demand some additional smooth structure.
+\subsection{Smooth manifolds}
+A smooth structure is a stronger condition than a topological manifold which can be described by a family of continuous functions.
+This assumption is justified by the theorem of Stone-Weierstrass which proves that every continuous function can be approximated arbitrarily exactly by a smooth one \cite{stone1948generalized}.
+Smooth manifolds are described by a family of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to a subset of $\mathbb{R}^{n}$ and cover an open neighborhood $(U_i,\varphi_i)$ in $M$.
+This family of such charts is called \emph{atlas} on $M$, i.e.
+$\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}.
+The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$.
+If the atlas is maximal in terms of inclusion, then it is a \emph{differentiable structure}.
+Each atlas for $M$ is included in a maximal atlas.
+Due to the differentiable structure, the maps in $\mathscr{A}$ are also compatible with all maps of $\mathscr{A}_{\text{max}}$.
+It is sufficient to describe smooth manifolds by a non maximal atlas.
 
-\subsection{Smooth manifolds} A smooth structure is a stronger condition than a topological manifold which can be described by a family of continuous functions.
-This assumption can be justified by the theorem of Stone-Weierstrass which proves that any continuous function can be approximated arbitrarily precisely by a smooth one \cite{stone1948generalized}.
-Smooth manifolds are described by a collection of local coordinate charts $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$, homeomorphic to $\mathbb{R}^{n}$ and an open neighborhood $(U_i,\varphi_i)$ in $M$ which is covered by this map.
-The family of such maps that cover $M$ is the \textit{atlas} $\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}.
-The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$.
-If the atlas is maximal with respect to inclusion, then it is a \emph{differentiable structure}.
-Each atlas for $M$ is included in a maximal atlas.
-Because of the differentiable structure, the maps in $\mathscr{A}$ are also compatible with all other maps of $\mathscr{A}_{\text{max}}$.
-Thus, it is sufficient to describe smooth manifolds by an atlas that is not maximal.
-
-As long as the activation functions are smooth and of the same differentiable structure, they are also compatible and the manifold does not change while propagating the data trough the layers.
-Only the coordinate representation changes.
+As long as the activation functions are of the same differentiable structure, they also are compatible with each other, and the manifold does not change as the data propagates through the layers.
+Only the coordinate system changes.
 
@@
 We apply a theorem from Lie theory, which states that each connected Abelian Lie group $(M,\bullet)$ of dimension $\dim M = n$ is isomorphic to a product space $\mathbb{R}^p \times \mathbb{T}^q$ with $p+q =n$, for a proof we refer to \cite[p.~116]{onishchik1993lie}.
-The $q$-torus $\mathbb{T}^q$ is a surface of revolution, which means it moves a curve around an axis of rotation.
+The $q$-torus $\mathbb{T}^q$ is a surface of revolution.
+It moves a curve around an axis of rotation.
 These axes are given by $1$-spheres, such that the $q$-torus is a product space $\mathbb{T}^q = \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$.
@@
 Recall, that the $(n-1)$-sphere is given by $\mathcal{S}^{n-1} := \{x \in \mathbb{R}^n \; \vert \; \norm{x}_2 = 1\}$.
-Thus, $\mathcal{S}^1$ can be embedded into a two-dimensional Euclidean space.
-Under our assumption, we can derive how many dimensions are at least needed for a suitable embedding.
+Thus, $\mathcal{S}^1$ can be embedded into $\mathbb{R}^2$.
+Next, we derive how many dimensions are at least needed for a suitable embedding.
 For each $1$-sphere we count two dimensions and for each real line accordingly one.
-Finally, we have to estimate from data how many factor spaces are necessary to represent the topology of the underlying data manifold sufficiently accurately.
+Finally, we have to estimate from data how the decomposition may look like to have the topology of the data manifold.
 
 \subsection{Persistent homology}
-Algebraic topology provides a computable tool to study not the topology of a set of points itself but an Abelian group which is attached to the topology.
-The core interest in this discipline lies in \emph{homotopy equivalences} of an equivalence class to which objects belong that are continuously deformable into one another.
-This term is much broader than a homeomorphism.
-For two topological spaces $X$ and $Y$ we are looking for a function $h: X \times I \rightarrow Y$, which gives the identity at the time $h_0(X) = X$ and for $h_1(X) = X$ a mapping into the other topological space.
+Algebraic topology provides a computable tool to study not the topology of a set of points themselfs but an Abelian group attached to them.
+The core interest in this discipline lies in \emph{homotopy equivalences}, an equivalence class to which objects belong that are continuously deformable into one another.
+This is much broader than homeomorphism.
+For two topological spaces $X$ and $Y$ we seek for a function $h: X \times I \rightarrow Y$, which gives the identity at time $h_0(X) = X$ and for $h_1(X) = X$ a mapping into another topological space.
 If this mapping is continuous with respect to its arguments, it is called \emph{homotopy}.
-Let us consider two functions $f,g: I \rightarrow X$ so that $f(1) = g(0)$.
+Consider two functions $f,g: I \rightarrow X$ so that $f(1) = g(0)$.
 Then there is a \emph{composition} of product paths $f \cdot g$, which pass first through $f$ and then through $g$ and which is defined as $f \cdot g (s) = f(2s)$ for $0 \leq s \leq 1/2$ and $g(2s - 1)$, for $1/2 \leq s \leq 1$.
 We first run $f$ at double speed up to $1/2$ and from $1/2$ to $1$ we run the function $g$ at double speed.
-In addition, suppose that we have a family of functions $f: I \rightarrow X$ that have the same start and end point $f(0) = f(1) = x_0 \in X$.
-These functions intuitively form a \emph{loop}.
+In addition, suppose that a family of functions $f: I \rightarrow X$ is given which have the same start and end point $f(0) = f(1) = x_0 \in X$.
+They intuitively form a \emph{loop}.
 The set of all homotopy classes $[f]$ of loops $f: I \rightarrow X$ at base point $x_0$ is noted as $\pi_1(X,x_0)$ and is called first homotopy group or fundamental group.
-The group operation is the product $[f][g] = [f \cdot g]$ of equivalence classes.
+The group operation is the product of equivalence classes $[f][g] = [f \cdot g]$.
 If we do not look at the interval $I$, but at mappings considering the unit cube $I^n$, we obtain the $n$-th homotopy group $\pi_n(X,x_0)$ by analogy.
-With the help of the homotopy groups we can study the connected components of a topological space for the $0$th group, the loops for the $1$th group, the cavities for the $2$th and so forth.
-Since homotopy groups are very difficult to compute, we resort to an algebraic variant, the \emph{homology groups}.
+With the help of homotopy groups we study connected components of a topological space for the $0$th group, the loops for the $1$th group, the cavities for the $2$th and so forth.
+Since homotopy groups are difficult to compute, we resort to an algebraic variant, the \emph{homology groups}.
 
@@
 \subsubsection{Simplicial complexes} A \emph{simplicial complex} $K$ is a set of simplices, such that any face of a simplex of $K$ is a simplex of $K$ and the intersection of any two simplices of $K$ is either empty or a common face of both \cite[p.~11]{boissonnat2018geometric}.
-Further, we require every $(k+1)$-element, with $k < n$ to be a $k$-simplex of $[\sigma] \subset K$.
-This definition means that the simplicial complex is also a topological space.
-The properties described ensure that simplices are added to this set in a certain way, namely so that they are only \emph{glued} edge to edge, point to point, and surface to surface.
+We require every $(k+1)$-element, with $k < n$ to be a $k$-simplex of $[\sigma] \subset K$.
+This defines the simplicial complex as a topological space.
+These properties ensure that simplices are added to the complex in a way, such that they are \emph{glued} edge to edge, point to point, and surface to surface.
 Thus, every $n$-simplex has $n+1$ distinct vertices and no other $n$-simplex has the same set of vertices.
 This is a unique combinatorial description of vertices together with a collection of sets $\{[\sigma]_1, \cdots, [\sigma]_m\}$ of $k$-simplices, which are $k+1$ element subsets of $K$ \cite[p.~107]{hatcher2002algebraic}.
-This set system can be realized geometrically with a homeomorphism, by a corresponding mapping into an Euclidean space.
+This set system can be realized geometrically with a homeomorphism into Euclidean space.
 
@@
 The group is commutative or Abelian.
-The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges is given here by negative and positive signs, so that $[\sigma] = -[\tau]$, iff $[\sigma] = [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations.
+The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges, by negative and positive signs, such that $[\sigma] = -[\tau]$, iff $[\sigma] = [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations.
 The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$.
@@
 \section{Neural nets}
-\label{sec:neuralnetworks}
+\label{sec:neuralnets}
 Neural nets are a composition of \emph{affine transformations} with a non-linear activation function.
@@
 It also preserves parallelism and partial relationships.
-The projection onto the $(l+1)$-th layer of such a network can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural net.
+The projection onto the $(l+1)$-th layer of such a net can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural net.
 
@@
 \cite{DinhSB17}.
-We use invertible architectures to avoid degenerated cases and to stay on the same differentiable structure during learning within the layers to be tested.
-A total of $5$ of these hidden layers are used, all with the same embedding dimension.
-This diagram describes our architecture:
-    \begin{align*}
-    x^{(0)}(M) \xrightarrow[]{\varphi^{(0)}}
-    x^{(1)}(U) \xleftrightarrow[]{\varphi^{(1)}}
-    x^{(2)}(U) \xdashleftrightarrow[]{\varphi^{(L-1)}} 
-    x^{(L-1)}(U) \xhookrightarrow[]{\varphi^{(L)}} 
-    x^{(L)}(M).
-    \end{align*}
+We use invertible architectures to stay on the same differentiable structure during learning.
+A total of $5$ hidden layers are used, all with the same embedding dimension.
 

### 2020-06-17 llncs2e/paper12.tex
paper12$

@@
 Recently there have been results in the direction of other manifolds on which neural nets can operate.
-The spherical neural nets operate on a manifold with a commutative group structure \cite{CohenGKW18,EstevesAMD20}.
+The spherical neural nets operate on a manifold with a commutative group structure \cite{CohenGKW18}.
 
@@
 
-\subsubsection{Persistence landscapes} Persistence landscapes give a statistical summary of the topology of a set of points embedded in a given topological manifold \cite{bubenik2015statistical,bubenikD17}.
+\subsubsection{Persistence landscapes} Persistence landscapes give a statistical summary of the topology of a set of points embedded in a given topological manifold \cite{bubenik2015statistical}.
 Looking at the points $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2}$ on the $k$-th persistence diagram $\text{PH}_{k}(X)$, one associates a piecewise linear function $\lambda^{\epsilon_i}_{\epsilon_{j+1}}: \mathbb{R} \rightarrow [0,\infty)$ with those points:
@@
 \section{Counting Betti numbers}
-Based on our assumption that data lies on a connected Abelian Lie group, we can decompose the data space into a product of realaxes and tori.
-In the following we refer to this Lie group as $G \simeq \mathbb{R}^p \times \mathbb{T}^q$.
-By this decomposition there is also an isomorphism of the homology groups of $G$ with those of $\mathbb{R}^p \times \mathbb{T}^q$, so that $H_k(G) \simeq H_k(\mathbb{R}^p \times \mathcal{S}^{1}_{1}\times\cdots\times\mathcal{S}^{1}_{q})$ with $p+q = \dim G$.
+According to our main assumption, we seek a decomposition of the data manifold into a product of realaxes and tori.
+We refer to this Lie group as $G \simeq \mathbb{R}^p \times \mathbb{T}^q$.
+Thus, there is also an isomorphism of the homology groups of $G$ with those of $\mathbb{R}^p \times \mathbb{T}^q$, so that $H_k(G) \simeq H_k(\mathbb{R}^p \times \mathcal{S}^{1}_{1}\times\cdots\times\mathcal{S}^{1}_{q})$ with $p+q = \dim G$.
 We demand from a neural net that these homology groups can all be approximated, i.e.
 that all topological structure of the input space can be represented sufficiently well.
-
-One might wonder to what extent the homology groups of a simplicial complex can be used to estimate the homology groups of the manifolds underlying the data.
-In this section we denote the simplicial homology groups as $H^{\Delta}_{\bullet}(X)$.
-The $\bullet$ is a proxy for all possible indices.
-Through the construction of continuous mappings, so-called singular simplices, from a simplex into a topological space, the singular homology groups $H_{\bullet}(X)$ can be constructed.
-We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a detailed introduction with proofs.
-This allows to assign Abelian groups to any topological space.
-If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{\bullet}(X)$, see for a proof \cite[pp.~473]{lee2013smooth}.
-Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{\bullet}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}.
-In summary, we have homology theories for simplicial complexes, topological spaces, a smooth variant on topological spaces and on smooth manifolds.
-We consider simplicial homology to be determined on the basis of a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space.
-However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}.
-This legitimizes our approach.
 
@@
 
+One might wonder to what extent homology groups of a simplicial complex can be used to estimate homology groups of smooth manifolds.
+Through construction of continuous maps from a simplex into a topological space, so-called singular simplices, the singular homology groups $H_{k}(X)$ are obtained.
+We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a proof.
+This allows to assign Abelian groups to any topological space.
+If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{k}(X)$, consider \cite[pp.~473]{lee2013smooth} for a proof.
+Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{k}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}.
+In summary, we have homology theories for simplicial complexes, topological spaces and on smooth manifolds.
+Simplicial homology is determined by a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space.
+However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}.
+This legitimizes our approach.
+
 \section{Experimental setting}
+\input{figure2.tex}
 We train an auto-encoder using as input heavily noisy images and map them to their noiseless original.
 We perturb each input vector $\textbf{x}^{(0)}$ with Gaussian noise $\epsilon \sim \mathcal{N}(0,\sigma^2 \mathbf{I})$, such that the input is weighted $0.5 \cdot \textbf{x}^{(0)} + 0.5 \cdot \epsilon$.
-The embedding dimension is increased for each experiment by $2$, i.e.
+The layer size is increased for each experiment by $2$, i.e.
 $2,4,6, \cdots, 784$.
@@
 Each line of Fig.
-\ref{figure2} c), d) represents one neural architecture as averaged loss function.
-\input{figure2.tex}
-\subsubsection*{Persistence landscapes} We use the Delaunay-complex for the filtration, according to Melodia et al.
+\ref{figure2} c), d) represents one neural architecture as averaged loss function.\\[0.4cm]
+\textbf{Persistence Landscapes:} We use the Delaunay-complex for the filtration, according to Melodia et al.
 \cite{lume}.
@@
 We implement the persistence diagrams and persistence landscapes using the \texttt{GUDHI} library \cite{gudhi:urm}.
-Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.
-\subsubsection*{Hyperparameters} The batch size is set to $128$, which should help to better recognize good parametrization.
-Higher batch size causes explosions of the gradient, which we visualize in Fig.
-\ref{figure2} c), d).
-For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$.
-The dataset is randomly shuffled.
-We implement in \texttt{Python} $3.7$.
-We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}.
-The neural training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.
-\input{figure3.tex}
-\subsubsection*{Neural net} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function.
+Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.\\[0.4cm]
+\textbf{Neural nets:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function.
 The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold.
@@
 A total of $5$ hidden layers are used, all with the same embedding dimension.
-
-\subsubsection*{Results} In Fig.
+The batch size is set to $128$, which should help to better recognize good parametrization.
+Higher batch size causes explosions of the gradient, which we visualize in Fig.
+\ref{figure2} c), d).
+For optimization we use \texttt{ADAM} \cite{KingmaB14} and a learning rate of $10^{-4}$.
+The dataset is randomly shuffled.
+We implement in \texttt{Python} $3.7$.
+We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}.
+The training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.\\[0.4cm]
+\textbf{Results:} In Fig.
 \ref{figure2} c) d) we colored the neural nets according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.
 The embedding dimension can be taken from Tab.
-\ref{figure3}, due to the trivial invertibility of our neural nets this integer must be doubled, see \cite{DinhSB17}.
-All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig.
+\ref{figure3}.\footnote{Due to the construction of invertibility the treshold must be doubled, see \cite{DinhSB17}.} All models below our dimensional treshold $\dim U$ show drastic explosions of the gradient, see Fig.
 \ref{figure2} c) d).
@@
 The closer we come to the treshold, the sparser the fluctuations of loss.
-We justified that the minimal embedding dimension does not lie below our treshold.
+\input{figure3.tex}
 
@@
 Based on the theory of Lie groups and persistent homology a method for neural nets has been developed to parameterize them using the assumption that the data lies on or near by some connected commutative Lie group.
-This forms an approximate solution for a special case of learning problems, allowing to restrict the domain to a commutative group beeing a connected smooth manifold.
+This forms an approximate solution for a special case of learning problems.
 
@@
 Using persistence landscapes the elements originating from some homology groups on the filtration were estimated.
-With numerical experiments we predicted near ideal embedding dimensions and could confirm, according to Tab.
-\ref{figure3}, that such a neural embedding delivers a loss function on the validation data set with the smallest fluctuations and the best reconstruction results.
+With numerical experiments we predicted near ideal embedding dimensions and could confirm, see Fig.
+\ref{figure2}, that a neural embedding above the treshold delivers a loss function on the validation data set with small fluctuations and best reconstruction results.
 We pose following open research questions:

### 2020-06-19 llncs2e/paper12.tex
data

@@
 %
-\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/Irhe}{https://github.com/karhunenloeve/Irhe}.}}
+\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/NNTopoLie}{https://github.com/karhunenloeve/NNTopoLie}.}}
 %

### 2020-06-19 llncs2e/paper12.tex
finish

@@
 %
-\titlerunning{Estimate of the Neural net Dimension}
+\titlerunning{Estimate of the Neural Net Dimension}
 
@@
 Bartlett et al.
-show an almost tight bound for VC dimension.
+show an almost tight bound for VC dimensions.
 A neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
@@
 The assumption that a set of points lies on a manifold is more accurate than the treatment in Euclidean space.
-A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that at any two points $x,y \in M$ have always open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty.
+A \emph{topological manifold} $M$ is a \emph{Hausdorff} space, which means that any two points $x,y \in M$ have always open neighborhoods $U_x,U_y \subset M$, so that their intersection is empty.
 This behaviour gives an intuitive feeling for the position of points in space.
@@
 Smooth manifolds are described by a family of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to a subset of $\mathbb{R}^{n}$ and cover an open neighborhood $(U_i,\varphi_i)$ in $M$.
-This family of such charts is called \emph{atlas} on $M$, i.e.
+A family of charts, which covers $M$, is called \emph{atlas} on $M$, i.e.
 $\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}.
@@
 For each $1$-sphere we count two dimensions and for each real line accordingly one.
-Finally, we have to estimate from data how the decomposition may look like to have the topology of the data manifold.
+Finally, we have to estimate from data how the decomposition may look like to yield the topology of the data manifold.
 
@@
 This is much broader than homeomorphism.
-For two topological spaces $X$ and $Y$ we seek for a function $h: X \times I \rightarrow Y$, which gives the identity at time $h_0(X) = X$ and for $h_1(X) = X$ a mapping into another topological space.
+For two topological spaces $X$ and $Y$ we seek for a function $h: X \times I \rightarrow Y$, which gives the identity at time $h_0(X) = X$ and for $h_1(X) = Y$ a mapping into another topological space.
 If this mapping is continuous with respect to its arguments, it is called \emph{homotopy}.
@@
 This is a unique combinatorial description of vertices together with a collection of sets $\{[\sigma]_1, \cdots, [\sigma]_m\}$ of $k$-simplices, which are $k+1$ element subsets of $K$ \cite[p.~107]{hatcher2002algebraic}.
-This set system can be realized geometrically with a homeomorphism into Euclidean space.
+This set system can be realized geometrically with an injective map into Euclidean space.
 
@@
 The group is commutative or Abelian.
-The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges, by negative and positive signs, such that $[\sigma] = -[\tau]$, iff $[\sigma] = [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations.
+The used coefficients $\lambda_i$ induce an orientation, intuitively a direction of the edges, by negative and positive signs, such that $[\sigma] = -[\tau]$, iff $[\sigma] \sim [\tau]$ and $[\sigma]$ and $[\tau]$ have different orientations.
 The objects of the chain group are called $k$-chains, for $[\sigma] \in C_k$.
@@
     \emptyset &= K^{\epsilon_0} \subset K^{\epsilon_1} \subset \cdots \subset K^{\epsilon_{n+1}}= K,\\
-    K^{\epsilon_{i+1}} &= K^{\epsilon_i} \cup \sigma^{\epsilon_{i+1}}, \quad \text{for} \; i \in \{0, \cdots, n-1\}.
+    K^{\epsilon_{i+1}} &= K^{\epsilon_i} \cup [\sigma]^{\epsilon_{i+1}}, \quad \text{for} \; i \in \{0, \cdots, n-1\}.
 \end{align}
@@
 \begin{equation}
-\mu^{\epsilon_{i},\epsilon_{j}}_k = (\beta_k^{\epsilon_{i},\epsilon_{j-1}} - \beta_k^{\epsilon_i,\epsilon_{j}})-(\beta_k^{\epsilon_{i-1},\epsilon_{j-1}}-\beta_k^{\epsilon_{i-1},\epsilon_{j}})
+\mu^{\epsilon_{i},\epsilon_{j}}_k = (\beta_k^{\epsilon_{i},\epsilon_{j-1}} - \beta_k^{\epsilon_i,\epsilon_{j}})-(\beta_k^{\epsilon_{i-1},\epsilon_{j-1}}-\beta_k^{\epsilon_{i-1},\epsilon_{j}}),
 \end{equation}
@@
 
-Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space $X \times Y$ is isomorphic to the direct sum of the tensor product of its $k$th homology groups, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
+Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space is isomorphic to the direct sum of the tensor product of $k$th homology groups from its factors, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
 \begin{align}

### 2020-07-05 llncs2e/paper12.tex
small correction

@@
 \end{align}
-By the isomorphism of simplicial, singular, smooth singular and De Rham cohomology, we use the theorem in accordance to our assumption.
-Thus, we count the representatives of homology classes in the persistence landscapes to derive the dimension of the sought manifold.
-The homology groups of the $1$-sphere are
+One might wonder to what extent homology groups of a simplicial complex can be used to estimate homology groups of smooth manifolds.
+Through construction of continuous maps from a simplex into a topological space, so-called singular simplices, the singular homology groups $H_{k}(X)$ are obtained.
+We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a proof.
+This allows to assign Abelian groups to any topological space.
+If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{k}(X)$, consider \cite[pp.~473]{lee2013smooth} for a proof.
+Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{k}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}.
+In summary, we have homology theories for simplicial complexes, topological spaces and on smooth manifolds.
+Simplicial homology is determined by a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space.
+However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}.
+By the isomorphism of simplicial, singular, smooth singular and De Rham cohomology, our approach is legitimized.
+
+We count the representatives of homology classes in the persistence landscapes to derive the dimension of the sought manifold. We get for $\mathcal{S}^1$:
 \begin{align}
@@
 \ref{figure3}.
-
-One might wonder to what extent homology groups of a simplicial complex can be used to estimate homology groups of smooth manifolds.
-Through construction of continuous maps from a simplex into a topological space, so-called singular simplices, the singular homology groups $H_{k}(X)$ are obtained.
-We refer to Hatcher \cite[p.~102]{hatcher2002algebraic} for a proof.
-This allows to assign Abelian groups to any topological space.
-If one constructs smooth mappings instead of continuous ones, one gets smooth singular homology groups $H^{\infty}_{k}(X)$, consider \cite[pp.~473]{lee2013smooth} for a proof.
-Finally a chain complex can be formed on a smooth manifold over the $p$-forms on $X$ and from this the so-called De Rham cohomology can be defined $H_{\text{dR}}^{k}(X)$, for a proof we refer to \cite[pp.~440]{lee2013smooth}.
-In summary, we have homology theories for simplicial complexes, topological spaces and on smooth manifolds.
-Simplicial homology is determined by a simplicial complex, which is merely a rough approximation of the triangulation of the underlying topological space.
-However, we can draw conclusions about a possible smooth manifold, since the discussed homology groups all have isomorphisms to each other known as the De Rham Theorem, see \cite[pp.~106, pp.~467]{hatcher2002algebraic,lee2013smooth}.
-This legitimizes our approach.
 

### 2020-07-07 llncs2e/paper12.tex
paper

@@
 %
-\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/NNTopoLie}{https://github.com/karhunenloeve/NNTopoLie}.}}
+\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/ neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
 %

### 2020-07-14 llncs2e/paper12.tex
corrections

@@
 \begin{abstract}
-In this paper we present an approach to determine the smallest possible number of perceptrons in a neural net in such a way that the topology of the input space can be learned sufficiently well.
+In this paper we present an approach to determine the smallest possible number of neurons in a neural net in such a way that the topology of the input space can be learned sufficiently well.
 We introduce a general procedure based on persistent homology to investigate topological invariants of the manifold on which we suspect the data set.
@@
 \section{Introduction}
-Since the development of deep neural nets, their parameterization, in particular the smallest possible number of perceptual units in a layer, has been studied.
+Since the development of deep neural nets, their parameterization, in particular the smallest possible number of neurons in a layer, has been studied.
 This number is of importance for auto-encoding tasks that require extrapolation or interpolation of data, such as blind source separation or super-resolution.
@@
 \section{Related work}
-We present chronologically relevant work on neural nets from a differential geometric perspective.
+We present relevant work on neural nets from a differential geometric perspective.
 Further, we summarize earlier attempts to determine depth and width of a neural net, depending on the data to be processed.
@@
 \subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean.
-Recently there have been results in the direction of other manifolds on which neural nets can operate.
-The spherical neural nets operate on a manifold with a commutative group structure \cite{CohenGKW18}.
+Recently there have been results in the direction of other manifolds on which neural nets can operate, such as the spherical neural nets \cite{CohenGKW18}.
+They operate on a manifold with a commutative group structure.
 
@@
 show an almost tight bound for VC dimensions.
-A neural net with $W$ weights and $L$ layers has $O(WL \log W)$ but $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
+A neural net with $W$ weights and $L$ layers is bounded by $O(WL \log W)$ VC dimensions but has $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
 Its depth was determined as a parameter depending on the moduli of continuity of the function to be approximated \cite{LinJ18}.
 
-To these fundamental results we contribute a topological approach to parameterization.
+To these fundamental results we contribute a topological approach to the parameterization of the minimal amount of neurons within a layer, such that the topology of the data manifold can be represented.
 Similar to Futagami et al.\cite{FutagamiYS19}, we investigate the manifold on which we suspect the data.
@@
 The topology $(M, \tau)$ is a set system that describes the structure of a geometric object.
-Here $M$ itself and $\emptyset$ must be contained in $\tau$, and any union of open sets from $\tau$ and any intersection of $\tau$ must be contained in $\tau$.
+Here $M$ itself and the empty set $\emptyset$ must be contained in $\tau$, and any union of open sets from $\tau$ and any intersection of $\tau$ must be contained in $\tau$.
 If this set system has a subset which generates the topology $\tau$ with finitely many unions, then this subset is a \emph{basis} of the topology.
@@
 \subsubsection{Lie groups} The study of group theory deals with symmetries which can be expressed algebraically.
-A pair $(M,\bullet)$, consisting of a map $\bullet: M\times M \rightarrow M$ is called group if the map is associative, $x \bullet (y \bullet z) = (x \bullet y) \bullet z$, has a neutral element so that $x \bullet e_{M} = x$ and has an inverse element to each element of the group, $x \bullet x^{-1} = e_{M}$.
+A pair $(M,\bullet)$, consisting of a map $\bullet: M\times M \rightarrow M$ is called group if the map is associative, i.e.
+$x \bullet (y \bullet z) = (x \bullet y) \bullet z$ and has a neutral element so that $x \bullet e_{M} = x$.
+We also require an inverse element to each element of the group, i.e.
+$x \bullet x^{-1} = e_{M}$.
 A Lie group is a smooth manifold which is equipped with a group structure such that the maps $\bullet:M \times  M \rightarrow  M, (x,y) \mapsto xy, $ and $\imath: M \rightarrow  M, x \mapsto x^{-1}$ are smooth.
@@
 
-In practice neural nets are used with a different number of perceptrons per layer.
+In practice neural nets are used with a different number of neurons per layer.
 However, a layer can represent the manifold of the previous one.

### 2020-07-14 llncs2e/paper12.tex
corrections

@@
 \begin{abstract}
-In this paper we present an approach to determine the smallest possible number of neurons in a neural net in such a way that the topology of the input space can be learned sufficiently well.
+In this paper, we present an approach to determine the smallest possible number of neurons in a neural net in such a way that the topology of the input space can be learned sufficiently well.
 We introduce a general procedure based on persistent homology to investigate topological invariants of the manifold on which we suspect the data set.
@@
 \subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean.
-Recently there have been results in the direction of other manifolds on which neural nets can operate, such as the spherical neural nets \cite{CohenGKW18}.
-They operate on a manifold with a commutative group structure.
+Nevertheless, it does often approximate a manifold with different structure.
+Recently there have been results in the direction of other manifolds on which neural nets can operate and that fit more the nature of data, such as the spherical neural nets \cite{CohenGKW18}.
+They operate on a model of a manifold with commutative group structure.
+Manifolds induce a notion of distance, with symmetric properties, also referred to as metric.
 
 A metric describes the geometric properties of a manifold.
-The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}.
+The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie Group.
 Further, it has been shown that dense neural nets cannot approximate all functions arbitrarily precisely \cite{Johnson19}.
@@
 completely covered with simplices.
-For a proof and illustrations we refer to \cite[p.~102]{hatcher2002algebraic,lume}.
+For a proof and illustrations we refer to \cite[p.~102]{hatcher2002algebraic}\cite{lume}.
 The concept of the triangle is too specific, so the \emph{$n$-simplices} are defined as a generalization.
@@
 \ref{simplex}.
-Together with addition as an operation, they form a group structure called $k$th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used.
+Together with addition as an operation, they form a group structure called $k$-th chain group $\langle C_k, +\rangle$ for the corresponding dimension of the $k$-simplices used.
 The group is commutative or Abelian.
@@
 
-The $k$th chain groups contain two more Abelian subgroups that behave normal to it.
+The $k$-th chain groups contain two more Abelian subgroups that behave normal to it.
 First, the so-called cycle groups $\langle Z_k, +\rangle$, which are defined as $Z_k := \ker \partial_k = \{[\sigma] \in C_k \; \vert \; \partial_k [\sigma] = \emptyset \}$.
@@
 
-In practice neural nets are used with a different number of neurons per layer.
+In practice, neural nets are used with a different number of neurons per layer.
 However, a layer can represent the manifold of the previous one.
@@
 
-Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space is isomorphic to the direct sum of the tensor product of $k$th homology groups from its factors, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
+Using Künneth's theorem, we know that the $k$-th singular homology group of a topological product space is isomorphic to the direct sum of the tensor product of $k$-th homology groups from its factors, for a proof we refer to \cite[p.~268]{hatcher2002algebraic}:
 \begin{align}
@@
 $2,4,6, \cdots, 784$.
-$392$ neural nets have been trained $\times 100$.
+$392$ neural nets around a hundred times.
 Each line of Fig.
@@
 Dense neural nets are used with bias term.
-The datasets are flattened from $(32,32,3)$ into $(3072)$.
+The datasets are flattened from $(32,32,3)$ into $(3072)$ for the purpose of fitting into a dense layer.
 The first layer is applied to the flattened dataset, which is also adapted by back propagation, i.e.
@@
 We implement in \texttt{Python} $3.7$.
-We use \texttt{Tensorflow} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}.
+We use \texttt{Tensorflow v.2.2.0} as backend \cite{tensorflow2015-whitepaper} with \texttt{Keras} as wrapper \cite{chollet2015keras}.
 The training is conducted on \texttt{Intel Core i}$7$ $9700$\texttt{K} processors.\\[0.4cm]

### 2020-07-14 llncs2e/paper12.tex
corrections

@@
 What decomposition allows to neglect the connectedness?
+    \item What do other representations of manifolds with a more complicated structure look like?
+For example, can manifolds like the Klein bottle be learned that are not orientable?
+And if so, how can one decide in advance whether or not the approximation to an orientable manifolds makes more sense than to a non orientable one?
 \end{itemize}

### 2020-07-14 llncs2e/paper12.tex
corrections

@@
 \hypersetup{
-    pdftitle={Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory},    % title
+    pdftitle={Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory},    % title
     pdfauthor={Luciano Melodia},     % author
-    pdfsubject={Estimate of the Neural Net Dimension},   % subject of the document
+    pdfsubject={Estimate of the Neural Network Dimension},   % subject of the document
     pdfcreator={Luciano Melodia},   % creator of the document
     pdfproducer={Luciano Melodia},  % producer of the document
-    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {Neural nets} {Manifold learning}, % list of keywords
+    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {neural networks} {Manifold learning}, % list of keywords
     pdfnewwindow=true,
@@
 %
-\title{Estimate of the Neural Net Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/ neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
+\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/ neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
 %
@@
 %
-\titlerunning{Estimate of the Neural Net Dimension}
+\titlerunning{Estimate of the neural network Dimension}
 
@@
 \begin{abstract}
-In this paper, we present an approach to determine the smallest possible number of neurons in a neural net in such a way that the topology of the input space can be learned sufficiently well.
+In this paper, we present an approach to determine the smallest possible number of neurons in a layer of a neural network in such a way that the topology of the input space can be learned sufficiently well.
 We introduce a general procedure based on persistent homology to investigate topological invariants of the manifold on which we suspect the data set.
@@
 
-\keywords{Embedding dimension \and Parameterization \and Persistent homology \and Neural nets \and Manifold learning.}
+\keywords{Embedding dimension \and Parameterization \and Persistent homology \and Neural Networks \and Manifold learning.}
 \end{abstract}
@@
 \section{Introduction}
-Since the development of deep neural nets, their parameterization, in particular the smallest possible number of neurons in a layer, has been studied.
+Since the development of deep neural networks, their parameterization, in particular the smallest possible number of neurons in a layer, has been studied.
 This number is of importance for auto-encoding tasks that require extrapolation or interpolation of data, such as blind source separation or super-resolution.
@@
     \item We study topological invariants using statistical summaries of persistent homology on a filtered simplicial complex on the data.
-    \item Using the theory of Lie groups, we specify the smallest possible embedding dimension so that the neural net is able to estimate a projection onto a space with the determined invariants.
+    \item Using the theory of Lie groups, we specify the smallest possible embedding dimension so that the neural network is able to estimate a projection onto a space with the determined invariants.
 \end{itemize}
@@
 \section{Related work}
-We present relevant work on neural nets from a differential geometric perspective.
-Further, we summarize earlier attempts to determine depth and width of a neural net, depending on the data to be processed.
+We present relevant work on neural networks from a differential geometric perspective.
+Further, we summarize earlier attempts to determine depth and width of a neural network, depending on the data to be processed.
 
-\subsubsection*{Differential geometry in neural nets} The manifold of a dense neural net is in most cases Euclidean.
+\subsubsection*{Differential geometry in neural networks} The manifold of a dense neural network is in most cases Euclidean.
 Nevertheless, it does often approximate a manifold with different structure.
-Recently there have been results in the direction of other manifolds on which neural nets can operate and that fit more the nature of data, such as the spherical neural nets \cite{CohenGKW18}.
+Recently there have been results in the direction of other manifolds on which neural networks can operate and that fit more the nature of data, such as the spherical neural networks \cite{CohenGKW18}.
 They operate on a model of a manifold with commutative group structure.
@@
 The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie Group.
-Further, it has been shown that dense neural nets cannot approximate all functions arbitrarily precisely \cite{Johnson19}.
-Thereupon neural nets with residual connections were investigated.
+Further, it has been shown that dense neural networks cannot approximate all functions arbitrarily precisely \cite{Johnson19}.
+Thereupon neural networks with residual connections were investigated.
 Residual nets add the output of one layer to a deeper one, bypassing some of the layers in between.
@@
 
-\subsubsection*{Embedding dimension of neural nets} Cybenko showed that dense neural nets with one hidden layer can approximate any continuous function with a bounded domain with arbitrarily small error \cite{Cybenko92}.
+\subsubsection*{Embedding dimension of neural networks} Cybenko showed that dense neural networks with one hidden layer can approximate any continuous function with a bounded domain with arbitrarily small error \cite{Cybenko92}.
 The upper bound of the embedding dimension for ReLU nets was recently determined to be $(n+4)$, where $n$ is the sample dimension \cite{RaghuPKGS17}.
@@
 show an almost tight bound for VC dimensions.
-A neural net with $W$ weights and $L$ layers is bounded by $O(WL \log W)$ VC dimensions but has $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
+A neural network with $W$ weights and $L$ layers is bounded by $O(WL \log W)$ VC dimensions but has $\Omega(WL \log W/L)$ VC dimensions \cite{BartlettHLM19}.
 Its depth was determined as a parameter depending on the moduli of continuity of the function to be approximated \cite{LinJ18}.
@@
 
-\section{Neural nets}
+\section{Neural Networks}
 \label{sec:neuralnets}
-Neural nets are a composition of \emph{affine transformations} with a non-linear activation function.
+Neural networks are a composition of \emph{affine transformations} with a non-linear activation function.
 This transformation obtains collinearity.
 It also preserves parallelism and partial relationships.
-The projection onto the $(l+1)$-th layer of such a net can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural net.
+The projection onto the $(l+1)$-th layer of such a net can be formalized as $\textbf{x}^{(l+1)} = f(\mathbf{W}^{(l+1)} \cdot \textbf{x}^{(l)}+\textbf{b}^{(l+1)}).$ The composition of multiple such maps is a deep neural network.
 
@@
 Commonly used are functions from the exponential family.
-As a result a neural net is a map $\varphi^{(l)}:\textbf{x}^{(l)}(M)$ $\rightarrow$ $(\varphi^{(l)} \circ \textbf{x}^{(l)})(M)$.
+As a result a neural network is a map $\varphi^{(l)}:\textbf{x}^{(l)}(M)$ $\rightarrow$ $(\varphi^{(l)} \circ \textbf{x}^{(l)})(M)$.
 Learning by back propagation can in this way be considered as a change of coordinate charts of a smooth manifold.
@@
 
-In practice, neural nets are used with a different number of neurons per layer.
+In practice, neural networks are used with a different number of neurons per layer.
 However, a layer can represent the manifold of the previous one.
 Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map does not change.
-Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural net \cite{HauserR17}.
+Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural network \cite{HauserR17}.
 
@@
 Thus, there is also an isomorphism of the homology groups of $G$ with those of $\mathbb{R}^p \times \mathbb{T}^q$, so that $H_k(G) \simeq H_k(\mathbb{R}^p \times \mathcal{S}^{1}_{1}\times\cdots\times\mathcal{S}^{1}_{q})$ with $p+q = \dim G$.
-We demand from a neural net that these homology groups can all be approximated, i.e.
-that all topological structure of the input space can be represented sufficiently well.
+We demand from a neural network that these homology groups can all be approximated, i.e.
+that all topological structure of the input space can be represented sufficiently well, by means of its topological invariants.
 
@@
 $2,4,6, \cdots, 784$.
-$392$ neural nets around a hundred times.
+$392$ neural networks around a hundred times.
 Each line of Fig.
@@
 Persistent homology is computed on a \texttt{NVIDIA Quadro P4000}.\\[0.4cm]
-\textbf{Neural nets:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function.
+\textbf{Neural networks:} We use $\text{L}(\textbf{x}^{(0)}, \textbf{y}) = 1/n \sum_{i=1}^{n} (\textbf{x}^{(0)}_i-\textbf{y}_i)^2$ as loss function.
 The sigmoid activation function $\sigma(\mathbf{x}^{(l)}) = 1 / 1 + e^{-\mathbf{x}^{(l)}}$ is applied throughout the net to yield the structure of a smooth manifold.
-Dense neural nets are used with bias term.
+Dense neural networks are used with bias term.
 The datasets are flattened from $(32,32,3)$ into $(3072)$ for the purpose of fitting into a dense layer.
@@
 \textbf{Results:} In Fig.
-\ref{figure2} c) d) we colored the neural nets according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.
+\ref{figure2} c) d) we colored the neural networks according to the chosen embedding, such that the lines show the following dimensions ${\color{red1}\bullet} = [2, \cdots, 148]$, ${\color{red2}\bullet} = [150, \cdots, 198]$, ${\color{red3}\bullet} = [200, \cdots, 292]$ and ${\color{red4}\bullet} = [\dim U \approx 294, \cdots, 784]$.
 The embedding dimension can be taken from Tab.
@@
 \section{Conclusion}
-Based on the theory of Lie groups and persistent homology a method for neural nets has been developed to parameterize them using the assumption that the data lies on or near by some connected commutative Lie group.
+Based on the theory of Lie groups and persistent homology a method for neural networks has been developed to parameterize them using the assumption that the data lies on or near by some connected commutative Lie group.
 This forms an approximate solution for a special case of learning problems.

### 2020-07-14 llncs2e/paper12.tex
final

@@
 Further, it has been shown that dense neural networks cannot approximate all functions arbitrarily precisely \cite{Johnson19}.
-Thereupon neural networks with residual connections were investigated.
-Residual nets add the output of one layer to a deeper one, bypassing some of the layers in between.
+Thereupon, neural networks with residual connections were investigated.
+Residual networks add the output of one layer to a deeper one, bypassing some of the layers in between.
 Indeed, they are universal function approximators \cite{LinJ18}.
@@
 \subsubsection*{Embedding dimension of neural networks} Cybenko showed that dense neural networks with one hidden layer can approximate any continuous function with a bounded domain with arbitrarily small error \cite{Cybenko92}.
-The upper bound of the embedding dimension for ReLU nets was recently determined to be $(n+4)$, where $n$ is the sample dimension \cite{RaghuPKGS17}.
+The upper bound of the embedding dimension for \texttt{ReLU} networks was recently determined to be $(n+4)$, where $n$ is the sample dimension \cite{RaghuPKGS17}.
 Bartlett et al.
@@
 Smooth manifolds are described by a family of local coordinate maps $\varphi: U \rightarrow \varphi(U) \subseteq \mathbb{R}^n$ which are homeomorphic to a subset of $\mathbb{R}^{n}$ and cover an open neighborhood $(U_i,\varphi_i)$ in $M$.
-A family of charts, which covers $M$, is called \emph{atlas} on $M$, i.e.
-$\mathscr{A} = (U_i,\varphi_i)_{i \in I}$ \cite{lee2013smooth}.
+A family of charts, which covers $M$, is called \emph{atlas} on $M$ \cite{lee2013smooth}, i.e.
+$\mathscr{A} = (U_i,\varphi_i)_{i \in I}$.
 The functions $x_1, \cdots, x_n: U \rightarrow \mathbb{R}$ are called local coordinates $\varphi(p) = (x_1(p), \cdots, x_n(p))$.
@@
 Due to the differentiable structure, the maps in $\mathscr{A}$ are also compatible with all maps of $\mathscr{A}_{\text{max}}$.
-It is sufficient to describe smooth manifolds by a non maximal atlas.
+It is sufficient to use a non maximal atlas.
 
@@
 \subsubsection{Lie groups} The study of group theory deals with symmetries which can be expressed algebraically.
-A pair $(M,\bullet)$, consisting of a map $\bullet: M\times M \rightarrow M$ is called group if the map is associative, i.e.
-$x \bullet (y \bullet z) = (x \bullet y) \bullet z$ and has a neutral element so that $x \bullet e_{M} = x$.
+A pair $(M,\circ)$, consisting of a map $\circ: M\times M \rightarrow M$ is called group if the map is associative, i.e.
+$x \circ (y \circ z) = (x \circ y) \circ z$ and has a neutral element so that $x \circ e_{M} = x$.
 We also require an inverse element to each element of the group, i.e.
-$x \bullet x^{-1} = e_{M}$.
-A Lie group is a smooth manifold which is equipped with a group structure such that the maps $\bullet:M \times  M \rightarrow  M, (x,y) \mapsto xy, $ and $\imath: M \rightarrow  M, x \mapsto x^{-1}$ are smooth.
+$x \circ x^{-1} = e_{M}$.
+A Lie group is a smooth manifold which is equipped with a group structure such that the maps $\circ: M \times  M \rightarrow  M, (x,y) \mapsto xy, $ and $\imath: M \rightarrow  M, x \mapsto x^{-1}$ are smooth.
 We call a space \emph{connected} if it cannot be divided into disjoint open neighbourhoods.
@@
 
-We apply a theorem from Lie theory, which states that each connected Abelian Lie group $(M,\bullet)$ of dimension $\dim M = n$ is isomorphic to a product space $\mathbb{R}^p \times \mathbb{T}^q$ with $p+q =n$, for a proof we refer to \cite[p.~116]{onishchik1993lie}.
+We apply a theorem from Lie theory, which states that each connected Abelian Lie group $(M,\circ)$ of dimension $\dim M = n$ is isomorphic to a product space $\mathbb{R}^p \times \mathbb{T}^q$ with $p+q =n$, for a proof we refer to \cite[p.~116]{onishchik1993lie}.
 The $q$-torus $\mathbb{T}^q$ is a surface of revolution.
@@
 These axes are given by $1$-spheres, such that the $q$-torus is a product space $\mathbb{T}^q = \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$.
-The initial decomposition of connected commutative Lie groups can be further simplified into $M \simeq \mathbb{R}^p \times (\mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q)$.
+The initial decomposition of connected commutative Lie groups can be further simplified into $M \simeq \mathbb{R}^p \times \mathcal{S}^1_1 \times \cdots \times \mathcal{S}^1_q$.
 Recall, that the $(n-1)$-sphere is given by $\mathcal{S}^{n-1} := \{x \in \mathbb{R}^n \; \vert \; \norm{x}_2 = 1\}$.
@@
 This defines the simplicial complex as a topological space.
+Examples for simplicial complexes can be found in Fig.
+\ref{figure2} b).
 These properties ensure that simplices are added to the complex in a way, such that they are \emph{glued} edge to edge, point to point, and surface to surface.
@@
 These elements form the boundary group $\langle B_k, +\rangle$, defined by $B_k := \text{im } \partial_{k+1} = \{[\sigma] \in C_k \; \vert \; \exists [\tau] \in C_{k+1}: [\sigma] = \partial_{k+1}[\tau]\}$.
-Now we define the homology groups as quotient of groups.
+The connection of these groups is illustrated in Fig.
+\ref{figure1} a).
+Now we define the homology groups as quotients of groups.
 Homology -- analogous to homotopy theory -- gives information about connected components, loops and higher dimensional holes in the simplicial complex:
@@
 We denote $K^{\epsilon} := K^{\epsilon}(X)$.
-Depending on how we vary the parameter $\epsilon_i$ of the chosen simplicial complex, the following sequence is generated, connected by inclusion and starting with the empty set cf.
-Fig.
-\ref{figure1} b):
+Depending on how we vary the parameter $\epsilon_i$ of the chosen simplicial complex, the following sequence is generated, connected by inclusion:
 \begin{align}
@@
 \end{align}
+An example of a filtration is given in Fig.
+\ref{figure1} b).
 The filtration has a discrete realization with a fixed $\epsilon = \epsilon_{i+1}-\epsilon_{i} = \min(x,y)$ for all $x,y \in X$.
@@
 \end{equation}
-for the $k$-th homology group and index pairs $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2} := \mathbb{R}^2 \cup \infty$ with indices $i \leq j$.
+for the $k$-th homology group and index pairs $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2} := \mathbb{R}^2 \cup \{-\infty,\infty\}$ with indices $i \leq j$.
 The Euclidean space is extended, as the very first connected component on the filtration remains connected.
@@
 \begin{equation}
-\text{PH}_{k}(X) := \left\{(\epsilon_i, \epsilon_{j+1}) \in \overline{\mathbb{R}^2} \; \bigg\vert \; \mu^{\epsilon_{i},\epsilon_{j+1}}_k = 1 \; \text{for all} \; i,j \in I \right\}.
+\text{Ph}_{k}(X) := \left\{(\epsilon_i, \epsilon_{j+1}) \in \overline{\mathbb{R}^2} \; \bigg\vert \; \mu^{\epsilon_{i},\epsilon_{j+1}}_k = 1 \; \text{for all} \; i,j \in I \right\}.
 \end{equation}
@@
 \subsubsection{Persistence landscapes} Persistence landscapes give a statistical summary of the topology of a set of points embedded in a given topological manifold \cite{bubenik2015statistical}.
-Looking at the points $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2}$ on the $k$-th persistence diagram $\text{PH}_{k}(X)$, one associates a piecewise linear function $\lambda^{\epsilon_i}_{\epsilon_{j+1}}: \mathbb{R} \rightarrow [0,\infty)$ with those points:
+Looking at the points $(\epsilon_i,\epsilon_{j+1}) \in \overline{\mathbb{R}^2}$ on the $k$-th persistence diagram $\text{Ph}_{k}(X)$, one associates a piecewise linear function $\lambda^{\epsilon_i}_{\epsilon_{j+1}}: \mathbb{R} \rightarrow [0,\infty)$ with those points:
 \begin{align}
-    &\text{If } x \not\in (\epsilon_i,\epsilon_{j+1}), \quad \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0,\\
-    &\text{if } x \in \left(\epsilon_{i},(\epsilon_i + \epsilon_{j+1})/2 \right], \quad \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = x - \epsilon_{i} \; \text{and}\\
-    &\text{if } x \in \left((\epsilon_i + \epsilon_{j+1})/2,\epsilon_{j+1} \right), \quad \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = \epsilon_{j+1} - x.
+    &\text{If } x \not\in (\epsilon_i,\epsilon_{j+1}), \; \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0,\\
+    &\text{if } x \in \left(\epsilon_{i},(\epsilon_i + \epsilon_{j+1})/2 \right], \; \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = x - \epsilon_{i} \; \text{and}\\
+    &\text{if } x \in \left((\epsilon_i + \epsilon_{j+1})/2,\epsilon_{j+1} \right), \; \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = \epsilon_{j+1} - x.
 \end{align}
-The summaries for the general persistence diagram are the disjoint union of the $k$-th persistence diagrams $\text{PH}(X) := \coprod_{i=0}^{k} \text{PH}_i(X)$.
-A persistence landscape $\text{PL}(X)$, contains the birth-death pairs $(\epsilon_i,\epsilon_{j+1})$, for an $i,j \in \{1, \cdots, n\}$ and is the function sequence $\Lambda_k : \mathbb{R} \rightarrow [0,\infty)$ for a $k \in \mathbb{N}$, where $\Lambda_k(x)$ denotes the $k$-th greatest value of $\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x)$, see Fig.
+The summaries for the general persistence diagram are the disjoint union of the $k$-th persistence diagrams $\text{Ph}(X) := \coprod_{i=0}^{k} \text{Ph}_i(X)$.
+A persistence landscape $\text{Pl}(X)$, contains the birth-death pairs $(\epsilon_i,\epsilon_{j+1})$, for an $i,j \in \{1, \cdots, n\}$ and is the function sequence $\Lambda_k : \mathbb{R} \rightarrow [0,\infty)$ for a $k \in \mathbb{N}$, where $\Lambda_k(x)$ denotes the $k$-th greatest value of $\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x)$, see Fig.
 \ref{figure1} c)4.
 Thus, $\Lambda_k(x) = 0$ if $k > n$.
-$\text{PL}(X)$ lies in a completely normed vector space, suited for statistical computations.
+$\text{Pl}(X)$ lies in a completely normed vector space, suited for statistical computations.
 
@@
 Manifolds can always be immersed and submersed as long as the rank of the Jacobian of the respective map does not change.
-Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural network \cite{HauserR17}.
+Thus, the dimension of the data manifold is equal to the width of the smallest possible layer of a neural network that is able to represent it \cite{HauserR17}.
+In other words, the layer with minimal width which does not loose information about the topological structure.
 
@@
 where $p$ indicates the number of connected components.
-The number of local maxima for a homology group in the persistence landscape is the cardinality of the set of points $(\epsilon_i,\epsilon_{j+1})$ with derivatives $d\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0 \; \text{and} \; d^2 \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) < 0$, see Fig.
+The number of local maxima for a homology group in the persistence landscape is the cardinality of the set of points $(\epsilon_i,\epsilon_{j+1})$ with derivatives $d\lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) = 0 \; \text{and} \; d^2 \lambda^{\epsilon_i}_{\epsilon_{j+1}}(x) < 0$, for an example of these maximum values within a persistence landscape see Fig.
 \ref{figure2} b) c).
@@
 What decomposition allows to neglect the connectedness?
+For example, the persistence diagrams could be used to identify connected components.
+These in return could be represented as independent connected manifolds.
     \item What do other representations of manifolds with a more complicated structure look like?
 For example, can manifolds like the Klein bottle be learned that are not orientable?
-And if so, how can one decide in advance whether or not the approximation to an orientable manifolds makes more sense than to a non orientable one?
+And if so, how can one decide in advance whether or not the approximation to an orientable manifold makes more sense than to a non orientable one?
 \end{itemize}

### 2020-07-14 llncs2e/paper12_anonymous.tex
anon

--- 2020-07-14 llncs2e/paper12.tex
@@
 \hypersetup{
-    pdftitle={Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory},    % title
-    pdfauthor={Luciano Melodia},     % author
-    pdfsubject={Estimate of the Neural Network Dimension},   % subject of the document
-    pdfcreator={Luciano Melodia},   % creator of the document
-    pdfproducer={Luciano Melodia},  % producer of the document
-    pdfkeywords={Embedding dimension} {Parametrization} {Persistent homology} {neural networks} {Manifold learning}, % list of keywords
-    pdfnewwindow=true,
     colorlinks,
@@
 	linkcolor=black,
-	urlcolor=purple
+	urlcolor=blue
 }
@@
 
-\author{Luciano Melodia\orcidID{0000-0002-7584-7287} \\
-\and Richard Lenz\orcidID{0000-0003-1551-4824}}
 %
-\authorrunning{L. Melodia et al.}
 % First names are abbreviated in the running head.

### 2020-07-20 llncs2e/paper12.tex
link mistake2

@@
 %
-\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/ neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
+\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
 %

### 2020-07-20 llncs2e/paper12.tex
typo notation
papers

@@
 A metric describes the geometric properties of a manifold.
-The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie Group.
+The change of coordinate systems and the learned metric tensor of a smooth manifold during backpropagation was formalized \cite{HauserR17}, which shows, that we operate often on a model of a Lie group.
 Further, it has been shown that dense neural networks cannot approximate all functions arbitrarily precisely \cite{Johnson19}.
@@
 \end{align} 
-We count in $PL(X)$ the elements of the $0$th homology group.
+We count in $\text{Pl}(X)$ the elements of the $0$th homology group.
 The amount of elements from higher homology groups correspond to $q$ (Eq.

### 2020-07-21 llncs2e/paper12.tex
typo
paper12

@@
 $2,4,6, \cdots, 784$.
-$392$ neural networks around a hundred times.
+$392$ neural networks have been trained around a hundred times each.
 Each line of Fig.

### 2020-08-01 llncs2e/paper12.tex
acknow

@@
 %
-\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory\thanks{The code can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.}}
+\title{Estimate of the Neural Network Dimension Using Algebraic Topology and Lie Theory}
 %
@@
 
+\subsubsection*{Acknowledgements} We would like to thank Noah Becker for proofreading and anonymous reviewers for their comments and suggestions for improvement.
+
+\subsubsection*{Code \& Data} The implementation, the data sets and experimental results can be found at: \href{https://github.com/karhunenloeve/neural-topology-lie}{https://github.com/karhunenloeve/neural-topology-lie}.
 \bibliographystyle{splncs04}
